### 2023-02-06 Manuscript/manuscriptPOL.tex
Update manuscript

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% This is a (brief) model paper using the achemso class
%% The document class accepts keyval options, which should include
%% the target journal and optionally the manuscript type.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[journal=jacsat,manuscript=article,layout=singlecolumn]{achemso}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional packages needed here.  Only include packages
%% which are essential, to avoid problems later.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{chemformula} % Formula subscripts using \ch{}
\usepackage[T1]{fontenc} % Use modern font encodings
\usepackage{epstopdf}
\usepackage{multirow}
\usepackage{adjustbox}
\usepackage{float}
\usepackage{placeins}
\usepackage{easy-todo}
\usepackage{verbatim}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If issues arise when submitting your manuscript, you may want to
%% un-comment the next line.  This provides information on the
%% version of every file you have used.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\listfiles

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional macros here.  Please use \newcommand* where
%% possible, and avoid layout-changing macros (which are not used
%% when typesetting).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand*\mycommand[1]{\texttt{\emph{#1}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Meta-data block
%% ---------------
%% Each author should be given as a separate \author command.
%%
%% Corresponding authors should have an e-mail given after the author
%% name as an \email command. Phone and fax numbers can be given
%% using \phone and \fax, respectively; this information is optional.
%%
%% The affiliation of authors is given after the authors; each
%% \affiliation command applies to all preceding authors not already
%% assigned an affiliation.
%%
%% The affiliation takes an option argument for the short name.  This
%% will typically be something like "University of Somewhere".
%%
%% The \altaffiliation macro should be used for new address, etc.
%% On the other hand, \alsoaffiliation is used on a per author basis
%% when authors are associated with multiple institutions.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author{Batuhan Kav}
\author{Hanne Antila}
\affiliation{Forschungszentrum Juelich, Germany}
\email{batuhankav@gmail.com}
\author{O. H. Samuli Ollila}
\author{Markus S. Miettinen}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\title[An \textsf{achemso} demo]
%  {A demonstration of the \textsf{achemso} \LaTeX\
%   class\footnote{A footnote for the title}}
\title{NMRLipids VI: Polarizable Force Fields}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some journals require a list of abbreviations or keywords to be
%% supplied. These should be set up here, and will be printed after
%% the title and author information, if needed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\abbreviations{IR,NMR,UV}
%\keywords{American Chemical Society, \LaTeX}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The manuscript does not need to include \maketitle, which is
%% executed automatically.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "tocentry" environment can be used to create an entry for the
%% graphical table of contents. It is given here as some journals
%% require that it is printed as part of the abstract page. It will
%% be automatically moved as appropriate.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{tocentry}

%\end{tocentry}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The abstract environment will automatically gobble the contents
%% if an abstract is not used by the target journal.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

	For the initial project description, please refer to \\ https://github.com/batukav/NMRlipidsVIpolarizableFFs/blob/master/Manuscript/manuscript.pdf

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Conventional classical molecular dynamics (MD) simulations only consider electrostatics as interactions between static point charges, set by molecular dynamics models (force fields) as an input in the beginning of the simulation. The dynamic effects arising from polarizability are ignored or at most enter the simulations in an averaged fashion in the force field parametrization process. However, since decades work has been underway to make MD simulations explicitly polarizable in the hopes of reaching a more accurate description of the system. In the case of biomembranes (lipid bilayers), spesifically, including polarizability the molecular (dielectric) environment varies dramatically while crossing the membrane from the water phase across the dipolar/charged headgroup interface to the hydrophobic tail region and including explicit polarizability has the potential to provide more accurate description of membrane binding, membrane crossing, and behaviour and interactions of molecules residing within membranes, such as membrane protein folding. There certainly is room for improvement: efforts from the NMRlipids opens science community have showed that the current classical lipid models cannot correctly reproduce the conformational ensemble~\cite{}, conformational dynamics~\cite{}, or ion biding~\cite{} within the simulations.

Historically, there has been three main approaches to modelling polarizability: 1) the induced point dipole/multipole model where polarizable point dipoles are placed on chosen sites on a molecule, 2) the classical Drude oscillator model where polarization is modelled by separating a core and a shell charge with a spring which stretches according to the enviroment giving the site a dipole moment, and 3) electronegativity equilization (fluctuating charge) model where atomic charges are no longer constant but can change according to the electronegativities of the molecule atoms and the electric fields from the molecular enviroment. In recent years the first two have surpassed the third in popularity. However, all of these approaches result in an increasing computational cost, e.g, by introducing new types of interactions, more interactions sites, or by requiring a shorter time step in the simulations.

%To avoid polarizability catastrophe where two closeby polarizable sites feed off from each other and the dipole moments grow infinitely

Considering the increasing popularity of both implicitly and explicitly polarizable force fields and the increased computational computational cost that they require, systematic benchmark study polarizable force field quality is in high need. A few published studies reviewing the state of the polarizable force fields for biomolecules did not go beyond citing the results from the existing studies~\cite{inakollu2020polarisable,jin,baker2015polarizable}. Recent publications show that implicit inclusion of polarizability improves cation binding simulations when evaluated against NMR data~\cite{Melcr:2018a, melcr2019improved}. These results suggest that ion binding would be better captured in simulations with polarizable force fields. 
At the moment, there are three main polarizable lipid force fields with explicit electronic polarization: CHARMM-Drude~\cite{li2017drude}, AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable}, and CHARMM-Fluctuating Charge (FQ)~\cite{lucas2012charge}. 
Here we evaluate the performance of two polarizable force fields with lipid parameters available, CHARMM-Drude~\cite{li2017drude} and AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable}, against NMR data. We use NMR order parameters to evaluate  structural properties and cation binding, whereas NMR relaxation rates and correlation times are utilized to benchmark the lipid conformational dynamics. Furthermore, we compare the results to the non-polarizable simulations from the NMRlipids databanks.

\section{Methods}
\subsection{Practical pitfalls of polarizable biomembrane simulations}

\subsubsection{Building a polarizable force field}

The atomic point charges describing the electrostatics in non-polarizable force fields are typically parameterized by fitting to the electrostatic potential energy surface around one or few conformers obtained from ab initio calculations. One then proceeds to build on top of that the parameters for Lennard-Jones interactions and the dihedrals using a mixture of experimental (such as heats of vaporizations) and quantum mechanical information. For a polarizable force fields, this process has to be redone as in the first step the polarizable contribution has to be separated from the "static" one. Furthermore, for the polarization models considered here, an additional cutoff/damping scheme needs to be introduced to prevent polarization catastrophe---a phenomenon where close-by dipoles feed of each others and their dipole moments (and energy) grows into infinity.

Both polarizable lipid force fields considered here use a bottom up approach where parameters are first obtained for smaller molecules or molecular fragments from which the lipid models are then built.  In the CHARMM-Drude model, the polarizabilities, the Thole damping factors used to prevent polarization catastrophe, and the partial charges where determined by fitting to a discretized QM potential energy surface around the molecule. More spesifically, the pertubations of the potential energy surface in the presence of a small charge where observed to separate the dynamic polarizability contribution. They then proceeded to use enthalpies of vaporation along with QM calculations of molecular interaction enegies to obtain the Lennard-Jones parameters, and on top of that dihedral potentials were introduced to reproduce the QM potential energy profiles along a dihedral rotation. Finally, the dihedrals of the lipid head groups where refined targetting the NMR order parameters.

In the Amoeba force field, a set of atomic multipoles were first obtained based on the QM electron densities using Stone's distributed multipole analysis. The atoms were then assigned polarizabilities according to the AMOEBA atom types and   induced dipole contribution was self consistently substracted from the permanent dipoles~\cite{shi2013proteinamoeba}. The charges (monopoles) were then kept constant while the multipoles were refined against the electrostatic potential energy surface around few conformers of the molecular fragments. %Multipoles are allowed to induce only dipoles at polarizable sites outside their immediate environment (polarization group). 
After electrostatic parameters were set, the Lennard-Jones parameters were taken from previous AMOEBA parameterizations of smaller molecules containing same functional groups~\cite{Ren2011polorganic,shi2011hydration}, and then refined against molecule - water (gas phase) interaction energies. After settling all the nonbonded parameters, dihedral potentials were parameterised based on the QM conformational energy profiles.

\begin{comment}
- polarization contribution has to be separated from point charge electrostatics in the parametrization process.
- for the models considered here (Drude, IDP) one has to prevent the polarization catastrophe somehow (Thole damping parameter)
- Brief explanation how the two FFs studied here were built.

Notes on CHARMM-Drude parametrization:
- bottom up approach. Permantent charge, polarizability and screening factors from QM. Polarizability decoupled from permanent by fitting to changes in PES? 
- Non-bonded electrostatics parameters calculated based on QM.
- LJ from enthalpies of vaporization together .with QM data
- Dihedrals from QM scans for model compounds, then refined to according to the order parameters of the headgroup using re-weighted MD and an genetic algorithm.

Notes on Amoeba parametrization (this is NOT well explained in the paper)
-bottom up approach
-Stone distributed multipole analysis to get the multipoles from the QM electron densities. Divided into permanent and induced.
-Then iterative fitting to QM-PES for electrostatics
-WdW from 1) optimization to hydration energy calculations and 2) QM fragment-water gas phase calculations.
-dihedrals optimized to QM potentials
-no mention of the screening for polarization? Do they use some AMOEBA value
\end{comment}
\subsubsection{Using a polarizable force field}

The increased level of detail portrayed by explicit treatment of electronic polarizability comes with certain costs. The first cost is the increased computational complexity reduces the simulation speed. For the Drude force field, this slowdown occurs both because the addition of the Drude particles increases the number of interaction pairs and the employed extended dual-Langevin thermostat requires 1~fs integration time step. Current AMOEBA force field can use a multi time-step integration algorithm where the non-electronic interactions are iterated with a 2~fs time step whereas more computationally unstable polarization terms are iterated with a smaller time step. Yet, this usage of a multi-time step integration scheme does not overcome the slowdown resulting from the computationally demanding polarization calculations. As a result, one experiences almost a ten-fold decrease in the simulation speed while using the polarizable force fields. The second cost is due to the limited software support for both running and analyzing the simulations. Currently only OpenMM supports both AMOEBA and Drude force fields. NAMD can only run the Drude force field whereas TINKER is only valid for AMOEBA force field and it does not support semi-isotropic pressure coupling, which decouples the pressure tensor in $xy$ and $z$-direction and is crucial for correctly evaluating the ion binding properties of the bilayers (\todo{This needs a reference, I remember Samuli mentioning this during a chat}). On the other hand, GROMACS, which has been the most popular MD engine within the bilayer simulation field, does not support any polarizable force field simulations. This is particularly important, as GROMACS also comes with many built-in functions for analyzing the trajectories and many of these functions require a GROMACS tpr file and subsequently cannot be utilized for analyzing the trajectories generated by other MD engines. This hinders the researchers from using well-tested analysis tools creating an overhead during the trajectory analysis while making results more prone to error. Therefore, open-source Python libraries for trajectory analysis like MDAnalysis, MDtraj, and LiPhyphilic (based on MDAnalysis) that can work on multiple trajectory and topology formats are gaining more attention.

Thanks to its recent update, CHARMM-GUI now can generate the topology and input parameters for the Drude force field, which greatly simplifies the employment of this model.

\subsection{Simulations with CHARMM-Drude parameters}

Current versions of the polarizable force fields require "good" starting structures. To achieve this, for every system, 200~ns long MD simulations with the non-polarizable CHARMM36 force field are run. The last frames from these simulations are fed into the CHARMM-GUI Drude Prepper to generate the input structures and OpenMM run files. Each bilayer contains 36 lipids in each leaflet. Simulations are run with the OpenMM MD engine.

The trajectories are publicly available on Zenodo.

\subsection{Simulations with AMOEBA parameters}

AMOEBA parameters are obtained from Githb [ref].

\subsection{Analysis of simulations}
Order parameters are calculated using the script "OrderParameters.py" under\\ "https://github.com/batukav/NMRlipidsVIpolarizableFFs/tree/master/scripts". \\
The order parameters are first calculated for each lipid as an average over time, and then averaged over the lipids. The error estimates are calculated as the standard error of the mean from the second average.
Dihedral distributions are calculated using the script "calcDihedral.py" under\\ "https://github.com/batukav/NMRlipidsVIpolarizableFFs/tree/master/scripts" .\\
The $R_{1}$ rates and effective correlation times, along with the accompanying error estimates, are quantified from the trajectories as elaborated in Ref.~\citenum{Antila2021}. In-house python script available at XXX was used for this purpose.
\subsection{Simulation Details and data availability}
This project is performed using the "NMRLipids Databank" format and all related files, except the trajectories, are available under\\ "https://github.com/batukav/NMRlipidsVIpolarizableFFs/"

% Please add the following required packages to your document preamble:
\newpage
%\begin{adjustbox}{angle=90}
\begin{table}[]
\begin{tabular}{cccccccccc}
	Lipid/counter-ions                & force field  & Ion (M) & N$_{l}$ & N$_{w}$ & N$_{ion}$ & T (K) & t$_{sim}$ (ns) & t$_{analysis}$ (ns) & files \\ \hline
POPC                              & CHARMM-Drude & NA      & NA       & 6400       & 0         & 303    & 500              & 400         &          \cite{kav_batuhan_2021_4604630}    \\ \hline
POPE                              & CHARMM-Drude & NA      & NA       & 6400       & 0         & 303    & 350              & 300         &          \cite{kav_batuhan_2021_4665773}    \\ \hline
	\multirow{4}{*}{POPC:NaCl}        & CHARMM-Drude & 0.350   & 144      & 6400     & 41         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_4683386}   \\
				  & CHARMM-Drude & 0.450   & 144      & 6400     & 51         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_4683398}   \\
				  & CHARMM-Drude & 0.650   & 144      & 6400     & 77         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_4683405}   \\
				  & CHARMM-Drude & 1.0     & 144      & 6400     & 115        & 303   & 500             & 400                  & \cite{kav_batuhan_2020_4683411}   \\ \hline
	\multirow{3}{*}{POPC:CaCl$_{2}$} & CHARMM-Drude & 0.350   & 144      & 6400     & 41         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_4683393}   \\
				  & CHARMM-Drude & 0.450   & 144      & 6400     & 52         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_4683391}   \\
				  & CHARMM-Drude & 0.650   & 144      & 6400     & 76         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_4683394} \\
				  & CHARMM-Drude & 1.0   & 144      & 6400     & 114         & 303   & 500             & 400                  & \cite{kav_batuhan_2021_4738966}

\end{tabular}
\end{table}
%\end{adjustbox}

\section{Results}

\subsection{Evaluation of lipid bilayer properties simulated with polarizable force fields}

\todo{Batuhan: Do we want to have a table showing the area per lipid, volume per lipid etc. type of values from these simulations?}
%\begin{figure}[!hbt]
%	\centering
%	\includegraphics{popc_order_parameters_nosalt.png}
%	\caption{The head group and glycerol backbone order parameters $S_{CH}$ for the POPC simulations without salt.}
%	\label{fig:popc_order_no_salt}
%        \todo{SAMULI: It would be good to put the experimental value also to second hydrogen even though it would be the same as for first.
%          Now it looks like that those values would not be known.} \\
%        \todo{SAMULI: y-axis should be more zoomed in.} \\
%        \todo{SAMULI: Similar figure for POPE (now figure 4) could be here as panel B).}
%\end{figure}

%\begin{figure}[!hbt]
%	\centering
%	\includegraphics{pope_order_parameters.png}
%	\caption{The head group and glycerol backbone order parameters $S_{CH}$ for the POPE simulations.}
%	\label{fig:pope_order}
%        \todo{SAMULI: Move to panel B) in Figure 1.} \\
%        \todo{SAMULI: y-axis should be more zoomed in.} 
%\end{figure}

\begin{figure}[!hbt]
    \centering
    \includegraphics{order_parameters.png}
    \caption{Order parameters calculated from the simulations}
    \label{fig:order_parameters}
    \todo{Batuhan: Exp for DOPC?}
    \todo{Batuhan: Charmm or any other simulation data for DOPC?}
\end{figure}

%\begin{figure}[!hbt]
%	\centering
%	\includegraphics{dihedral_distributions_for_all.png}
%	\caption{The distributions of the torsion angles for the head group atoms}
%	\label{fig:dihedral}
%        \todo{SAMULI: I am not sure if we need this figure at all.
%          Or were you planning to discuss why order parameter response to ions is qualitatively incorrect in Drude PC?
%          If yes, then maybe remove POPE from this figure.
%        }
%\end{figure}

\begin{figure}[!hbt]
	\centering
	\includegraphics{dihedral_distributions_no_salt.png}
	\caption{The distributions of the torsion angles for the head group atoms without the presence of salt.}
	\label{fig:dihedral_no_salt}
        \todo{SAMULI: I am not sure if I see POPE CHARMM36 lines in this figure.
          Also, colors, lines styles and/or widths could be better used to distinguish common and different elements in each line.
        For example, CHARMM36 with thick lines, drude thin lines, POPC black, POPE red, or something like that.}
        \todo{Batuhan: I'll fix the color scheme, add amoeba results, and move it to the SI section}
\end{figure}

\begin{figure}[!hbt]
	\centering
	\includegraphics{correlation_times.eps}
	\caption{R1 times calculated for the POPC system. Experimental values are obtained from Ref.~\cite{antila2020quasi}}
	\label{fig:correlation_times}
        \todo{SAMULI: Are ''No Salt'' and all ion containing systems from Drude? Is it really true that addition of any salt increases the
        values to the same independently on the amount?}
        \todo{HANNE: are you going to add the teffs? You got the lower limit out of them right? Or maybe we can just mention that the correlation times were super slow which compounds to the already heavy computational load (one needs to simulate even longer to get convergence) }
\end{figure}

\clearpage

\subsection{Ion binding to membranes in simulations with polarizable force fields}

%\begin{figure}[!hbt]
%	\centering
%	\includegraphics{popc_order_parameters_for_all.png}
%	\caption{The head group and glycerol backbone order parameters $S_{CH}$ for the POPC simulations. The experimental values on the dashed lines mean that the different isomers 1/2 experimentally could not be resolved.}
%	\label{fig:popc_order}
%        \todo{SAMULI: I am not sure if we need this kind of figure. Systems without ions are in separate figures as well as dependence on ions.}
%\end{figure}

\begin{figure}[!hbt]
	\centering
	\includegraphics{order_parameter_change.png}
	\caption{The change in the head group order parameters with respect to the simulations without salt as a function of the salt concentration.}
	\label{fig:popc_order_parameter_change}
        \todo{SAMULI: y-axis scale should be much more zoomed in here, particularly the difference in experiments between sodium and calcium must be visible.} \\
        \todo{SAMULI: We should also show the ion density profiles along membrane normal and compare the ones with 1000 mM NaCl and 350 mM CaCl$_2$
          (concentration in bulk water) to the best profiles in the literature (Fig. 5 in http://dx.doi.org/10.1021/acs.jpcb.7b12510).
          This will show the actual difference in ion binding to the best performing models available.}
\end{figure}

\begin{figure}[!hbt]
    \centering
    \includegraphics{ion_density_profiles_with_chloride.png}
    \caption{Density profiles alongside the membrane normal for the Drude POPC (first row) and AMOEBA DOPC (second row). Solid and dashed lines are for the systems with the NaCl and CaCl$_{2}$, respectively.}
    \label{fig:ion_density_profiles}
\end{figure}

\clearpage

\subsection{Discussion}

\subsection{Conclusions}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "Acknowledgement" section can be given in all manuscript
%% classes.  This should be given within the "acknowledgement"
%% environment, which will make the correct section or running title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgement}

\end{acknowledgement}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The same is true for Supporting Information, which should use the
%% suppinfo environment.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{suppinfo}
\end{suppinfo}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The appropriate \bibliography command should be placed here.
%% Notice that the class file automatically sets \bibliographystyle
%% and also names the section correctly.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{nmrlipidsPOL.bib,references.bib}

\end{document}

### 2023-06-16 manuscriptPOL.tex
Update organization of the git and pdf
Add figure
Update

--- 2023-02-06 Manuscript/manuscriptPOL.tex
@@
 
+\author{Hanne Antila}
 \author{Batuhan Kav}
-\author{Hanne Antila}
 \affiliation{Forschungszentrum Juelich, Germany}
 \email{batuhankav@gmail.com}
+\author{Markus S. Miettinen}
 \author{O. H. Samuli Ollila}
-\author{Markus S. Miettinen}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The dynamic effects arising from polarizability are ignored or at most enter the simulations in an averaged fashion in the force field parametrization process.
-However, since decades work has been underway to make MD simulations explicitly polarizable in the hopes of reaching a more accurate description of the system.
-In the case of biomembranes (lipid bilayers), spesifically, including polarizability the molecular (dielectric) environment varies dramatically while crossing the membrane from the water phase across the dipolar/charged headgroup interface to the hydrophobic tail region and including explicit polarizability has the potential to provide more accurate description of membrane binding, membrane crossing, and behaviour and interactions of molecules residing within membranes, such as membrane protein folding.
-There certainly is room for improvement: efforts from the NMRlipids opens science community have showed that the current classical lipid models cannot correctly reproduce the conformational ensemble~\cite{}, conformational dynamics~\cite{}, or ion biding~\cite{} within the simulations.
+However, since decades work~\cite{Thole1981,ando2001stable,Grossfield2003,
+lamoreux2003,Antila2013 ,Lemkul2016, baker2015polarizable,jing2019polarizable } has been underway to make MD simulations explicitly polarizable in the hopes of reaching a more accurate description of the system.
 
-Historically, there has been three main approaches to modelling polarizability: 1) the induced point dipole/multipole model where polarizable point dipoles are placed on chosen sites on a molecule, 2) the classical Drude oscillator model where polarization is modelled by separating a core and a shell charge with a spring which stretches according to the enviroment giving the site a dipole moment, and 3) electronegativity equilization (fluctuating charge) model where atomic charges are no longer constant but can change according to the electronegativities of the molecule atoms and the electric fields from the molecular enviroment.
+In the case of biomembranes (lipid bilayers), specifically, the molecular (dielectric) environment varies dramatically while crossing the membrane from the water phase across the dipolar/charged headgroup interface to the hydrophobic tail region.
+Including explicit polarizability in biomembrane modelling has the potential to provide more accurate description of membrane binding, membrane crossing, and behaviour and interactions of molecules residing within membranes, such as membrane protein folding.
+There certainly is room for improvement: efforts from the NMRlipids open science community have shown that the current non-polarizable lipid models cannot correctly reproduce the conformational ensemble~\cite{Botan2015,Antila2019}, conformational dynamics~\cite{Antila2021}, or ion biding~\cite{Catte2016} within MD simulations.
+
+Historically, there has been three main approaches to modelling polarizability: 1) the induced point dipole/multipole model where polarizable point dipoles are placed on chosen sites on a molecule, 2) the classical Drude oscillator model where polarization is modelled by separating a core and a shell charge with a spring which stretches according to the enviroment giving the site a dipole moment, and 3) the electronegativity equilization (fluctuating charge) model where atomic charges are no longer constant but can change according to the electronegativities of the molecule atoms and the electric fields from the molecular enviroment.
 In recent years the first two have surpassed the third in popularity.
@@
 
-Considering the increasing popularity of both implicitly and explicitly polarizable force fields and the increased computational computational cost that they require, systematic benchmark study polarizable force field quality is in high need.
-A few published studies reviewing the state of the polarizable force fields for biomolecules did not go beyond citing the results from the existing studies~\cite{inakollu2020polarisable,jin,baker2015polarizable}.
-Recent publications show that implicit inclusion of polarizability improves cation binding simulations when evaluated against NMR data~\cite{Melcr:2018a, melcr2019improved}.
-These results suggest that ion binding would be better captured in simulations with polarizable force fields.
-At the moment, there are three main polarizable lipid force fields with explicit electronic polarization: CHARMM-Drude~\cite{li2017drude}, AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable}, and CHARMM-Fluctuating Charge (FQ)~\cite{lucas2012charge}.
-Here we evaluate the performance of two polarizable force fields with lipid parameters available, CHARMM-Drude~\cite{li2017drude} and AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable}, against NMR data.
+Considering the increasing popularity of both implicitly and explicitly polarizable force fields and the increased computational cost that they require, a systematic benchmark study on polarizable force field quality is in great need.
+The few published studies reviewing the state of polarizable force fields for biomolecule simulations did not go beyond citing results from the existing studies~\cite{inakollu2020polarisable,jing2019polarizable,baker2015polarizable}.
+Recent publications show that \textit{implicit} inclusion of polarizability improves cation binding simulations when evaluated against NMR data~\cite{Melcr:2018a, melcr2019improved}, which suggest that including polarization explicitly could result in better description of ion binding as well.
+Currently, there are three main polarizable lipid force fields with explicit electronic polarization: CHARMM-Drude~\cite{li2017drude}, AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable}, and CHARMM-Fluctuating Charge (FQ)~\cite{lucas2012charge}.
+Here we evaluate the performance of two actively developed polarizable force fields with lipid parameters available, CHARMM-Drude~\cite{li2017drude} and AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable}, against NMR data.
 We use NMR order parameters to evaluate  structural properties and cation binding, whereas NMR relaxation rates and correlation times are utilized to benchmark the lipid conformational dynamics.
-Furthermore, we compare the results to the non-polarizable simulations from the NMRlipids databanks.
+Small angle scattering (SAXS) data is used to estimate the description of electron density profile of the membrane.
+Furthermore, we compare the results to the best performing non-polarizable simulations from the NMRlipids Databank\cite{Databank}.
+
+This project has been performed within the NMRLipids collaboration (https://databank.nmrlipids.fi/)~\cite{Databank}.
+All simulation data, analysis scripts, and results are publicly available in their corresponding links.
+No plausible nor implausible request is required to download, use, and disseminate the data.
+
+\section{Results}
+
+\subsection{Evaluation of lipid bilayer properties simulated with polarizable force fields}
+
+We start by exploring the performance of polarizable force fields using the quality metrics developed in the NMRlipids Databank (https://databank.nmrlipids.fi/)~\cite{Databank}---a quality evaluated databank for lipid simulations.
+These metrics describe the ability of the models to reproduce the C-H NMR order parameters (quantifying the conformations) and the SAXS data (measuring the electron density variation across the membrane).
+For the former the molecule is divided into three parts, one containing the headgroup and the glycerol backbone carbons and one for each of the carbonyl tails, and the quality of each is evaluated separately.
+The quality metrics are summarized in Table~\ref{tab:my_label} and the performance of the force fields is visualized along with the underlying experimental measurables in Fig~\ref{fig:order_parameters}.
+
+One can clearly see that the non-polarizable force fields, CHARMM36 and OPLS3e, outperform the polarizable force fields with respect to the NMR measurables for all parts of the molecule (higher $P$-values).
+For CHARMM-Drude force field, this is likely due to the headgroup and glycerol parametrization having been done targetting absolute average values of the order parameters instead of taking account the potential order parameter forking and order parameter sign~\cite{Antila2022}.
+The Amoeba force field reproduces the headgroup order parameters well [\todo{HANNE: refer to the hg flexibility?
+@Samuli, I didn't quite understand the argument?
+Refer there to the dihedral figure}] but clearly fails to depict the tail conformations, in particular in the double bond region (Fig.~\ref{fig:order_parameters}).
+In the tail regions, the CHARMM-Drude model is clearly better out of the two polarizable force fields.
+
+The higher $FFq$ values (Table~\ref{tab:my_label}) of the polarizable force fields indicate that they are not able to reproduce the SAXS data as well as the best non-polarizable force fields.
+That said, CHARMM-Drude is as good as (POPC) or better (POPE) than its non-polarizable counterpart CHARMM36.
+AMOEBA, on the other hand, captures the SAXS profile of POPE very poorly (Table~\ref{tab:my_label} and Fig.~\ref{fig:order_parameters}D top).
+Note that the method for estimating the charge density of bilayers from the simulations places electrons as point charges at the atom centers in the system and does not consider any redistribution of charge density due to the polarizability.
+We expect this to results into a good estimate of the charge density.
+
+We proceed to quantify (Fig.~\ref{fig:correlation_times}) the structural dynamics of the polarizable models againts NMR $R_{1}$ rates, measuring a window of dynamics around $\sim$0.1-1~ns, and effective correlation times $\tau_{\mathrm{eff}}$ which quantifies a larger range of dynamical processes up to $\sim$1000~ns.
+We include the same measurables quantified from non-polarizable CHARMM36 and SLipids force fields, which are the best MD force fields regarding the dynamics~\cite{Antila2021}, as a point of comparison.
+The Amoeba force field captures the dynamics with respect to both measurables well,  in par with the best non-polarizable models.
+In contrast, the dynamics of CHARMM-Drude appears to be faulty in the fast range ($R_{1}$ rates) and, furthermore, $\tau_{\mathrm{eff}}$ values reveal vastly too slow overall dynamics.
+This indicates that not only are the simulations with this polarizable force field computationally costly but one would have to simulate longer to obtain converged results.
+Note that the experimental data here is from POPC.
+However, chemically identical the headgroup region the dynamics is assumed to be very similar in DOPC, due to the decoupling of the dynamics between the headgroup and the glycerol regions~\cite{Antila2022rot, Klauda08}.
+
+Finally, we address the premise that adding polarizability to force fields improves the depiction of ion binding to biomembranes in simulations.
+For this, we use the concept of "lipid electrometer"~\cite{Catte2016} where the amount of ion binding to the membrane can be quantified by the change in $\alpha$ and $\beta$ order parameters upon increasing salt concentration.
+These changes in order parameters are depicted in Fig.~\ref{fig:popc_order_parameter_change} whereas Fig.~\ref{fig:ion_density_profiles} gives the density profiles of ions with respect to the bilayer normal.
+The CHARMM36-Drude appears to have strong sodium binding, evidenced both by large change in the order parameters and the ion density profiles.
+The binding of monovalent sodium is, in fact, stronger than the divalent calcium.
+In strong contrast, experimentally one sees no sodium binding and some binding of calcium ions.
+
+As opposed to the CHARMM36-Drude model, the order parameters of the AMOEBA model are almost completely unaffected by the ion addition, both for NaCl and CaCl$_{2}$.
+The ion density profiles confirm that very little ions penetrate the membrane region where the lipid phosphates reside.
+While this is in line with experiments regarding sodium ions, there should be some binding of calcium ions to the membrane.
+Again, the non-polarizable CHARMM36 appears to depict this calcium binding the best out of the models compared here.
+
+%\begin{figure}[!hbt]
+%	\centering
+%	\includegraphics{popc_order_parameters_nosalt.png}
+%	\caption{The head group and glycerol backbone order parameters $S_{CH}$ for the POPC simulations without salt.}
+%	\label{fig:popc_order_no_salt}
+%        \todo{SAMULI: It would be good to put the experimental value also to second hydrogen even though it would be the same as for first.
+%          Now it looks like that those values would not be known.} \\
+%        \todo{SAMULI: y-axis should be more zoomed in.} \\
+%        \todo{SAMULI: Similar figure for POPE (now figure 4) could be here as panel B).}
+%\end{figure}
+
+%\begin{figure}[!hbt]
+%	\centering
+%	\includegraphics{pope_order_parameters.png}
+%	\caption{The head group and glycerol backbone order parameters $S_{CH}$ for the POPE simulations.}
+%	\label{fig:pope_order}
+%        \todo{SAMULI: Move to panel B) in Figure 1.} \\
+%        \todo{SAMULI: y-axis should be more zoomed in.} 
+%\end{figure}
+
+\begin{figure}[!hbt]
+    \centering
+    %\includegraphics{Figures/order_parameters.png}
+    \includegraphics[width=\textwidth]{Figures/quality.pdf}
+    \caption{X-ray scattering form factors, and C-H bond order parameters for headgroup and glycerol backbone, and acyl chains (from top to bottom) compared between simulations and experiments using the NMRlipids databank.}
+    \label{fig:order_parameters}
+
+\end{figure}
+
+\begin{table}[]
+    \centering
+    \begin{tabular}{c c c c c c c}
+        Lipid & Force field & $P^{\mathrm{headgroup}}$ & $P^{sn-1}$ & $P^{sn-2}$ & $FF_{q}$ & APL \\
+        \hline
+        POPC & OPLS3e       & 0.76 & 0.87 & 0.85 & 0.15 & 66.5\\
+        POPC & CHARMM36     & 0.70 & 0.54 & 0.69 & 1.16 & 65.0\\
+        POPC & CHARMM-Drude & 0.51 & 0.22 & 0.47 & 1.16 & 62.5\\
+        \hline
+        POPE & GROMOS-CKP & 0.29 & 0.83 & 0.48 & 0.40 & 59.6\\
+        POPE & CHARMM36   & 0.54 & 0.52 & 0.27 & 1.30 & 57.2 \\
+        POPE & CHARMM-Drude & 0.06 & 0.53 & 0.27 & 0.80 & 56.6  \\
+        POPE & AMOEBA & 0.21 & 0.10 & 0.23 & 3.80 & 66.9\\
+        \hline
+        DOPC & AMOEBA & 0.60 & 0.60 & 0.54 & - & 70.2\\
+    \end{tabular}
+    \caption{NMRlipids databank quality evaluations~\cite{Databank} and area per lipids compared between the best simulations found from the NMRlipids databank, simulations with CHARMM36 force field, and simulations with polarizable force fields.
+Experimental estimations for areas per lipid are POPC: 64.3$\pm1$\AA$^{2}$\cite{kucerka2011}, DOPC: 67.5$\pm1$\AA$^{2}$\cite{kucerka2008}, and POPE: 56.7$\pm$3\AA$^{2}$\cite{Rickeard2020}}
+    \label{tab:my_label}
+\end{table}
+
+%\begin{figure}[!hbt]
+%	\centering
+%	\includegraphics{dihedral_distributions_for_all.png}
+%	\caption{The distributions of the torsion angles for the head group atoms}
+%	\label{fig:dihedral}
+%        \todo{SAMULI: I am not sure if we need this figure at all.
+%          Or were you planning to discuss why order parameter response to ions is qualitatively incorrect in Drude PC?
+%          If yes, then maybe remove POPE from this figure.
+%        }
+%\end{figure}
+
+\begin{figure}[!hbt]
+	\centering
+	\includegraphics{Figures/dihedral_distributions_without_ions.eps}
+	\caption{The distributions of the torsion angles for the head group atoms.
+For each torsion angle, the upper and lower rows contain the DOPC (AMOEBA) - POPC (Drude/Drude2023) and POPE lipids, respectively.}
+	\label{fig:dihedral_no_salt}
+\end{figure}
+
+\begin{figure}[!hbt]
+	\centering
+	\includegraphics{Figures/both_log.pdf}
+	\caption{Effective correlation times (top) and R$_1$ rates (bottom) for the polarizable models and best performing non-polarizable force fields.
+Note, top panel axis is logarithmic to visualize the slow CHARMM-drude dynamics.
+Experimental values are obtained from Ref.~\citenum{Antila2022rot}.
+All simulations used here were salt free.}
+	\label{fig:correlation_times}
+
+\end{figure}
+
+\clearpage
+
+\subsection{Ion binding to membranes in simulations with polarizable force fields}
+
+%\begin{figure}[!hbt]
+%	\centering
+%	\includegraphics{popc_order_parameters_for_all.png}
+%	\caption{The head group and glycerol backbone order parameters $S_{CH}$ for the POPC simulations.
+The experimental values on the dashed lines mean that the different isomers 1/2 experimentally could not be resolved.}
+%	\label{fig:popc_order}
+%        \todo{SAMULI: I am not sure if we need this kind of figure. Systems without ions are in separate figures as well as dependence on ions.}
+%\end{figure}
+
+\begin{figure}[!hbt]
+	\centering
+	\includegraphics{Figures/order_parameter_change.png}
+	\caption{The change in the lipid head group order parameters $\beta$ (top row) $\alpha$ (bottom row) upon increasing ion concentration with respect to the simulations without salt. }
+	\label{fig:popc_order_parameter_change}
+\end{figure}
+
+\begin{figure}[!hbt]
+    \centering
+    \includegraphics{Figures/ion_density_profiles_with_chloride.eps}
+    \caption{Density profiles along the membrane normal for the Drude 2017 and 2023 POPC (first and second rows) and AMOEBA DOPC (third row).
+Solid and dashed lines are for the systems with the NaCl and CaCl$_{2}$, respectively.}
+    \label{fig:ion_density_profiles}
+            \todo{SAMULI: We should also show the ion density profiles along membrane normal and compare the ones with 1000 mM NaCl and 350 mM CaCl$_2$
+          (concentration in bulk water) to the best profiles in the literature (Fig. 5 in http://dx.doi.org/10.1021/acs.jpcb.7b12510).
+          This will show the actual difference in ion binding to the best performing models available.}
+          \todo{Batuhan: Data generated from http://dx.doi.org/10.1021/acs.jpcb.7b12510}
+\end{figure}
+
+\clearpage
+
+%\subsection{Discussion}
+
+\subsection{Conclusions}
+After comparing the polarizable lipid models against high-resolution experimental data and the best performing non-polarizable force fields, we cannot confirm that adding polarizablity to the force fields has yet solved many of the issues it was hoped to address.
+Ion binding to the model membrane seems to be overestimated by the CHARMM36-Drude model, whereas the calcium binding is underestimated by the Amoeba force field.
+The electron density profiles of the bilayers appear to be better reproduced by the best non-polarizable force fields, within the confines of the approximations used to calculate them from the trajectory.
+Same applies for the dynamics and conformational sampling of the models, although the Amoeba force field produces results, spesifically in the headgroup region of PC lipids, that are in par or better than the best non-polarizable force fields with respect to these measurables.
+In summary, polarizable models appear still to be in their infancy, and added complexity of polarizability has not still realised its potential to the fullest; The non-polarizable force fields with longer and more intense refinement behind them still outpeform.
+Considering the complexity and cost of using these models, polarizability should be parameterized with more thorough consideration to the experimental benchmarks.
 
@@
 For a polarizable force fields, this process has to be redone as in the first step the polarizable contribution has to be separated from the "static" one.
-Furthermore, for the polarization models considered here, an additional cutoff/damping scheme needs to be introduced to prevent polarization catastrophe---a phenomenon where close-by dipoles feed of each others and their dipole moments (and energy) grows into infinity.
+Furthermore, for the polarization models considered here, an additional cutoff/damping scheme needs to be introduced to prevent polarization catastrophe~\cite{Thole1981}---a phenomenon where close-by dipoles feed of each others and their dipole moments (and energy) grows into infinity.
 
@@
 In the CHARMM-Drude model, the polarizabilities, the Thole damping factors used to prevent polarization catastrophe, and the partial charges where determined by fitting to a discretized QM potential energy surface around the molecule.
-More spesifically, the pertubations of the potential energy surface in the presence of a small charge where observed to separate the dynamic polarizability contribution.
-They then proceeded to use enthalpies of vaporation along with QM calculations of molecular interaction enegies to obtain the Lennard-Jones parameters, and on top of that dihedral potentials were introduced to reproduce the QM potential energy profiles along a dihedral rotation.
+More specifically, the perturbations of the potential energy surface in the presence of a small charge where observed to separate the dynamic polarizability contribution.
+They then proceeded to use enthalpies of vaporation along with QM calculations of molecular interaction energies to obtain the Lennard-Jones parameters, and on top of that dihedral potentials were introduced to reproduce the QM potential energy profiles along a dihedral rotation.
 Finally, the dihedrals of the lipid head groups where refined targetting the NMR order parameters.
 
-In the Amoeba force field, a set of atomic multipoles were first obtained based on the QM electron densities using Stone's distributed multipole analysis.
-The atoms were then assigned polarizabilities according to the AMOEBA atom types and   induced dipole contribution was self consistently substracted from the permanent dipoles~\cite{shi2013proteinamoeba}.
+In the Amoeba force field, a set of atomic multipoles were first obtained based on the QM electron densities using Stone's distributed multipole analysis~\cite{Stone1981}.
+The atoms were then assigned polarizabilities according to the AMOEBA atom types and induced dipole contribution was self consistently subtracted from the permanent dipoles~\cite{shi2013proteinamoeba}.
 The charges (monopoles) were then kept constant while the multipoles were refined against the electrostatic potential energy surface around few conformers of the molecular fragments.
@@
 Currently only OpenMM supports both AMOEBA and Drude force fields.
-NAMD can only run the Drude force field whereas TINKER is only valid for AMOEBA force field and it does not support semi-isotropic pressure coupling, which decouples the pressure tensor in $xy$ and $z$-direction and is crucial for correctly evaluating the ion binding properties of the bilayers (\todo{This needs a reference, I remember Samuli mentioning this during a chat}).
-On the other hand, GROMACS, which has been the most popular MD engine within the bilayer simulation field, does not support any polarizable force field simulations.
+NAMD can only run the Drude force field whereas TINKER is only compatible with AMOEBA force field and does not support semi-isotropic pressure coupling, which decouples the pressure tensor in $xy$ and $z$-direction and is crucial for correctly evaluating the ion binding properties of the bilayers (\todo{This needs a reference, I remember Samuli mentioning this during a chat}).
+On the other hand, GROMACS, which has been the most popular MD engine within the bilayer simulation field, only has limited support for the Drude polarizable force field via an unofficial git-branch [https://github.com/gromacs/gromacs/tree/drude] but does not support AMOEBA at all.
+\begin{comment}
 This is particularly important, as GROMACS also comes with many built-in functions for analyzing the trajectories and many of these functions require a GROMACS tpr file and subsequently cannot be utilized for analyzing the trajectories generated by other MD engines.
@@
 Therefore, open-source Python libraries for trajectory analysis like MDAnalysis, MDtraj, and LiPhyphilic (based on MDAnalysis) that can work on multiple trajectory and topology formats are gaining more attention.
-
+\end{comment}
 Thanks to its recent update, CHARMM-GUI now can generate the topology and input parameters for the Drude force field, which greatly simplifies the employment of this model.
@@
 
-AMOEBA parameters are obtained from Githb [ref].
+AMOEBA force field parameters for OpenMM are obtained from GitHub [https://github.com/Inniag/openmm-scripts-amoeba,~\cite{klesse2020induced}].
+The initial bilayer structures, similar to our simulations with the Drude2017 force field, are initially obtained from CHARMM-GUI membrane builder and equilibriated with the CHARMM36 force field for 200~ns.
 
@@
 \begin{table}[]
+\begin{small}
 \begin{tabular}{cccccccccc}
-	Lipid/counter-ions                & force field  & Ion (M) & N$_{l}$ & N$_{w}$ & N$_{ion}$ & T (K) & t$_{sim}$ (ns) & t$_{analysis}$ (ns) & files \\ \hline
+	Lipid/ion                & force field  & Ion (M) & N$_{l}$ & N$_{w}$ & N$_{ion}$ & T (K) & t$_{sim}$ (ns) & t$_{anal}$ (ns) & files \\ \hline
 POPC                              & CHARMM-Drude & NA      & NA       & 6400       & 0         & 303    & 500              & 400         &          \cite{kav_batuhan_2021_4604630}    \\ \hline
@@
 \end{tabular}
+\end{small}
+\caption{Systems simulated for this work.
+Column N$_l$ gives the number of lipids, N$_w$ gives the number of water molecules, T(K) denotes the temperature in kelvins.
+Simulated time is listed in column t$_{sim}$ and time used for analysis in t$_{anal}$.
+Column "files" gives the reference to open access simulation data.}
 \end{table}
 %\end{adjustbox}
-
-\section{Results}
-
-\subsection{Evaluation of lipid bilayer properties simulated with polarizable force fields}
-
-\todo{Batuhan: Do we want to have a table showing the area per lipid, volume per lipid etc. type of values from these simulations?}
-%\begin{figure}[!hbt]
-%	\centering
-%	\includegraphics{popc_order_parameters_nosalt.png}
-%	\caption{The head group and glycerol backbone order parameters $S_{CH}$ for the POPC simulations without salt.}
-%	\label{fig:popc_order_no_salt}
-%        \todo{SAMULI: It would be good to put the experimental value also to second hydrogen even though it would be the same as for first.
-%          Now it looks like that those values would not be known.} \\
-%        \todo{SAMULI: y-axis should be more zoomed in.} \\
-%        \todo{SAMULI: Similar figure for POPE (now figure 4) could be here as panel B).}
-%\end{figure}
-
-%\begin{figure}[!hbt]
-%	\centering
-%	\includegraphics{pope_order_parameters.png}
-%	\caption{The head group and glycerol backbone order parameters $S_{CH}$ for the POPE simulations.}
-%	\label{fig:pope_order}
-%        \todo{SAMULI: Move to panel B) in Figure 1.} \\
-%        \todo{SAMULI: y-axis should be more zoomed in.} 
-%\end{figure}
-
-\begin{figure}[!hbt]
-    \centering
-    \includegraphics{order_parameters.png}
-    \caption{Order parameters calculated from the simulations}
-    \label{fig:order_parameters}
-    \todo{Batuhan: Exp for DOPC?}
-    \todo{Batuhan: Charmm or any other simulation data for DOPC?}
-\end{figure}
-
-%\begin{figure}[!hbt]
-%	\centering
-%	\includegraphics{dihedral_distributions_for_all.png}
-%	\caption{The distributions of the torsion angles for the head group atoms}
-%	\label{fig:dihedral}
-%        \todo{SAMULI: I am not sure if we need this figure at all.
-%          Or were you planning to discuss why order parameter response to ions is qualitatively incorrect in Drude PC?
-%          If yes, then maybe remove POPE from this figure.
-%        }
-%\end{figure}
-
-\begin{figure}[!hbt]
-	\centering
-	\includegraphics{dihedral_distributions_no_salt.png}
-	\caption{The distributions of the torsion angles for the head group atoms without the presence of salt.}
-	\label{fig:dihedral_no_salt}
-        \todo{SAMULI: I am not sure if I see POPE CHARMM36 lines in this figure.
-          Also, colors, lines styles and/or widths could be better used to distinguish common and different elements in each line.
-        For example, CHARMM36 with thick lines, drude thin lines, POPC black, POPE red, or something like that.}
-        \todo{Batuhan: I'll fix the color scheme, add amoeba results, and move it to the SI section}
-\end{figure}
-
-\begin{figure}[!hbt]
-	\centering
-	\includegraphics{correlation_times.eps}
-	\caption{R1 times calculated for the POPC system. Experimental values are obtained from Ref.~\cite{antila2020quasi}}
-	\label{fig:correlation_times}
-        \todo{SAMULI: Are ''No Salt'' and all ion containing systems from Drude? Is it really true that addition of any salt increases the
-        values to the same independently on the amount?}
-        \todo{HANNE: are you going to add the teffs?
-You got the lower limit out of them right?
-Or maybe we can just mention that the correlation times were super slow which compounds to the already heavy computational load (one needs to simulate even longer to get convergence) }
-\end{figure}
-
-\clearpage
-
-\subsection{Ion binding to membranes in simulations with polarizable force fields}
-
-%\begin{figure}[!hbt]
-%	\centering
-%	\includegraphics{popc_order_parameters_for_all.png}
-%	\caption{The head group and glycerol backbone order parameters $S_{CH}$ for the POPC simulations.
-The experimental values on the dashed lines mean that the different isomers 1/2 experimentally could not be resolved.}
-%	\label{fig:popc_order}
-%        \todo{SAMULI: I am not sure if we need this kind of figure. Systems without ions are in separate figures as well as dependence on ions.}
-%\end{figure}
-
-\begin{figure}[!hbt]
-	\centering
-	\includegraphics{order_parameter_change.png}
-	\caption{The change in the head group order parameters with respect to the simulations without salt as a function of the salt concentration.}
-	\label{fig:popc_order_parameter_change}
-        \todo{SAMULI: y-axis scale should be much more zoomed in here, particularly the difference in experiments between sodium and calcium must be visible.} \\
-        \todo{SAMULI: We should also show the ion density profiles along membrane normal and compare the ones with 1000 mM NaCl and 350 mM CaCl$_2$
-          (concentration in bulk water) to the best profiles in the literature (Fig. 5 in http://dx.doi.org/10.1021/acs.jpcb.7b12510).
-          This will show the actual difference in ion binding to the best performing models available.}
-\end{figure}
-
-\begin{figure}[!hbt]
-    \centering
-    \includegraphics{ion_density_profiles_with_chloride.png}
-    \caption{Density profiles alongside the membrane normal for the Drude POPC (first row) and AMOEBA DOPC (second row).
-Solid and dashed lines are for the systems with the NaCl and CaCl$_{2}$, respectively.}
-    \label{fig:ion_density_profiles}
-\end{figure}
-
-\clearpage
-
-\subsection{Discussion}
-
-\subsection{Conclusions}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\todo{batuhan: update the bib file to have 10ps fixed dt trajectories}
 \bibliography{nmrlipidsPOL.bib,references.bib}

### 2023-08-11 manuscriptPOL.tex
Updates from Overleaf

@@
 \section{Introduction}
+
+\textcolor{red}{some ideas for the intro/conclusion: 1) amoeba and drude are mostly used for transmembrane protein simulations.
+https://pubs.acs.org/doi/abs/10.1021/acsnano.1c06443 https://www.sciencedirect.com/science/article/abs/pii/S0959440X2100110X https://pubs.rsc.org/en/content/articlehtml/2021/sc/d1sc01887f etc.
+Validation/comparison of these simulations with the experiments are done for the proteins, not the lipids.
+Now since we know that the lipid dynamics and conformations are not very correct with the polarizable force fields, can these affect the protein conformations?
+is there a way to possibly connect this to the previous nmrlipids jacs paper?\\
+}
+
 Conventional classical molecular dynamics (MD) simulations only consider electrostatics as interactions between static point charges, set by molecular dynamics models (force fields) as an input in the beginning of the simulation.
@@
 Historically, there has been three main approaches to modelling polarizability: 1) the induced point dipole/multipole model where polarizable point dipoles are placed on chosen sites on a molecule, 2) the classical Drude oscillator model where polarization is modelled by separating a core and a shell charge with a spring which stretches according to the enviroment giving the site a dipole moment, and 3) the electronegativity equilization (fluctuating charge) model where atomic charges are no longer constant but can change according to the electronegativities of the molecule atoms and the electric fields from the molecular enviroment.
-In recent years the first two have surpassed the third in popularity.
-However, all of these approaches result in an increasing computational cost, e.g, by introducing new types of interactions, more interactions sites, or by requiring a shorter time step in the simulations.
+In lipid modelling, the first two have surpassed the third in popularity.
+All of these approaches result in an increasing computational cost, e.g, by introducing new types of interactions, more interactions sites, or by requiring a shorter time step in the simulations.
 
@@
 
-Considering the increasing popularity of both implicitly and explicitly polarizable force fields and the increased computational cost that they require, a systematic benchmark study on polarizable force field quality is in great need.
+Considering the growing popularity of both implicitly and explicitly polarizable force fields and the increased computational cost that they require, a systematic benchmark study on polarizable force field quality is in great need.
 The few published studies reviewing the state of polarizable force fields for biomolecule simulations did not go beyond citing results from the existing studies~\cite{inakollu2020polarisable,jing2019polarizable,baker2015polarizable}.
@@
 
-\section{Results}
+\section{Results and Discussion}
 
@@
 
-One can clearly see that the non-polarizable force fields, CHARMM36 and OPLS3e, outperform the polarizable force fields with respect to the NMR measurables for all parts of the molecule (higher $P$-values).
-For CHARMM-Drude force field, this is likely due to the headgroup and glycerol parametrization having been done targetting absolute average values of the order parameters instead of taking account the potential order parameter forking and order parameter sign~\cite{Antila2022}.
-The Amoeba force field reproduces the headgroup order parameters well [\todo{HANNE: refer to the hg flexibility?
-@Samuli, I didn't quite understand the argument?
-Refer there to the dihedral figure}] but clearly fails to depict the tail conformations, in particular in the double bond region (Fig.~\ref{fig:order_parameters}).
-In the tail regions, the CHARMM-Drude model is clearly better out of the two polarizable force fields.
+One can clearly see that the non-polarizable force fields, CHARMM36 and OPLS3e, consistently outperform the polarizable force fields with respect to the NMR measurables for all parts of the molecule (higher $P$-values).
+For older CHARMM-Drude force field, this is likely due to the headgroup and glycerol parametrization having been done targetting absolute average values of the order parameters instead of taking account the potential order parameter forking ( measurably different order parameters for different C--H bods at one carbon) and order parameter sign~\cite{Antila2022}.
+This was then remedied in the newer version of the model, which also provides better description of the carbon tails.
+In fact, for POPE sn-2 tail the CHARMM-Drude2023 gives the best description of the models investigated here.
+The AMOEBA force field reproduces the most of the headgroup order parameters well for both DOPC and POPE but vastly overestimates one of the g$_3$ values, resulting in too large forking of oder parameters in this particular carbon.
+AMOEBA also clearly struggles to depict the tail conformations, in particular in the double bond region of DOPC (Fig.~\ref{fig:order_parameters}).
+In the tail regions, the CHARMM-Drude2023 appears to be the best out of the polarizable models.
 
-The higher $FFq$ values (Table~\ref{tab:my_label}) of the polarizable force fields indicate that they are not able to reproduce the SAXS data as well as the best non-polarizable force fields.
-That said, CHARMM-Drude is as good as (POPC) or better (POPE) than its non-polarizable counterpart CHARMM36.
-AMOEBA, on the other hand, captures the SAXS profile of POPE very poorly (Table~\ref{tab:my_label} and Fig.~\ref{fig:order_parameters}D top).
-Note that the method for estimating the charge density of bilayers from the simulations places electrons as point charges at the atom centers in the system and does not consider any redistribution of charge density due to the polarizability.
-We expect this to results into a good estimate of the charge density.
+The higher $FFq$ values (Table~\ref{tab:my_label}) of the polarizable force fields for POPC indicate that they are not able to reproduce the SAXS data quite as well as the best non-polarizable force field, OPLS3e.
+That said, CHARMM-Drude2023, in particular, is clearly better than its non-polarizable counterpart CHARMM36 and the best out of all the models here in capturing the POPE SAXS data.
+AMOEBA, on the other hand, captures the SAXS profile of POPE very poorly (Table~\ref{tab:my_label} and Fig.~\ref{fig:order_parameters}D top) which is likely connected to the overall description of the lipid tails in the force fields.
+Note that owing the simulation size dependency of the SAXS data~\cite{Databank} one should focus on the extrema $q_{z}$[\AA$^{-1}$] when comparing the simulated and experimental data.
+In particular $FFq$ only measures the difference between the first minima of the simulated and experimental curves.
+Also note, that the method for estimating the charge density of bilayers from the simulations places electrons as point charges at the atom centers in the system and does not consider any redistribution of charge density due to the polarizability.
+Nevertheless, we expect this results into a good estimate of the charge density for the calculation of the SAXS curve.
 
-We proceed to quantify (Fig.~\ref{fig:correlation_times}) the structural dynamics of the polarizable models againts NMR $R_{1}$ rates, measuring a window of dynamics around $\sim$0.1-1~ns, and effective correlation times $\tau_{\mathrm{eff}}$ which quantifies a larger range of dynamical processes up to $\sim$1000~ns.
-We include the same measurables quantified from non-polarizable CHARMM36 and SLipids force fields, which are the best MD force fields regarding the dynamics~\cite{Antila2021}, as a point of comparison.
-The Amoeba force field captures the dynamics with respect to both measurables well,  in par with the best non-polarizable models.
+\subsection{Evaluation of conformational dynamics of lipids}
+
+We proceed to evaluate (Fig.~\ref{fig:correlation_times}) the structural dynamics of the polarizable models againts NMR $R_{1}$ rates, here measured at a magnetic frequency that captures a window of dynamics around $\sim$0.1-1~ns, and effective correlation times $\tau_{\mathrm{eff}}$ which quantifies a larger range of dynamical processes up to $\sim$1000~ns.
+We include the same measurables calculated from non-polarizable CHARMM36 and SLipids force fields---the best MD force fields regarding the dynamics~\cite{Antila2021}---as a point of comparison.
+The AMOEBA force field captures the dynamics with respect to both measurables very well,  in par with the best non-polarizable models.
 In contrast, the dynamics of CHARMM-Drude appears to be faulty in the fast range ($R_{1}$ rates) and, furthermore, $\tau_{\mathrm{eff}}$ values reveal vastly too slow overall dynamics.
+The updated CHARMM-Drude2023 does not significantly improve the dynamics at the fast range investigated here.
+The effective correlation times of the newer Drude model are improved but still an order of magnitude too slow.
 This indicates that not only are the simulations with this polarizable force field computationally costly but one would have to simulate longer to obtain converged results.
+The non-polarizable counterpart of the Drude model, CHARMM36, exhibits much more realistic, faster effective correlation times.
+Interestingly, it has also been observed that CHARMM-Drude models have slower dynamic of water hydrogen bonding around amino acids compared to their non-polarizable counterpart~\cite{Ngo2019} which might align with an overall slower dynamics of the model in addition to enhanced water binding.
 Note that the experimental data here is from POPC.
-However, chemically identical the headgroup region the dynamics is assumed to be very similar in DOPC, due to the decoupling of the dynamics between the headgroup and the glycerol regions~\cite{Antila2022rot, Klauda08}.
+However, chemically identical the headgroup region the dynamics is assumed to be practically identical in DOPC, due to the decoupling of the dynamics between the headgroup and the glycerol regions~\cite{Antila2022rot, Klauda08}.
+
+\subsection{Cation binding to membranes in polarizable simulations}
 
@@
 These changes in order parameters are depicted in Fig.~\ref{fig:popc_order_parameter_change} whereas Fig.~\ref{fig:ion_density_profiles} gives the density profiles of ions with respect to the bilayer normal.
-The CHARMM36-Drude appears to have strong sodium binding, evidenced both by large change in the order parameters and the ion density profiles.
-The binding of monovalent sodium is, in fact, stronger than the divalent calcium.
+The CHARMM-Drude2017 exhibits strong sodium binding, evidenced both by large change in the order parameters and the ion density profiles.
+Based on the change in the order parameters, the sodium binding is even stronger than the divalent calcium binding.
 In strong contrast, experimentally one sees no sodium binding and some binding of calcium ions.
+Too strong sodium binding in the CHARMM-Drude model has been indicated before for peptides/amino acids~\cite{Ngo2019, Kav2022} and deep-eutectic solvents~\cite{shayestehpour2022ion} along with a too strong calcium interactions~\cite{Tan2022} in line with what we observe here for the lipids.
+
+The faulty sodium ion binding seems to have been somewhat remedied in the CHARMM-Drude2023 force field: the change of the order parameters is diminished to be more in line with experiments and only the some accumulation of sodium ions near the phosphate oxygen is seen in the ion distributions.
+Based on the change in the order parameters, the calcium binding also appears to be better represented in the updated model.
+The ion distributions reveal that binding of calcium has moved deeper into the bilayer and there is a stronger formation of ionic double layer by subsequent attraction of chloride into the bilayer.
+However, this the ion layering is most prominently present in the original CHARMM36 force field, for for sodium and calcium ions.
+Note, that both for sodium and calcium ion additionin the CHARMM-Drude2023force field reintroduces the forking of order parameters at the $\beta$ carbon in particular.
+These might also be a sign of poor convergence of the simulations owing to the slow conformational dynamics detected above, which would also not be fully reflected in the calculation of error estimates for the order parameters.
 
@@
 While this is in line with experiments regarding sodium ions, there should be some binding of calcium ions to the membrane.
-Again, the non-polarizable CHARMM36 appears to depict this calcium binding the best out of the models compared here.
+Again, the change in the order parameters indicate the non-polarizable CHARMM36 depicts sodium binding the best out of the models compared here, being slightly more realistic than its most recent polarizable counterpart, CHARMM-Drude2023.
+Regarding calcium binding, CHARMM36 captures the $\Delta^{\alpha}_{CH}$ about as well as the CHARMM-Drude2023, but the change in the $\beta$ carbon is better represented in the CHARMM-Drude2023 and AMOEBA models.
+This suggest that some accuracy might have been gained in modelling the divalent ion binding with polarizable models but the advantage is not clear.
+No force field, polarizable or non-polarizable, is able to model both monovalent and divalent ion binding flawlessly.
+This is of note, as polarizable force fields are often assumed superior and concequently a popupar choice in modelling, eg, ion channels~\cite{sun2017 jing2019polarizable, klesse2020induced}\todo{I think there is a mistake in the C36 data, firsly, the concentration should probably be 950mM instead of 1000mM and secondly the monovalent anion and cation curves look identical} ~\todo{batuhan: I fxed the anion and cation curves.
+Yes, C36 data is at 950mM but explicitly adding this as a new color could clutter the figure.
+Should I just mention this in the caption?}.
 
-%\begin{figure}[!hbt]
-%	\centering
-%	\includegraphics{popc_order_parameters_nosalt.png}
-%	\caption{The head group and glycerol backbone order parameters $S_{CH}$ for the POPC simulations without salt.}
-%	\label{fig:popc_order_no_salt}
-%        \todo{SAMULI: It would be good to put the experimental value also to second hydrogen even though it would be the same as for first.
-%          Now it looks like that those values would not be known.} \\
-%        \todo{SAMULI: y-axis should be more zoomed in.} \\
-%        \todo{SAMULI: Similar figure for POPE (now figure 4) could be here as panel B).}
-%\end{figure}
+\begin{figure}[t]
+    \centering
+    %\includegraphics{Figures/order_parameters.png}
+    \includegraphics[width=0.4\textwidth]{Figures/molecules.pdf}
+    \caption{Molecules and atom labelling used.}
+    \label{fig:molecules}
 
-%\begin{figure}[!hbt]
-%	\centering
-%	\includegraphics{pope_order_parameters.png}
-%	\caption{The head group and glycerol backbone order parameters $S_{CH}$ for the POPE simulations.}
-%	\label{fig:pope_order}
-%        \todo{SAMULI: Move to panel B) in Figure 1.} \\
-%        \todo{SAMULI: y-axis should be more zoomed in.} 
-%\end{figure}
+\end{figure}
 
@@
     %\includegraphics{Figures/order_parameters.png}
-    \includegraphics[width=\textwidth]{Figures/quality.pdf}
-    \caption{X-ray scattering form factors, and C-H bond order parameters for headgroup and glycerol backbone, and acyl chains (from top to bottom) compared between simulations and experiments using the NMRlipids databank.}
+    \includegraphics[width=0.8\textwidth]{Figures/quality.pdf}
+    \caption{X-ray scattering form factors, and C-H bond order parameters for headgroup and glycerol backbone, and acyl chains (from top to bottom) compared between simulations and experiments using the NMRlipids databank.
+The atom labelling is as depicted in Fig.~\ref{fig:molecules}.}
     \label{fig:order_parameters}
@@
         POPC & CHARMM-Drude & 0.51 & 0.22 & 0.47 & 1.16 & 62.5\\
+        POPC & CHARMM-Drude2023 & 0.50 & 0.49 & 0.53 & 0.82 & 64.5\\
         \hline
@@
         POPE & CHARMM-Drude & 0.06 & 0.53 & 0.27 & 0.80 & 56.6  \\
+        POPE & CHARMM-Drude2023 & 0.28 & 0.59 & 0.54 & 0.00 & 61.4  \\
         POPE & AMOEBA & 0.21 & 0.10 & 0.23 & 3.80 & 66.9\\
@@
     \caption{NMRlipids databank quality evaluations~\cite{Databank} and area per lipids compared between the best simulations found from the NMRlipids databank, simulations with CHARMM36 force field, and simulations with polarizable force fields.
-Experimental estimations for areas per lipid are POPC: 64.3$\pm1$\AA$^{2}$\cite{kucerka2011}, DOPC: 67.5$\pm1$\AA$^{2}$\cite{kucerka2008}, and POPE: 56.7$\pm$3\AA$^{2}$\cite{Rickeard2020}}
+Experimental estimations for areas per lipid are POPC: 64.3$\pm1$\AA$^{2}$\cite{kucerka2011}, DOPC: 67.5$\pm1$\AA$^{2}$\cite{kucerka2008}, and POPE: 56.7$\pm$3\AA$^{2}$\cite{Rickeard2020}.}
     \label{tab:my_label}
@@
 	\centering
-	\includegraphics{Figures/both_log.pdf}
+	\includegraphics{Figures/both_inkscp.pdf}
 	\caption{Effective correlation times (top) and R$_1$ rates (bottom) for the polarizable models and best performing non-polarizable force fields.
-Note, top panel axis is logarithmic to visualize the slow CHARMM-drude dynamics.
+Note, top panel axis is logarithmic to visualize the slow Drude dynamics.
 Experimental values are obtained from Ref.~\citenum{Antila2022rot}.
@@
 \subsection{Ion binding to membranes in simulations with polarizable force fields}
-
-%\begin{figure}[!hbt]
-%	\centering
-%	\includegraphics{popc_order_parameters_for_all.png}
-%	\caption{The head group and glycerol backbone order parameters $S_{CH}$ for the POPC simulations.
-The experimental values on the dashed lines mean that the different isomers 1/2 experimentally could not be resolved.}
-%	\label{fig:popc_order}
-%        \todo{SAMULI: I am not sure if we need this kind of figure. Systems without ions are in separate figures as well as dependence on ions.}
-%\end{figure}
 
@@
     \includegraphics{Figures/ion_density_profiles_with_chloride.eps}
-    \caption{Density profiles along the membrane normal for the Drude 2017 and 2023 POPC (first and second rows) and AMOEBA DOPC (third row).
-Solid and dashed lines are for the systems with the NaCl and CaCl$_{2}$, respectively.}
+    \caption{Density profiles along the membrane normal for the CHARMM36 (first row)~\cite{Catte2016}, Drude 2017 and 2023 POPC (second and third rows), AMOEBA DOPC (fourth row), and electronic continuum correction (ECC, last row)~\cite{Melcr:2018a}.
+Solid and dashed lines are for the systems with the NaCl and CaCl$_{2}$, respectively.
+Please note that only 350~mM and 450~mM CaCl$_{2}$ concentrations are shown for the Drude and AMOEBA force fields, respectively, while 1000~mM NaCl concentration is shown for all force fields.
+CHARMM36 NaCl simulations are at 950~mM.}
     \label{fig:ion_density_profiles}
@@
 
-%\subsection{Discussion}
+\subsection{Discussion}
+
+\textcolor{red}{For the ion binding discussion: drude calcium overbinds to proteins. this is known here https://pubs.acs.org/doi/full/10.1021/acsphyschemau.1c00039
+drude sodium overbinds to proteins, as shown here https://pubs.rsc.org/en/content/articlepdf/2022/ra/d2ra01478e and https://onlinelibrary.wiley.com/doi/abs/10.1002/adts.201800106.
+Venable also cautioned us against using any ion parameters with the drude force fields.
+This is important as 1- polarizable force fields are supposed to be better for ion parameters and 2- they are generally used to study the ion channels.}
 
@@
 After comparing the polarizable lipid models against high-resolution experimental data and the best performing non-polarizable force fields, we cannot confirm that adding polarizablity to the force fields has yet solved many of the issues it was hoped to address.
-Ion binding to the model membrane seems to be overestimated by the CHARMM36-Drude model, whereas the calcium binding is underestimated by the Amoeba force field.
-The electron density profiles of the bilayers appear to be better reproduced by the best non-polarizable force fields, within the confines of the approximations used to calculate them from the trajectory.
-Same applies for the dynamics and conformational sampling of the models, although the Amoeba force field produces results, spesifically in the headgroup region of PC lipids, that are in par or better than the best non-polarizable force fields with respect to these measurables.
-In summary, polarizable models appear still to be in their infancy, and added complexity of polarizability has not still realised its potential to the fullest; The non-polarizable force fields with longer and more intense refinement behind them still outpeform.
+Monovalent ion binding is best represented in the non-polarizable CHARMM36 and the polarizable AMOEBA.
+However, AMOEBA does not model well the divalent ion binding to bilayers.
+For calcium binding the best models are CHARMM36 and the polarizable CHARMM-Drude2023.
+ The electron density profiles of the bilayers are quite well captured by CHARMM-Drude2023 but rather poorly by AMOEBA, within the confines of the approximations used to calculate them from the trajectories.
+AMOEBA's poor performance regarding electron density curves might be related to the poor representation of lipid tail conformations in the force field, revealed by the order parameters.
+While CHARMM-Drude2023 captures tail and the headgroup conformations well, the non-polarizable models still outperform when comparing the order parameters.
+When investigating the conformational dynamics the AMOEBA force field produces results, specifically in the headgroup region of PC lipids, that are in par or better than the best non-polarizable force fields.
+However, the Drude models exhibit an extremely slow dynamics which leads to deficient sampling and compounding simulation time cost when using the model.
+In summary, the added complexity of polarizability has not still realised its potential to the fullest; The non-polarizable force fields with longer and more intense refinement behind them still outperform the polarizable models in some respects.
 Considering the complexity and cost of using these models, polarizability should be parameterized with more thorough consideration to the experimental benchmarks.
+When doing so, potential for improvement is significant, as exemplified by the refinement of the CHARMM-Drude2017 model to the most recent CHARMM-Drude2023 version.
 
@@
 
-In the Amoeba force field, a set of atomic multipoles were first obtained based on the QM electron densities using Stone's distributed multipole analysis~\cite{Stone1981}.
+In the AMOEBA force field, a set of atomic multipoles were first obtained based on the QM electron densities using Stone's distributed multipole analysis~\cite{Stone1981}.
 The atoms were then assigned polarizabilities according to the AMOEBA atom types and induced dipole contribution was self consistently subtracted from the permanent dipoles~\cite{shi2013proteinamoeba}.
@@
 
-Notes on Amoeba parametrization (this is NOT well explained in the paper)
+Notes on AMOEBA parametrization (this is NOT well explained in the paper)
 -bottom up approach
@@
 NAMD can only run the Drude force field whereas TINKER is only compatible with AMOEBA force field and does not support semi-isotropic pressure coupling, which decouples the pressure tensor in $xy$ and $z$-direction and is crucial for correctly evaluating the ion binding properties of the bilayers (\todo{This needs a reference, I remember Samuli mentioning this during a chat}).
-On the other hand, GROMACS, which has been the most popular MD engine within the bilayer simulation field, only has limited support for the Drude polarizable force field via an unofficial git-branch [https://github.com/gromacs/gromacs/tree/drude] but does not support AMOEBA at all.
+On the other hand, GROMACS, which has been the most popular MD engine within the bilayer simulation field, only has limited support for the Drude polarizable force field via an unofficial git-branch~\cite{drudegithub} but does not support AMOEBA at all.
 \begin{comment}
@@
 
-Current versions of the polarizable force fields require "good" starting structures.
-To achieve this, for every system, 200~ns long MD simulations with the non-polarizable CHARMM36 force field are run.
-The last frames from these simulations are fed into the CHARMM-GUI Drude Prepper to generate the input structures and OpenMM run files.
-Each bilayer contains 36 lipids in each leaflet.
-Simulations are run with the OpenMM MD engine.
+CHARMM-Drude 2017 force field parameters for OpenMM were extracted from CHARMM-GUI~\cite{jo2008charmm,lee2016charmm} \textit{Membrane Builder}~\cite{wu2014charmm,jo2009charmm,jo2007automated,lee2018charmm} and \textit{Drude Prepper}~\cite{kognole2022charmm}.
+CHARMM-Drude simulations were performed with OpenMM 7.5.0~\cite{eastman2017openmm}.
+Initial membrane structures have been equilibrated for 200~ns using the CHARMM36 force field~\cite{klauda2010update}.
+The last frames of these simulations were used to generate the starting structures for the polarizable force field simulations.
 
-The trajectories are publicly available on Zenodo.
+CHARMM-Drude 2023 force field parameters are currently not integrated into CHARMM-GUI.
+Therefore, the starting structures for these simulations with NaCl and CaCl$_{2}$ have been generated following the instructions in the original CHARMM-Drude 2023 paper~\cite{yu2023drude} using CHARMM~\cite{brooks2009charmm} and the last frame of 200~ns long CHARMM36 simulations.
+CHARMM-Drude 2023 simulations without any ions have been directly obtained from Zenodo.
+Please see Table~\ref{table:sim_details} for the full list of used simulations and their locations.
+All simulations were stored in the NMRlipids databank~\cite{Databank}.
+
+For both systems, a dual Langevin thermostat was employed to keep the Drude particles and the rest of the system at 1.0~K and 303~K, respectively.
+A Drude-hardwall of 0.02~nm was used to keep the Drude particles close to their parent atoms.
+Semi-isotropic Monte Carlo barostat was used to couple pressure to 1~bar independently in membrane plane and normal directions.
+%while the pressure fluctuations along the $xy$-plane were coupled.
+A constant surface tension of 0.0~dyne/cm was applied throughout the simulations.
+Bonds containing the hydrogens have been constrained.
+For CHARMM-Drude 2017, particle mesh Ewald was used to compute the Coulomb interactions.
+van der Waals interactions have been cut to 0 between 1.0~nm and 1.2~nm using a switching function.
+For the CHARMM-Drude 2023, Lennard-Jones particle-mesh Ewald (LJ-PME) method have been used to compute the long-range dispersions~\cite{wennberg2013lennard}.
+Simulation frames have been saved in every 10~ps.
 
@@
 
-AMOEBA force field parameters for OpenMM are obtained from GitHub [https://github.com/Inniag/openmm-scripts-amoeba,~\cite{klesse2020induced}].
-The initial bilayer structures, similar to our simulations with the Drude2017 force field, are initially obtained from CHARMM-GUI membrane builder and equilibriated with the CHARMM36 force field for 200~ns.
+AMOEBA force field parameters for OpenMM were obtained from GitHub~\cite{amoebagithub,klesse2020induced}.
+All AMOEBA simulations have been run on OpenMM 7.1.1~\cite{eastman2017openmm}.
+The same initial structures as the CHARMM-Drude simulations have been used to create the simulation setups.
+A multi-time step Langevin integrator was used to iterate the bonded and remaining interactions in every 0.5~fs and 2.0~fs, respectively.
+A non-bonded cutoff of 1.2~nm has been applied while semi-isotropic Monte Carlo barostat was used to couple pressure to 1~bar independently in membrane plane and normal directions.
+%keeping the pressure at 1~bar using a Monte Carlo barostat.
+Pressure fluctuations in the $xy$-plane has been coupled using a semi-isotropic pressure coupling and a constant surface tension at 0.0~dyn/cm has been applied throughout the simulations.
+Simulation frames have been saved in every 10~ps. Further simulations details can be found in the input files of the respective simulations (Table~\ref{table:sim_details})
+All simulations were stored in the NMRlipids databank~\cite{Databank}.
 
 \subsection{Analysis of simulations}
+
+Area per lipids, X-ray scattering form factors, and $S_{CH}$ order parameters automatically calculated by the NMRLipids Databank~\cite{Databank} were used.
+The $R_{1}$ rates and effective correlation times, along with the accompanying error estimates, are quantified from the trajectories as elaborated in Ref.~\citenum{Antila2021}.
+In-house python script available at XXX was used for this purpose.
+Source codes for all scripts and a detailed methodological explanation of the computed values are available in the referenced cited herein.
+
+\begin{comment}
 Order parameters are calculated using the script "OrderParameters.py" under\\ "https://github.com/batukav/NMRlipidsVIpolarizableFFs/tree/master/scripts". \\
@@
 This project is performed using the "NMRLipids Databank" format and all related files, except the trajectories, are available under\\ "https://github.com/batukav/NMRlipidsVIpolarizableFFs/"
+\end{comment}
 
-% Please add the following required packages to your document preamble:
 \newpage
@@
 \begin{tabular}{cccccccccc}
-	Lipid/ion                & force field  & Ion (M) & N$_{l}$ & N$_{w}$ & N$_{ion}$ & T (K) & t$_{sim}$ (ns) & t$_{anal}$ (ns) & files \\ \hline
-POPC                              & CHARMM-Drude & NA      & NA       & 6400       & 0         & 303    & 500              & 400         &          \cite{kav_batuhan_2021_4604630}    \\ \hline
-POPE                              & CHARMM-Drude & NA      & NA       & 6400       & 0         & 303    & 350              & 300         &          \cite{kav_batuhan_2021_4665773}    \\ \hline
-	\multirow{4}{*}{POPC:NaCl}        & CHARMM-Drude & 0.350   & 144      & 6400     & 41         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_4683386}   \\
-				  & CHARMM-Drude & 0.450   & 144      & 6400     & 51         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_4683398}   \\
-				  & CHARMM-Drude & 0.650   & 144      & 6400     & 77         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_4683405}   \\
-				  & CHARMM-Drude & 1.0     & 144      & 6400     & 115        & 303   & 500             & 400                  & \cite{kav_batuhan_2020_4683411}   \\ \hline
-	\multirow{3}{*}{POPC:CaCl$_{2}$} & CHARMM-Drude & 0.350   & 144      & 6400     & 41         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_4683393}   \\
-				  & CHARMM-Drude & 0.450   & 144      & 6400     & 52         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_4683391}   \\
-				  & CHARMM-Drude & 0.650   & 144      & 6400     & 76         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_4683394} \\
-				  & CHARMM-Drude & 1.0   & 144      & 6400     & 114         & 303   & 500             & 400                  & \cite{kav_batuhan_2021_4738966}
+	Lipid/ion                & force field  & Ion (M) & N$_{l}$ & N$_{w}$ & N$_{ion}$ & T (K) & t$_{sim}$ (ns) & t$_{analysis}$ (ns) & files \\ \hline
+\multirow{2}{*}{POPC}            & Drude2017 & 0      & 144       & 6400       & 0         & 303    & 500              & 400         &          \cite{kav_batuhan_2021_7607436}    \\ 
+                                 & Drude2023 & 0      & 72       & 2239       & 0         & 303      & 300              & 200         &          \cite{kav_batuhan_2023_7916287, richard_m_venable_2023_7871949}    \\ \hline
+\multirow{3}{*}{POPE}                              & Drude2017 & 0      & 144       & 6400       & 0         & 303    & 350              & 300         &          \cite{kav_batuhan_2021_7604627}    \\                                            & Drude2023 & 0       & 72        & 2304      &  0         & 303    & 300             & 200         &           \cite{richard_m_venable_2023_7872447,kav_batuhan_2023_7916494} \\
+                                                  & AMOEBA     & 0       & 72        & 2880      &  0        & 303     & 305.94          & 305.94      & \cite{kav_batuhan_2023_7622838} \\ \hline
+	\multirow{5}{*}{POPC:NaCl}        & Drude2017 & 0.350   & 144      & 6400     & 41         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_7586915}   \\
+				  & Drude2017 & 0.450   & 144      & 6400     & 51         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_7591753}   \\
+				  & Drude2017 & 0.650   & 144      & 6400     & 77         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_7596011}   \\
+				  & Drude2017 & 1.0     & 144      & 6400     & 115        & 303   & 500             & 400                  & \cite{kav_batuhan_2020_7600326}   \\ 
+                    & Drude2023 & 1.0      & 128       & 6400      & 41          & 303    & 223.97      &       220.14       & \cite{kav_batuhan_2023_8000095} \\     
+                    & Drude2023 & 1.0      & 128       & 6400      & 115          & 303    & 220.14      &       220.14       & \cite{kav_batuhan_2023_8000133} \\ \hline
+
+	\multirow{5}{*}{POPC:CaCl$_{2}$} & Drude2017 & 0.350   & 144      & 6400     & 41         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_7600827}   \\
+				  & Drude2017 & 0.450   & 144      & 6400     & 52         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_7605016}   \\
+				  & Drude2017 & 0.650   & 144      & 6400     & 76         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_7604040} \\
+				  & Drude2017 & 1.0   & 144      & 6400     & 114         & 303   & 500             & 400                  & \cite{kav_batuhan_2023_7658975}\\
+                    & Drude2023 & 0.350 & 128      & 6400     & 41          & 303   & 219.59          & 219.59               & \cite{kav_batuhan_2023_8000065}\\
+                    & Drude2023 & 0.790 & 128      & 6400     & 91          & 303   & 214.85          & 214.85               & \cite{kav_batuhan_2023_7992137} \\ \hline
+DOPC           & AMOEBA   & 0 & 72 & 2880 & 0 & 303 & 201.61 & 201.61 & \cite{kav_batuhan_2022_7604681} \\ \hline
+\multirow{2}{*}{DOPC:NaCl} & AMOEBA & 0.450 & 72 & 2880 & 17 & 303 & 218.41 & 218.41 & \cite{kav_batuhan_2022_7604711} \\
+                           & AMOEBA & 1.0 & 72 & 2880 &  35 & 303 & 201.61 & 201.61 & \cite{kav_batuhan_2023_7625844} \\ \hline
+\multirow{2}{*}{DOPC:CaCl$_{2}$} & AMOEBA  & 0.450 & 72 & 2880 & 16 & 303 & 218.41 & 218.41 & \cite{kav_batuhan_2022_7604842} \\
+                                 & AMOEBA & 1.0 & 72 & 2880  & 36 & 303 & 218.41 & 218.41 & \cite{kav_batuhan_2022_7604810}
 
@@
 Column N$_l$ gives the number of lipids, N$_w$ gives the number of water molecules, T(K) denotes the temperature in kelvins.
-Simulated time is listed in column t$_{sim}$ and time used for analysis in t$_{anal}$.
+Simulated time is listed in column t$_{sim}$ and time used for analysis in t$_{analysis}$.
 Column "files" gives the reference to open access simulation data.}
+\label{table:sim_details}
 \end{table}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\todo{batuhan: update the bib file to have 10ps fixed dt trajectories}
-\bibliography{nmrlipidsPOL.bib,references.bib}
+
+\bibliography{references.bib}
 

### 2023-09-11 manuscriptPOL.tex
revisions of the manuscript text (J.J.M.)

@@
 \email{batuhankav@gmail.com}
+\author{Jesper J. Madsen}
 \author{Markus S. Miettinen}
@@
 %   class\footnote{A footnote for the title}}
-\title{NMRLipids VI: Polarizable Force Fields}
-
+%\title{NMRLipids VI: Polarizable Force Fields}
+\title{Do explicitly polarizable force fields give better description of a biomembrane?}
+%\title{Comparison of polarizable lipid models to high-resolution experimental data reveals room for improvement}
+%\title{Benchmarking polarizable lipid force field}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\textcolor{red}{some ideas for the intro/conclusion: 1) amoeba and drude are mostly used for transmembrane protein simulations.
-https://pubs.acs.org/doi/abs/10.1021/acsnano.1c06443 https://www.sciencedirect.com/science/article/abs/pii/S0959440X2100110X https://pubs.rsc.org/en/content/articlehtml/2021/sc/d1sc01887f etc.
-Validation/comparison of these simulations with the experiments are done for the proteins, not the lipids.
-Now since we know that the lipid dynamics and conformations are not very correct with the polarizable force fields, can these affect the protein conformations?
-is there a way to possibly connect this to the previous nmrlipids jacs paper?\\
-}
+The dielectric environment in and around a lipid bilayer varies dramatically as you move from the water phase through the interfacial headgroup region to the membrane interior.
+The water phase highly polar, while the interfacial region contains dipolar headgroups, and the membrane interior is predominantly hydrophobic and nonpolar due to the aliphatic lipid tails.
 
-Conventional classical molecular dynamics (MD) simulations only consider electrostatics as interactions between static point charges, set by molecular dynamics models (force fields) as an input in the beginning of the simulation.
-The dynamic effects arising from polarizability are ignored or at most enter the simulations in an averaged fashion in the force field parametrization process.
-However, since decades work~\cite{Thole1981,ando2001stable,Grossfield2003,
-lamoreux2003,Antila2013 ,Lemkul2016, baker2015polarizable,jing2019polarizable } has been underway to make MD simulations explicitly polarizable in the hopes of reaching a more accurate description of the system.
+In conventional molecular dynamics (MD) simulations, electrostatic interactions are described by assigning static point charges to the atoms and molecules at the beginning of the simulation based on the force field definitions.
+Dynamic effects arising from molecular polarizability are traditionally not explicitly included and are only considered in an averaged fashion in the force field parametrization process when fitting to macroscopic observables.
+Significant efforts have been dedicated to introduce explicit polarizability into MD simulations, in the hopes of reaching a more accurate representation of the system~\cite{Thole1981,ando2001stable,Grossfield2003,
+lamoreux2003,Antila2013 ,Lemkul2016, baker2015polarizable,jing2019polarizable }.
+For bilayer membrane simulation, including polarizability in the lipids is expected to provide more accurate descriptions, especially in situations pertaining to membrane binding processes, translocation of charged biomolecules across the membrane, and the behavior of molecules residing within membranes, such as membrane proteins~\cite{baker2015polarizable,Lemkul2016, li2017drude,chu2018anionicpolarizable,Lynch21,Chen2021,Melcr:2018a, melcr2019improved,nencini22}.
+%Indeed, even implicit inclusion of polarizability appears to provide a better description of cation binding to membranes~\cite{Melcr:2018a, melcr2019improved}. 
+The main lipid force fields with explicit electronic polarization are CHARMM-Drude~\cite{li2017drude, yu2023drude}, AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable}, and CHARMM-Fluctuating Charge (FQ)~\cite{lucas2012charge}.
+The underlying treatment of polarizability differ among these models in the following way:  1) the classical Drude oscillator (CHARMM-Drude) models polarization by separating a core and a shell charge with a spring which stretches according to the envinronment giving the site a dipole moment~\cite{Lemkul2016},
+2) the induced point dipole/multipole (AMOEBA) uses polarizable point dipoles placed on chosen sites of the molecule~\cite{ponder2010current},
+and 3) the electronegativity equilization (FQ) employs atomic charges that are no longer constant but can change according to the electronegativities of the molecule atoms and the electric fields from the molecular enviroment~\cite{patel2004charmm}.
+All of these approaches result in an increasing computational cost, e.g, by introducing new types of interactions, more interactions sites, or by requiring a shorter time step in the simulations.
+As an alternative approach, electronic continuum correction (ECC) has been proposed as a computational efficient method to implicitly include polarizability by scaling the atom partial charges~\cite{Melcr:2018a,dijon20,Antila2021}.
 
-In the case of biomembranes (lipid bilayers), specifically, the molecular (dielectric) environment varies dramatically while crossing the membrane from the water phase across the dipolar/charged headgroup interface to the hydrophobic tail region.
-Including explicit polarizability in biomembrane modelling has the potential to provide more accurate description of membrane binding, membrane crossing, and behaviour and interactions of molecules residing within membranes, such as membrane protein folding.
-There certainly is room for improvement: efforts from the NMRlipids open science community have shown that the current non-polarizable lipid models cannot correctly reproduce the conformational ensemble~\cite{Botan2015,Antila2019}, conformational dynamics~\cite{Antila2021}, or ion biding~\cite{Catte2016} within MD simulations.
+%Hence, there certainly is room for improvement: efforts from the NMRlipids open science community have shown that the current non-polarizable lipid models cannot correctly reproduce the conformational ensemble~\cite{Botan2015,Antila2019}, conformational dynamics~\cite{Antila2021}, or ion biding~\cite{Catte2016} within MD simulations.
 
-Historically, there has been three main approaches to modelling polarizability: 1) the induced point dipole/multipole model where polarizable point dipoles are placed on chosen sites on a molecule, 2) the classical Drude oscillator model where polarization is modelled by separating a core and a shell charge with a spring which stretches according to the enviroment giving the site a dipole moment, and 3) the electronegativity equilization (fluctuating charge) model where atomic charges are no longer constant but can change according to the electronegativities of the molecule atoms and the electric fields from the molecular enviroment.
-In lipid modelling, the first two have surpassed the third in popularity.
-All of these approaches result in an increasing computational cost, e.g, by introducing new types of interactions, more interactions sites, or by requiring a shorter time step in the simulations.
+Our previous efforts in benchmarking state-of-the-art non-polarizable force fields have demonstrated that the quality of predictions for important membrane properties greatly vary between different force fields, particularly for lipid headgroup conformational ensembles and ion binding affinities~\cite{Botan2015,Catte2016,Antila2019,bacle21,Antila2021,Antila2022,Databank}.
+This emphasizes the critical importance of rigorous comparative quality evaluations of simulations performed using different force fields, which helps to estimate the reliability of simulation results and select the best models for specific applications.
+This is particularly important when employing polarizable force fields, as they offer the potential to capture a broader range of physical phenomena pertinent to enhancing our understanding of in particular polar membrane regions, at an increased computational cost.
+Unfortunately, systematic benchmark studies assessing the quality of polarizable force fields are,
+%is in great need. The few published studies reviewing the state of polarizable force fields for biomolecule simulations did not go beyond citing results from the existing studies
+to our knowledge, entirely missing from the field~\cite{inakollu2020polarisable,jing2019polarizable,baker2015polarizable}.
+Previous simulations using polarizable lipid force fields center around protein systems, with experimental validation (if performed at all) fovcusing on the properties of the proteins~\cite{jing2019polarizable, Lynch21}.
+Assessing the realism of the lipid models against experimental evidence from biophysical measurements is crucial since these membrane properties are foundational to more complicated applications, such as studies of transmembrane proteins.
+For example, it appears to be important that a wide variety of conformations is available for lipid headgroups in the model for realistic protein-lipid binding~\cite{bacle21}.
 
-%To avoid polarizability catastrophe where two closeby polarizable sites feed off from each other and the dipole moments grow infinitely
+Here we set out to assess the quality of available polarizable force fields, on par with previous evaluations of non-polarizable force fields, using the resources given by the NMRLipids open collaboration framework (https://databank.nmrlipids.fi/).
+Specifically, we evaluated the quality of POPC, DOPC and POPE lipid bilayer simulations with CHARMM-Drude~\cite{li2017drude,yu2023drude} and AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable} parameters against experimental NMR and small angle X-ray scattering (SAXS) data.
+These two force fields are increasingly utilized in biomolecular simulations and have lipid parameters available for the selected systems with corresponding experimental data available from the NMRlipids databank~\cite{Databank}.
+Following the quality metrics defined in the NMRlipids project, structural properties and cation binding were evaluated against NMR order parameters and SAXS form factors~\cite{Botan2015,ollila16,Catte2016,Databank}, while lipid conformational dynamics was evaluated in comparison to data from NMR spin relaxation rate experiments~\cite{ferreira15,Antila2021}.
+%Small angle scattering (SAXS) data is used to estimate the description of electron density profile of the membrane. 
+Furthermore, we compare the results to those obtained from the best performing non-polarizable simulations from the NMRlipids Databank\cite{Databank}.
 
-Considering the growing popularity of both implicitly and explicitly polarizable force fields and the increased computational cost that they require, a systematic benchmark study on polarizable force field quality is in great need.
-The few published studies reviewing the state of polarizable force fields for biomolecule simulations did not go beyond citing results from the existing studies~\cite{inakollu2020polarisable,jing2019polarizable,baker2015polarizable}.
-Recent publications show that \textit{implicit} inclusion of polarizability improves cation binding simulations when evaluated against NMR data~\cite{Melcr:2018a, melcr2019improved}, which suggest that including polarization explicitly could result in better description of ion binding as well.
-Currently, there are three main polarizable lipid force fields with explicit electronic polarization: CHARMM-Drude~\cite{li2017drude}, AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable}, and CHARMM-Fluctuating Charge (FQ)~\cite{lucas2012charge}.
-Here we evaluate the performance of two actively developed polarizable force fields with lipid parameters available, CHARMM-Drude~\cite{li2017drude} and AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable}, against NMR data.
-We use NMR order parameters to evaluate  structural properties and cation binding, whereas NMR relaxation rates and correlation times are utilized to benchmark the lipid conformational dynamics.
-Small angle scattering (SAXS) data is used to estimate the description of electron density profile of the membrane.
-Furthermore, we compare the results to the best performing non-polarizable simulations from the NMRlipids Databank\cite{Databank}.
+ %n lipid modelling, the first two have surpassed the third in popularity.
 
-This project has been performed within the NMRLipids collaboration (https://databank.nmrlipids.fi/)~\cite{Databank}.
-All simulation data, analysis scripts, and results are publicly available in their corresponding links.
-No plausible nor implausible request is required to download, use, and disseminate the data.
+%All simulation data, analysis scripts, and results are publicly available in their corresponding links. No plausible nor implausible request is required to download, use, and disseminate the data.
 
@@
 
-\subsection{Evaluation of lipid bilayer properties simulated with polarizable force fields}
+\subsection{Evaluation of lipid bilayer structural properties}
 
-We start by exploring the performance of polarizable force fields using the quality metrics developed in the NMRlipids Databank (https://databank.nmrlipids.fi/)~\cite{Databank}---a quality evaluated databank for lipid simulations.
-These metrics describe the ability of the models to reproduce the C-H NMR order parameters (quantifying the conformations) and the SAXS data (measuring the electron density variation across the membrane).
-For the former the molecule is divided into three parts, one containing the headgroup and the glycerol backbone carbons and one for each of the carbonyl tails, and the quality of each is evaluated separately.
-The quality metrics are summarized in Table~\ref{tab:my_label} and the performance of the force fields is visualized along with the underlying experimental measurables in Fig~\ref{fig:order_parameters}.
-
-One can clearly see that the non-polarizable force fields, CHARMM36 and OPLS3e, consistently outperform the polarizable force fields with respect to the NMR measurables for all parts of the molecule (higher $P$-values).
-For older CHARMM-Drude force field, this is likely due to the headgroup and glycerol parametrization having been done targetting absolute average values of the order parameters instead of taking account the potential order parameter forking ( measurably different order parameters for different C--H bods at one carbon) and order parameter sign~\cite{Antila2022}.
-This was then remedied in the newer version of the model, which also provides better description of the carbon tails.
-In fact, for POPE sn-2 tail the CHARMM-Drude2023 gives the best description of the models investigated here.
-The AMOEBA force field reproduces the most of the headgroup order parameters well for both DOPC and POPE but vastly overestimates one of the g$_3$ values, resulting in too large forking of oder parameters in this particular carbon.
-AMOEBA also clearly struggles to depict the tail conformations, in particular in the double bond region of DOPC (Fig.~\ref{fig:order_parameters}).
-In the tail regions, the CHARMM-Drude2023 appears to be the best out of the polarizable models.
-
-The higher $FFq$ values (Table~\ref{tab:my_label}) of the polarizable force fields for POPC indicate that they are not able to reproduce the SAXS data quite as well as the best non-polarizable force field, OPLS3e.
-That said, CHARMM-Drude2023, in particular, is clearly better than its non-polarizable counterpart CHARMM36 and the best out of all the models here in capturing the POPE SAXS data.
-AMOEBA, on the other hand, captures the SAXS profile of POPE very poorly (Table~\ref{tab:my_label} and Fig.~\ref{fig:order_parameters}D top) which is likely connected to the overall description of the lipid tails in the force fields.
-Note that owing the simulation size dependency of the SAXS data~\cite{Databank} one should focus on the extrema $q_{z}$[\AA$^{-1}$] when comparing the simulated and experimental data.
-In particular $FFq$ only measures the difference between the first minima of the simulated and experimental curves.
-Also note, that the method for estimating the charge density of bilayers from the simulations places electrons as point charges at the atom centers in the system and does not consider any redistribution of charge density due to the polarizability.
-Nevertheless, we expect this results into a good estimate of the charge density for the calculation of the SAXS curve.
-
-\subsection{Evaluation of conformational dynamics of lipids}
-
-We proceed to evaluate (Fig.~\ref{fig:correlation_times}) the structural dynamics of the polarizable models againts NMR $R_{1}$ rates, here measured at a magnetic frequency that captures a window of dynamics around $\sim$0.1-1~ns, and effective correlation times $\tau_{\mathrm{eff}}$ which quantifies a larger range of dynamical processes up to $\sim$1000~ns.
-We include the same measurables calculated from non-polarizable CHARMM36 and SLipids force fields---the best MD force fields regarding the dynamics~\cite{Antila2021}---as a point of comparison.
-The AMOEBA force field captures the dynamics with respect to both measurables very well,  in par with the best non-polarizable models.
-In contrast, the dynamics of CHARMM-Drude appears to be faulty in the fast range ($R_{1}$ rates) and, furthermore, $\tau_{\mathrm{eff}}$ values reveal vastly too slow overall dynamics.
-The updated CHARMM-Drude2023 does not significantly improve the dynamics at the fast range investigated here.
-The effective correlation times of the newer Drude model are improved but still an order of magnitude too slow.
-This indicates that not only are the simulations with this polarizable force field computationally costly but one would have to simulate longer to obtain converged results.
-The non-polarizable counterpart of the Drude model, CHARMM36, exhibits much more realistic, faster effective correlation times.
-Interestingly, it has also been observed that CHARMM-Drude models have slower dynamic of water hydrogen bonding around amino acids compared to their non-polarizable counterpart~\cite{Ngo2019} which might align with an overall slower dynamics of the model in addition to enhanced water binding.
-Note that the experimental data here is from POPC.
-However, chemically identical the headgroup region the dynamics is assumed to be practically identical in DOPC, due to the decoupling of the dynamics between the headgroup and the glycerol regions~\cite{Antila2022rot, Klauda08}.
-
-\subsection{Cation binding to membranes in polarizable simulations}
-
-Finally, we address the premise that adding polarizability to force fields improves the depiction of ion binding to biomembranes in simulations.
-For this, we use the concept of "lipid electrometer"~\cite{Catte2016} where the amount of ion binding to the membrane can be quantified by the change in $\alpha$ and $\beta$ order parameters upon increasing salt concentration.
-These changes in order parameters are depicted in Fig.~\ref{fig:popc_order_parameter_change} whereas Fig.~\ref{fig:ion_density_profiles} gives the density profiles of ions with respect to the bilayer normal.
-The CHARMM-Drude2017 exhibits strong sodium binding, evidenced both by large change in the order parameters and the ion density profiles.
-Based on the change in the order parameters, the sodium binding is even stronger than the divalent calcium binding.
-In strong contrast, experimentally one sees no sodium binding and some binding of calcium ions.
-Too strong sodium binding in the CHARMM-Drude model has been indicated before for peptides/amino acids~\cite{Ngo2019, Kav2022} and deep-eutectic solvents~\cite{shayestehpour2022ion} along with a too strong calcium interactions~\cite{Tan2022} in line with what we observe here for the lipids.
-
-The faulty sodium ion binding seems to have been somewhat remedied in the CHARMM-Drude2023 force field: the change of the order parameters is diminished to be more in line with experiments and only the some accumulation of sodium ions near the phosphate oxygen is seen in the ion distributions.
-Based on the change in the order parameters, the calcium binding also appears to be better represented in the updated model.
-The ion distributions reveal that binding of calcium has moved deeper into the bilayer and there is a stronger formation of ionic double layer by subsequent attraction of chloride into the bilayer.
-However, this the ion layering is most prominently present in the original CHARMM36 force field, for for sodium and calcium ions.
-Note, that both for sodium and calcium ion additionin the CHARMM-Drude2023force field reintroduces the forking of order parameters at the $\beta$ carbon in particular.
-These might also be a sign of poor convergence of the simulations owing to the slow conformational dynamics detected above, which would also not be fully reflected in the calculation of error estimates for the order parameters.
-
-As opposed to the CHARMM36-Drude model, the order parameters of the AMOEBA model are almost completely unaffected by the ion addition, both for NaCl and CaCl$_{2}$.
-The ion density profiles confirm that very little ions penetrate the membrane region where the lipid phosphates reside.
-While this is in line with experiments regarding sodium ions, there should be some binding of calcium ions to the membrane.
-Again, the change in the order parameters indicate the non-polarizable CHARMM36 depicts sodium binding the best out of the models compared here, being slightly more realistic than its most recent polarizable counterpart, CHARMM-Drude2023.
-Regarding calcium binding, CHARMM36 captures the $\Delta^{\alpha}_{CH}$ about as well as the CHARMM-Drude2023, but the change in the $\beta$ carbon is better represented in the CHARMM-Drude2023 and AMOEBA models.
-This suggest that some accuracy might have been gained in modelling the divalent ion binding with polarizable models but the advantage is not clear.
-No force field, polarizable or non-polarizable, is able to model both monovalent and divalent ion binding flawlessly.
-This is of note, as polarizable force fields are often assumed superior and concequently a popupar choice in modelling, eg, ion channels~\cite{sun2017 jing2019polarizable, klesse2020induced}\todo{I think there is a mistake in the C36 data, firsly, the concentration should probably be 950mM instead of 1000mM and secondly the monovalent anion and cation curves look identical} ~\todo{batuhan: I fxed the anion and cation curves.
-Yes, C36 data is at 950mM but explicitly adding this as a new color could clutter the figure.
-Should I just mention this in the caption?}.
+To evaluate the structural properties of lipid bilayers in simulations with polarizable force fields, we simulated POPC and POPE lipid bilayers with CHARMM-Drude2017~\cite{li2017drude} and CHARMM-Drude2023~\cite{yu2023drude} parameters, and POPE and DOPC bilayers with AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable} parameters.
+These systems were selected based on simultaneous availability of force field parameters, and experimental data~\cite{Databank}.
+After the data production, simulation trajectories were added into the NMRlipids Databank (https://github.com/NMRLipids/Databank/) and the NMRlipids quality metrics was used to evaluate each simulation against experiments~\cite{Databank}.
+This metric measures the quality of conformational ensemble of individual lipids in two aspects: first, it evaluates agreement with C-H bond order parameters (from NMR), and second, it examines consistency of membrane dimensions compared to small angle X-ray scattering (SAXS) data~\cite{Databank,ollila16}.
+%(measuring the electron density variation across the membrane). 
+The former is divided into three parts that separately describe the average quality of headgroup, glycerol backbone and acyl chain qualities as described in the NMRlipids databank~\cite{Databank}.
+While order parameters primarily reflect lipid conformations, acyl chain order parameters are also a good proxy for membrane packing since a smaller area per lipid tends to increase the magnitude of the order parameters~\cite{Databank}.
+%one for each of the carbonyl tails, and the quality of each is evaluated separately.
 
@@
 
-\begin{figure}[!hbt]
-    \centering
-    %\includegraphics{Figures/order_parameters.png}
-    \includegraphics[width=0.8\textwidth]{Figures/quality.pdf}
-    \caption{X-ray scattering form factors, and C-H bond order parameters for headgroup and glycerol backbone, and acyl chains (from top to bottom) compared between simulations and experiments using the NMRlipids databank.
-The atom labelling is as depicted in Fig.~\ref{fig:molecules}.}
-    \label{fig:order_parameters}
-
-\end{figure}
-
 \begin{table}[]
@@
         POPC & CHARMM36     & 0.70 & 0.54 & 0.69 & 1.16 & 65.0\\
-        POPC & CHARMM-Drude & 0.51 & 0.22 & 0.47 & 1.16 & 62.5\\
-        POPC & CHARMM-Drude2023 & 0.50 & 0.49 & 0.53 & 0.82 & 64.5\\
+        POPC & CHARMM-Drude & 0.52 & 0.29 & 0.53 & 1.06 & 62.5\\
+        POPC & CHARMM-Drude2023 & 0.63 & 0.60 & 0.57 & 0.96 & 64.5\\
         \hline
@@
     \end{tabular}
-    \caption{NMRlipids databank quality evaluations~\cite{Databank} and area per lipids compared between the best simulations found from the NMRlipids databank, simulations with CHARMM36 force field, and simulations with polarizable force fields.
+    \caption{NMRlipids databank quality metrics~\cite{Databank} and area per lipids (APL) compared between the best simulations found from the NMRlipids databank (OPLS3e for POPC and GROMOS-CKP for POPE), simulations with CHARMM36 force field, and simulations with polarizable force fields.
+Order  parameter quality measures, $P^{\mathrm{headgroup}}$, $P^{sn-1}$, and $P^{sn-2}$, present the average probability of order parameters within a given segment to agree with experiments (larger probabilities mean higher quality).
+Form factor quality measure, $FF_{q}$, presents the difference in essential features between simulated and experimental form factors (smaller values indicate higher quality).
 Experimental estimations for areas per lipid are POPC: 64.3$\pm1$\AA$^{2}$\cite{kucerka2011}, DOPC: 67.5$\pm1$\AA$^{2}$\cite{kucerka2008}, and POPE: 56.7$\pm$3\AA$^{2}$\cite{Rickeard2020}.}
@@
 
-%\begin{figure}[!hbt]
-%	\centering
-%	\includegraphics{dihedral_distributions_for_all.png}
-%	\caption{The distributions of the torsion angles for the head group atoms}
-%	\label{fig:dihedral}
-%        \todo{SAMULI: I am not sure if we need this figure at all.
-%          Or were you planning to discuss why order parameter response to ions is qualitatively incorrect in Drude PC?
-%          If yes, then maybe remove POPE from this figure.
-%        }
-%\end{figure}
+\begin{figure}[!p]
+    \centering
+    %\includegraphics{Figures/order_parameters.png}
+    \includegraphics[width=0.8\textwidth]{Figures/quality.pdf}
+    \caption{X-ray scattering form factors, and C-H bond order parameters for headgroup and glycerol backbone, and acyl chains (from top to bottom) compared between simulations and experiments using the NMRlipids databank.
+The atom labelling is as depicted in Fig.~\ref{fig:molecules}.
+Experimental data in this figure are originally reported in Refs.~\citenum{ferreira13,bacle21,Databank,NMRlipidsIII,kucerka2015}.
+For CHARMM-Drude2023 simulations we selected representative replicas among three available simulations, for example for differences between replicas for POPC see Fig.~\ref{fig:order_parameters_replicas}.}
+    \label{fig:order_parameters}
 
-\begin{figure}[!hbt]
-	\centering
-	\includegraphics{Figures/dihedral_distributions_without_ions.eps}
-	\caption{The distributions of the torsion angles for the head group atoms.
-For each torsion angle, the upper and lower rows contain the DOPC (AMOEBA) - POPC (Drude/Drude2023) and POPE lipids, respectively.}
-	\label{fig:dihedral_no_salt}
 \end{figure}
+
+Direct comparisons between simulated and experimental data are visualized in Fig~\ref{fig:order_parameters} and the resulting quality metrics are shown in Table~\ref{tab:my_label}.
+Similarly to the non-polarizable CHARMM36 simulations with Gromacs, CHARMM-Drude simulation predicts slightly too packed membrane with excessively negative order parameters with respect to experiments and simulations with the highest quality in the NMRlipids databank (OPLS3e)~\cite{Databank}.
+However,
+%consistently outperform the polarizable force fields with respect to the NMR measurables for all parts of the molecule (higher $P$-values). For older CHARMM-Drude force field, 
+the quality of headgroup conformations in the CHARMM-Drude simulations is worse than in its non-polarizable counterpart.
+This is likely because headgroup and glycerol backbone parameters were optimized to reproduce absolute average values of the experimentally determined order parameters, instead of taking into account the order parameter forking (meaning measurably different order parameters for different C--H bonds at a single carbon atom) and order parameter sign~\cite{Antila2022}.
+A better description of the headgroup and glycerol backbone is provided by CHARMM36-Drude2023, yet its quality is still below that offered by the non-polarizable CHARMM36 parameters (see also discussion about conformational dynamics below).
+Also the quality of membrane packing and acyl chain order are improved in CHARMM36-Drude2023 compared to the earlier version but again, it did not perform at a quality comparable to that of the best available simulations.
+%also provides better description of the carbon tails. 
+%In fact, for POPE sn-2 tail the CHARMM-Drude2023 gives the best description of the models investigated here.
+
+ AMOEBA simulations capture the headgroup and glycerol backbone order parameters reasonably well with the exception of g$_1$,
+%ll for both DOPC and POPE but vastly overestimates one of the g$_3$ values, resulting in 
+where forking of order parameters is unacceptably large.
+However, experimentally observed low order parameters at double bond region in both DOPC acyl chains and {\it sn}-2 chain of POPE are not captured.
+These low order parameters signal an important mechanism through which acyl chain double bonds affect membrane properties~\cite{ollila07} and are well reproduced in all state-of-the art MD simulation parameters~\cite{ollila16}.
+Furthermore, AMOEBA parameters substantially overestimate the area per lipid in POPE simulations, leading to too disordered acyl chains.
+Similar issues are evident also in membrane data presented in the recent publication for the AMOEBA cholesterol model~\cite{Li23chol}.
+%also clearly struggles to depict the tail conformations, in particular in the double bond region of DOPC (Fig.~\ref{fig:order_parameters}). 
+%parameters do not capture the  
+%In the tail regions, the CHARMM-Drude2023 appears to be the best out of the polarizable models.
+
+The unsatisfactory description of the lipid tail region and the area per lipid is further reflected in the inability of AMOEBA to capture the SAXS curves.
+Thus, we can conclude that our simulations with the AMOEBA parameters did not reproduce essential
+membrane properties.
+The higher $FFq$ values (Table~\ref{tab:my_label}) of the polarizable force fields in general indicate that they are not able confer consistency with the SAXS data as well as the best non-polarizable force fields (OPLS3e).
+That said, CHARMM-Drude2023 is on par with or better than its non-polarizable counterpart CHARMM36 and the best out of all the models here in capturing the POPE SAXS data, benefiting from the improvements made in the description of acyl chains.
+
+\subsection{Evaluation of conformational dynamics of lipids}
+C-H bond order parameters are sensitive to the conformational ensemble but the correspondence is not unique.
+Instead, they essentially describe the average of the distributions and do not contain information of the dynamics of the confomational sampling.
+Therefore, a simulation that reproduces the order parameters has an ensemble that is only potentially correct, and even the correct ensemble may not have the experimentally observed dynamics.
+To further elucidate this aspect in the case of polarizable force fields, we compared NMR $^{13}$C spin relaxation $R_{1}$ rates and C-H bond effective correlation times $\tau_{\mathrm{eff}}$ with experiments~\cite{ferreira15} and best non-polarizable simulations from our previous study~\cite{Antila2021} for simulations with PC headgroups and the glycerol backbone in Fig.~\ref{fig:correlation_times}.
+This focus was chosen based on the availability of both experimental data and polarizable simulations.
+$R_{1}$ rates measured with typical magnetic field strengths are sensitive to rotational dynamics of C-H bonds with the timescales around $\sim$0.1-1~ns, while effective correlation times respond to a larger range of dynamical processes up to $\sim$1000~ns~\cite{ferreira15}.
+The effective correlation time can be interpreted as an average measure for how fast molecular conformations sample the phase space that leads to the average order parameters.
+
+The effective correlation times in CHARMM-Drude2017 and CHARMM-Drude2023 simulations are approximately two  and one orders of magnitude slower, respectively, than the values suggested by experiments and best available simulations (Fig.~\ref{fig:correlation_times}).
+This indicates that not only are the simulations with this polarizable force field computationally costlier for equivalent lengths of trajectory, but one would also have to simulate longer to obtain converged results.
+The non-polarizable counterpart of the Drude models, CHARMM36, exhibits much more realistic, faster effective correlation times.
+Also the dynamics in the fast range ($R_{1}$ rates), are on average slightly more realistic in the non-polarizable CHARMM36 compared to both of the polarizable counterparts.
+Interestingly, CHARMM-Drude models have been reported to have slower water hydrogen bonding dynamics around amino acids compared to their non-polarizable counterpart,~\cite{Ngo2019} which might align with an overall slower dynamics of the model in addition to enhanced water binding.
+
+In contrast to the Drude-based models investigated here, the $R_1$ spin relaxation times and effective correlation times from DOPC simulations with AMOEBA force field are close to experimental data from POPC and on par with the best non-polarizable models (Slipids and CHARMM36).
+The small difference in acyl chain composition (DOPC vs.
+POPC) is not expected to strongly affect headgroup dynamics due to effective decoupling between hydrophilic and hydrophobic membrane regions~\cite{Antila2022rot, Klauda08}.
+%appears to be faulty in the fast range ($R_{1}$ rates) and, furthermore, $\tau_{\mathrm{eff}}$ values reveal vastly too slow overall dynamics.
+The updated CHARMM-Drude2023 does not significantly improve the dynamics at the fast range investigated here.
+The effective correlation times of the newer Drude model are improved but still an order of magnitude too slow.
+This indicates that not only are the simulations with this polarizable force field computationally costly but one would have to simulate longer to obtain converged results.
+The non-polarizable counterpart of the Drude model, CHARMM36, exhibits much more realistic, faster effective correlation times.
+%Interestingly, it has also been observed that CHARMM-Drude models have slower dynamic of water hydrogen bonding around amino acids compared to their non-polarizable counterpart~\cite{Ngo2019} which might align with an overall slower dynamics of the model in addition to enhanced water binding.
+%Note that the experimental data here is from POPC.
+However, chemically identical the headgroup region the dynamics is assumed to be practically identical in DOPC, due to the decoupling of the dynamics between the headgroup and the glycerol regions~\cite{Antila2022rot, Klauda08}.
+
+%We include the same measurables calculated from non-polarizable CHARMM36 and SLipids force fields---the best MD force fields regarding the dynamics~\cite{Antila2021}---as a point of comparison.
 
@@
 	\label{fig:correlation_times}
-
 \end{figure}
 
-\clearpage
+\subsection{Cation binding to membranes in polarizable simulations}
 
-\subsection{Ion binding to membranes in simulations with polarizable force fields}
+Given the abundance of cations in biological systems, accurately capturing their interactions with membranes has the uttermost importance in MD simulations.
+A wealth of  experimental evidence shows that monovalent ions, such as sodium (except for lithium), exhibit very weak binding affinity of PC lipid bilayers, while multivalent ions like calcium bind more strongly~\cite{Catte2016}.
+However, simulations with non-polarizable force fields without any additional corrections systematically overestimate cation binding to lipid bilayers~\cite{Catte2016}.
+Implicit inclusion of electronic continuum correction (ECC) to both ion and lipid parameters can substantially improve the situation~\cite{Melcr:2018a,melcr2019improved,bacle21}, suggesting that electronic polarizability plays an important role in ion binding to membranes, which might further lead one to expect that that simulations with polarizable force fields will perform better in this aspect and more accurately describe ion binding to membranes.
+%Finally, we address the premise that adding polarizability to force fields improves the depiction of ion binding to biomembranes in simulations. 
+To test this notion, we evaluated sodium and calcium cation binding to membranes against the experimental NMR data using the "lipid electrometer," which involves quantifing the amount of ion binding to the membrane by monitoring the change in lipid headgroup order parameters in response to increasing salt concentration~\cite{Catte2016}.
+Changes in these order parameters induced by increasing NaCl and CaCl$_2$ are shown in Fig.~\ref{fig:popc_order_parameter_change} for simulations and experiments, whereas Fig.~\ref{fig:ion_density_profiles} shows the density profiles of ions with respect to the bilayer normal.
+Results from AMBER-based ECC-lipid model are shown for the reference that gives a good agreement with experiments for cation binding.
 
@@
 
-\clearpage
+The CHARMM-Drude2017 predicts similar calcium ion density profile to ECC-lipids, in good agreement with experiments.
+However, the sodium binding is equally strong in these simulations in contrast with the experimental evidence~\cite{Catte2016}.
+%, evidenced both by large change in the order parameters and the ion density profiles.
+Based on the change in the order parameters, the sodium binding is even stronger than the divalent calcium binding.
+In strong contrast, experimentally one sees no sodium binding and some binding of calcium ions.
+Excessive sodium binding in the CHARMM-Drude model has been noticed before for systems containing peptides or amino acids~\cite{Ngo2019, Kav2022} and deep-eutectic solvents~\cite{shayestehpour2022ion}.
+Excessive calcium interactions, in line with what we observe here for the lipids, have also been reported~\cite{Tan2022}.
+Furthermore, the response of headgroup order parameters to the bound ions is not in qualitative agreement with experiments for CHARMM-Drude2017 simulations according to Fig.~\ref{fig:popc_order_parameter_change}.
+Particularly, increasing $\alpha$ carbon order parameters and forking upon ion binding are not observed in experiments.
+This is in contrast with the results from non-polarizable simulations, where experimentally observed decrease of order parameters to more negative values upon ion binding were observed in all simulations, even though the binding affinity was often inaccurately predicted~\cite{Catte2016}.
+This qualitative discrepancy in CHARMM-Drude2017 simulations most likely results from incorrect lipid headgroup conformational ensemble (see previous sections), which leads to inaccuracies in the structural response of the ensemble to ion binding.
+
+In simulations with CHARMM-Drude2023 parameters, sodium ion binding is in line with ECC-lipid simulations when comparing density profiles,
+%force field: the change of the order parameters is diminished to be more in line with experiments and only the some accumulation of sodium ions near the phosphate oxygen is seen in the ion distributions.
+%However, this the ion layering is most prominently present in the original CHARMM36 force field, for for sodium and calcium ions. 
+but addition of sodium or calcium ions seems to induce forking of order parameters in contrast to experiments. 
+%at the $\beta$ carbon in particular. 
+This might indicate inaccurate structural response to ion binding, but poor convergence of the simulations owing to the slow conformational dynamics detected above cannot be ruled out in this case or in the case of the older 2017 version.
+%, which would also not be fully reflected in the calculation of error estimates for the order parameters. 
+Based on density profiles, calcium binding in CHARMM-Drude2023 simulations is in line with the ECC-lipid model, although binding affinity is slightly larger, 
+%also appears to be better represented in the updated model. The 
+ions penetrate slightly
+%reveal that binding of calcium has moved 
+deeper into the bilayer, and there is a stronger formation of ionic double layer by subsequent attraction of chloride into the bilayer. 
+%As opposed to the CHARMM36-Drude model, the order parameters of the AMOEBA model are almost completely unaffected by the ion addition, both for NaCl and CaCl$_{2}$.
+The ion density profiles confirm that very little ions penetrate the membrane region where the lipid phosphates reside.
+While this is in line with experiments regarding sodium ions, there should be some binding of calcium ions to the membrane.
+In AMOEBA simulations, sodium binds weakly and does not affect order parameters, consistent with experiments and ECC-lipid simulations.
+However, order parameters do not change upon calcium binding in accordance with experiments and other simulations.
+This can be explained by the binding position of calcium, which is outside phosphate density peak in AMOEBA simulations, whereas calcium penetrates to phosphate region or deeper in other simulations and thereby reorients the headgroup dipole, giving rise to changes in order parameters~\cite{Catte2016}.
+
+%Again, the change in the order parameters indicate the non-polarizable CHARMM36 depicts sodium binding the best out of the models compared here, being slightly more realistic than its most recent polarizable counterpart, CHARMM-Drude2023.
+Regarding calcium binding, CHARMM36 captures the $\Delta^{\alpha}_{CH}$ about as well as the CHARMM-Drude2023, but the change in the $\beta$ carbon is better represented in the CHARMM-Drude2023 and AMOEBA models.
+This suggest that some accuracy might have been gained in modelling the divalent ion binding with polarizable models but the advantage is not clear.
+No force field, polarizable or non-polarizable, is able to model both monovalent and divalent ion binding flawlessly.
+This is of note, as polarizable force fields are often assumed superior and concequently a popupar choice in modelling, eg, ion channels~\cite{sun2017 jing2019polarizable, klesse2020induced}
+\todo{I think there is a mistake in the C36 data, firsly, the concentration should probably be 950mM instead of 1000mM and secondly the monovalent anion and cation curves look identical} ~\todo{batuhan: I fxed the anion and cation curves.
+-> @batuhan will plot the c36 data in a different color for the 950mm case.s}.
+
+%\begin{figure}[!hbt]
+%	\centering
+%	\includegraphics{dihedral_distributions_for_all.png}
+%	\caption{The distributions of the torsion angles for the head group atoms}
+%	\label{fig:dihedral}
+%        \todo{SAMULI: I am not sure if we need this figure at all.
+%          Or were you planning to discuss why order parameter response to ions is qualitatively incorrect in Drude PC?
+%          If yes, then maybe remove POPE from this figure.
+%        }
+%\end{figure}
+
+%\clearpage
+
+%\clearpage
+
+In conclusion, incorporating explicit polarizability as done in CHARMM-Drude or AMOEBA force fields do not necessarily lead to improved description of cation binding to phospolipid membranes.
+However, it is difficult to isolate the explicit influence of polarizability per se on ion binding based on the results presented here.
+What is clear, however, is that the headgroup conformational ensemble, which presumably plays a role in ion binding, is poorly described by the CHARMM-Drude parameters.
+In addition, the AMOEBA parameters struggle to capture the basic membrane properties that are reproduced by all current state-of-the-art MD simulation models.
 
@@
 \subsection{Conclusions}
-After comparing the polarizable lipid models against high-resolution experimental data and the best performing non-polarizable force fields, we cannot confirm that adding polarizablity to the force fields has yet solved many of the issues it was hoped to address.
-Monovalent ion binding is best represented in the non-polarizable CHARMM36 and the polarizable AMOEBA.
-However, AMOEBA does not model well the divalent ion binding to bilayers.
-For calcium binding the best models are CHARMM36 and the polarizable CHARMM-Drude2023.
- The electron density profiles of the bilayers are quite well captured by CHARMM-Drude2023 but rather poorly by AMOEBA, within the confines of the approximations used to calculate them from the trajectories.
-AMOEBA's poor performance regarding electron density curves might be related to the poor representation of lipid tail conformations in the force field, revealed by the order parameters.
-While CHARMM-Drude2023 captures tail and the headgroup conformations well, the non-polarizable models still outperform when comparing the order parameters.
-When investigating the conformational dynamics the AMOEBA force field produces results, specifically in the headgroup region of PC lipids, that are in par or better than the best non-polarizable force fields.
-However, the Drude models exhibit an extremely slow dynamics which leads to deficient sampling and compounding simulation time cost when using the model.
-In summary, the added complexity of polarizability has not still realised its potential to the fullest; The non-polarizable force fields with longer and more intense refinement behind them still outperform the polarizable models in some respects.
-Considering the complexity and cost of using these models, polarizability should be parameterized with more thorough consideration to the experimental benchmarks.
-When doing so, potential for improvement is significant, as exemplified by the refinement of the CHARMM-Drude2017 model to the most recent CHARMM-Drude2023 version.
+We have rigorously compared simulation output of model lipid bilayers using polarizable lipid force fields with high-resolution experimental data as well as results from best-performing non-polarizable force fields.
+From this exercise, it is evident that current polarizable force fields for lipids fall short in important aspects.
+In most cases, the best non-polarizable models exhibits superior performance and accuracy across all the categories investigated in this study.
+
+Evaluation order parameters and scattering form factors reveals that the AMOEBA force field has difficulties in producing membranes with reasonable physical properties.
+In constrast, the CHARMM-Drude2023 model has improved performance in this regard, notably surpassing its predecessor, CHARMM-Drude2017.
+Nevertheless, top-performing non-polarizable models generally reproduce experimental conformations (order parameters) better, although not perfectly so.
+
+In regard to conformational dynamics, the AMOEBA force field produces results, specifically in the headgroup region of PC lipids, that are either on par with or better than the best non-polarizable force fields.
+However, the CHARMM-Drude models exhibit an extremely slow dynamics, which can lead to deficient sampling and a substantial increase in simulation simulation time when using this model.
+
+In general, the ECC model exhibits thet most accurate representation of ion binding, even though it lacks an explicit description of polarization.
+The non-polarizable CHARMM36, when used with the NBFix ion model designed to provide more accurate binding, does XXX.
+The polarizable AMOEBA and Drude-2023 models show similar ion binding levels to the ECC model, but they exhibit variation in binding depth and the consequent structural response of the lipids do not align with experiments.
+The incorrect response to ion binding likely connects to the the other inaccuracies described above, including the misrepresentation of overall membrane structure and diversity in the conformational ensemble, as well as slow dynamics that lead to poor convergence of simulations.
+
+ In summary, the full potential and promise of explicitly polarizable force fields have still not been fully realized.
+Non-polarizable force fields benefit from a longer history of intense scrutiny and out-perform their polarizable counterparts in many respects, as we here show for the case of certain phospholipid bilayers.
+Considering the complexity and cost of using these models, it is crucial to parameterize polarizability with more thorough consideration to the experimental benchmarks.
+Doing so holds great potential for improvement, as demonstrated by the refinement of the CHARMM-Drude2017 model to the most recent CHARMM-Drude2023 version.
 
 \section{Methods}
-\subsection{Practical pitfalls of polarizable biomembrane simulations}
+%\subsection{Practical pitfalls of polarizable biomembrane simulations}
+
+\begin{comment}
 
@@
 After settling all the nonbonded parameters, dihedral potentials were parameterised based on the QM conformational energy profiles.
-
+\end{comment}
 \begin{comment}
@@
 \end{comment}
-\subsubsection{Using a polarizable force field}
+\subsection{Using a polarizable force field for membrane simulations}
 
-The increased level of detail portrayed by explicit treatment of electronic polarizability comes with certain costs.
-The first cost is the increased computational complexity reduces the simulation speed.
-For the Drude force field, this slowdown occurs both because the addition of the Drude particles increases the number of interaction pairs and the employed extended dual-Langevin thermostat requires 1~fs integration time step.
-Current AMOEBA force field can use a multi time-step integration algorithm where the non-electronic interactions are iterated with a 2~fs time step whereas more computationally unstable polarization terms are iterated with a smaller time step.
-Yet, this usage of a multi-time step integration scheme does not overcome the slowdown resulting from the computationally demanding polarization calculations.
-As a result, one experiences almost a ten-fold decrease in the simulation speed while using the polarizable force fields.
-The second cost is due to the limited software support for both running and analyzing the simulations.
-Currently only OpenMM supports both AMOEBA and Drude force fields.
-NAMD can only run the Drude force field whereas TINKER is only compatible with AMOEBA force field and does not support semi-isotropic pressure coupling, which decouples the pressure tensor in $xy$ and $z$-direction and is crucial for correctly evaluating the ion binding properties of the bilayers (\todo{This needs a reference, I remember Samuli mentioning this during a chat}).
-On the other hand, GROMACS, which has been the most popular MD engine within the bilayer simulation field, only has limited support for the Drude polarizable force field via an unofficial git-branch~\cite{drudegithub} but does not support AMOEBA at all.
-\begin{comment}
-This is particularly important, as GROMACS also comes with many built-in functions for analyzing the trajectories and many of these functions require a GROMACS tpr file and subsequently cannot be utilized for analyzing the trajectories generated by other MD engines.
-This hinders the researchers from using well-tested analysis tools creating an overhead during the trajectory analysis while making results more prone to error.
-Therefore, open-source Python libraries for trajectory analysis like MDAnalysis, MDtraj, and LiPhyphilic (based on MDAnalysis) that can work on multiple trajectory and topology formats are gaining more attention.
-\end{comment}
-Thanks to its recent update, CHARMM-GUI now can generate the topology and input parameters for the Drude force field, which greatly simplifies the employment of this model.
+Considering the increased interest in polarizable force fields in membrane simulations, we mention here some hurdles in their practical usage.
+Firstly, the increased level of detail obtained by explicit treatment of electronic polarizability always comes with increasing computational cost.
+For the Drude-based models, this slowdown occurs both because the addition of the Drude particles increases the number of interaction pairs and the employed extended dual-Langevin thermostat requires a shorter 1~fs integration time step, compared to the 2~fs typically used for non-polarizable membrane simulations.
+The AMOEBA force field can use a multi time-step integration algorithm where the non-electronic interactions are iterated with a 2~fs time step whereas more computationally unstable polarization terms are iterated with a smaller time step.
+However, the multi-time step scheme  only partly mitigates the computational cost: one typically experiences almost a ten-fold decrease in the simulation speed while using the polarizable force fields.
+
+Secondly, there are compability issues with the most used MD engines and the polarizable models and that support might be version dependent.
+Currently, only OpenMM supports both AMOEBA and Drude force fields.
+NAMD can only run the Drude force field whereas TINKER is only compatible with AMOEBA and does not support semi-isotropic pressure coupling, which decouples the pressure tensor in $xy$ and $z$-direction and is crucial for correctly modelling membrane systems~\cite{xxx}.
+GROMACS---the most popular MD engine within the bilayer simulation field---only has limited support for the Drude polarizable force field via an unofficial Git-branch~\cite{drudegithub} and no support for AMOEBA.
+
+In general, easy access to MD engines, as well as adaptability of common analysis tools and libraries for polarizable force fields, will be crucial in facilitating more wide-spread utilization of these models.
+Fortunately, CHARMM-GUI now can generate the topology and input parameters for the Drude force field, which greatly simplifies the employment of this model in particular~\cite{kognole2022charmm}.
 
@@
 
-CHARMM-Drude 2017 force field parameters for OpenMM were extracted from CHARMM-GUI~\cite{jo2008charmm,lee2016charmm} \textit{Membrane Builder}~\cite{wu2014charmm,jo2009charmm,jo2007automated,lee2018charmm} and \textit{Drude Prepper}~\cite{kognole2022charmm}.
-CHARMM-Drude simulations were performed with OpenMM 7.5.0~\cite{eastman2017openmm}.
-Initial membrane structures have been equilibrated for 200~ns using the CHARMM36 force field~\cite{klauda2010update}.
-The last frames of these simulations were used to generate the starting structures for the polarizable force field simulations.
+Table~\ref{table:sim_details} for the full list of simulations conducted for this study and links to the openly available trajectories.
+Data not mentioned here was obtained by analysis pre-existing trajectories in the NMRlipids databank.
+
+CHARMM-Drude2017 simulations were performed with OpenMM 7.5.0~\cite{eastman2017openmm} using parameters extracted with \textit{Membrane Builder}~\cite{wu2014charmm,jo2009charmm,jo2007automated,lee2018charmm} and \textit{Drude Prepper}~\cite{kognole2022charmm} from CHARMM-GUI~\cite{jo2008charmm,lee2016charmm}.
+Before starting the simulations, membrane structures were equilibrated for 200~ns using the non-polarizable CHARMM36 force field~\cite{klauda2010update}, and the last frames of these simulations were used to generate the starting structures for the polarizable force field simulations.
 
 CHARMM-Drude 2023 force field parameters are currently not integrated into CHARMM-GUI.
-Therefore, the starting structures for these simulations with NaCl and CaCl$_{2}$ have been generated following the instructions in the original CHARMM-Drude 2023 paper~\cite{yu2023drude} using CHARMM~\cite{brooks2009charmm} and the last frame of 200~ns long CHARMM36 simulations.
-CHARMM-Drude 2023 simulations without any ions have been directly obtained from Zenodo.
-Please see Table~\ref{table:sim_details} for the full list of used simulations and their locations.
-All simulations were stored in the NMRlipids databank~\cite{Databank}.
+Therefore, the simulation setups with NaCl and CaCl$_{2}$ have been generated following the instructions in the original CHARMM-Drude 2023 paper~\cite{yu2023drude} using CHARMM~\cite{brooks2009charmm} and the last frame of 200~ns long CHARMM36 simulations.
+CHARMM-Drude 2023 simulations without any ions have been directly obtained from Zenodo~\cite{richard_m_venable_2023_7872447, richard_m_venable_2023_7871949}.
 
-For both systems, a dual Langevin thermostat was employed to keep the Drude particles and the rest of the system at 1.0~K and 303~K, respectively.
-A Drude-hardwall of 0.02~nm was used to keep the Drude particles close to their parent atoms.
+A dual Langevin thermostat was employed to keep the Drude particles and the rest of the system at 1.0~K and 303~K, respectively.
+A Drude-hardwall of 0.02~nm was utilized to keep the Drude particles close to their parent atoms.
 Semi-isotropic Monte Carlo barostat was used to couple pressure to 1~bar independently in membrane plane and normal directions.
@@
 A constant surface tension of 0.0~dyne/cm was applied throughout the simulations.
-Bonds containing the hydrogens have been constrained.
+Bonds containing the hydrogens were constrained.
 For CHARMM-Drude 2017, particle mesh Ewald was used to compute the Coulomb interactions.
 van der Waals interactions have been cut to 0 between 1.0~nm and 1.2~nm using a switching function.
-For the CHARMM-Drude 2023, Lennard-Jones particle-mesh Ewald (LJ-PME) method have been used to compute the long-range dispersions~\cite{wennberg2013lennard}.
+For the CHARMM-Drude 2023, Lennard-Jones particle-mesh Ewald (LJ-PME) method was used to compute the long-range dispersions~\cite{wennberg2013lennard}.
 Simulation frames have been saved in every 10~ps.
@@
 The same initial structures as the CHARMM-Drude simulations have been used to create the simulation setups.
-A multi-time step Langevin integrator was used to iterate the bonded and remaining interactions in every 0.5~fs and 2.0~fs, respectively.
+A multi-time step Langevin integrator was used to iterate the bonded and non-bonded interactions in every 0.5~fs and 2.0~fs, respectively.
 A non-bonded cutoff of 1.2~nm has been applied while semi-isotropic Monte Carlo barostat was used to couple pressure to 1~bar independently in membrane plane and normal directions.
@@
 Simulation frames have been saved in every 10~ps. Further simulations details can be found in the input files of the respective simulations (Table~\ref{table:sim_details})
-All simulations were stored in the NMRlipids databank~\cite{Databank}.
 
 \subsection{Analysis of simulations}
+ All simulations were first uploaded to the NMRlipids databank~\cite{Databank}. 
+Area per lipids, X-ray scattering form factors, and order parameters ($S_{CH}$) are automatically calculated by the NMRLipids Databank~\cite{Databank} and were extracted from therein.
+Quality evaluation metrics were quantified as detailed in the NMRlipids databank~\cite{Databank} with the exception that POPC simulations at 303\,K were coupled with the experimental data measured at 300\,K, while in the NMRlipids databank simulations are coupled to experiments with the maximum temperature difference of two degrees.
+Shortly, order parameter qualities, $P^{\mathrm{headgroup}}$, $P^{sn-1}$, and $P^{sn-2}$, presents an average probability for order parameters of a given segment to locate within experimental values, taking the error bars of both values into the account.
+Qualities of SAXS form factors, $FF_{q}$, are measured from the distance between first form factor minimum in simulated vs.
+experimetal curve to avoid the effects from the simulation size dependency of the SAXS data.
+%~\cite{Databank} one should focus on the extrema $q_{z}$[\AA$^{-1}$] when comparing the simulated and experimental data.
+In particular $FFq$ only measures the difference between the first minima of the simulated and experimental curves.
+Note that in quantifying the bilayer electron densities for calculation of SAXS curves  the NMRlipids databank analysis algorithm places electrons as point charges at the atom centers without considering redistribution of charge density due to the polarizability.
+Nevertheless, we expect that this approximation does not have significant effect on the resulting SAXS form factors.
 
-Area per lipids, X-ray scattering form factors, and $S_{CH}$ order parameters automatically calculated by the NMRLipids Databank~\cite{Databank} were used.
 The $R_{1}$ rates and effective correlation times, along with the accompanying error estimates, are quantified from the trajectories as elaborated in Ref.~\citenum{Antila2021}.
@@
                                  & Drude2023 & 0      & 72       & 2239       & 0         & 303      & 300              & 200         &          \cite{kav_batuhan_2023_7916287, richard_m_venable_2023_7871949}    \\ \hline
-\multirow{3}{*}{POPE}                              & Drude2017 & 0      & 144       & 6400       & 0         & 303    & 350              & 300         &          \cite{kav_batuhan_2021_7604627}    \\                                            & Drude2023 & 0       & 72        & 2304      &  0         & 303    & 300             & 200         &           \cite{richard_m_venable_2023_7872447,kav_batuhan_2023_7916494} \\
+\multirow{3}{*}{POPE}                              & Drude2017 & 0      & 144       & 6400       & 0         & 308.15    & 350              & 300         &          \cite{kav_batuhan_2021_7604627}    \\                                            & Drude2023 & 0       & 72        & 2304      &  0         & 303    & 300             & 200         &           \cite{richard_m_venable_2023_7872447,kav_batuhan_2023_7916494} \\
                                                   & AMOEBA     & 0       & 72        & 2880      &  0        & 303     & 305.94          & 305.94      & \cite{kav_batuhan_2023_7622838} \\ \hline
@@
 \end{table}
+\todo{check the simualation settings (especially the temperature}
 %\end{adjustbox}
@@
 \begin{acknowledgement}
+
+J.J.M. thanks Huiying Chu (from the Li lab) for technical discussions and Sameer Varma for useful suggestions.
 
@@
 
+\section{Supplementary figures}
+
+\begin{figure}[!h]
+    \centering
+    %\includegraphics{Figures/order_parameters.png}
+    \includegraphics[width=0.8\textwidth]{Figures/quality_Drude2023_replicas.pdf}
+    \caption{X-ray scattering form factors, and C-H bond order parameters for headgroup and glycerol backbone, and acyl chains (from top to bottom) compared between simulations and experiments using the NMRlipids databank.
+The atom labelling is as depicted in Fig.~\ref{fig:molecules}.}
+    \label{fig:order_parameters_replicas}
+
+\end{figure}
+
+\begin{figure}[!hbt]
+	\centering
+	\includegraphics{Figures/dihedral_distributions_without_ions.eps}
+	\caption{The distributions of the torsion angles for the head group atoms.
+For each torsion angle, the upper and lower rows contain the DOPC (AMOEBA) - POPC (Drude/Drude2023) and POPE lipids, respectively.}
+	\label{fig:dihedral_no_salt}
+\end{figure}
+
 \end{document}

### 2023-10-14 manuscriptPOL.tex
Updates from Overleaf

@@
 %\title{NMRLipids VI: Polarizable Force Fields}
-\title{Do explicitly polarizable force fields give better description of a biomembrane?}
+\title{Do explicitly polarizable force fields give better description of biomembranes?}
 %\title{Comparison of polarizable lipid models to high-resolution experimental data reveals room for improvement}
@@
 \begin{abstract}
+With the increase of available computational capabilities, polarizable molecular dynamics force fields have gained popularity in modelling biomolecular systems.
+In the case of biomembranes in particular, they are expected to provide better description, eg., by capturing the varying dielectric environment across the lipid bilayer.
+Here, we use the framework established within the NMRlipids open science project and the related NMRlipids databank to test the commonly used polarizable lipid models, CHARMM-Drude and Amoeba, against high fidelity experimental data and best non-polarizable models.
+We find that, for each property tested, the best non-polarizable models surpass the polarizable ones.
+The flaws detected range from inaccuracies in describing the conformational space of lipid tails (AMOEBA) to the drastically too slow conformational dynamics of CHARMM-Drude models.
+That said, there is no model, polarizable or non-polarizable, that is consistently the best across all the properties tested.
 
-	For the initial project description, please refer to \\ https://github.com/batukav/NMRlipidsVIpolarizableFFs/blob/master/Manuscript/manuscript.pdf
+%	For the initial project description, please refer to \\ https://github.com/batukav/NMRlipidsVIpolarizableFFs/blob/master/Manuscript/manuscript.pdf
 
@@
 lamoreux2003,Antila2013 ,Lemkul2016, baker2015polarizable,jing2019polarizable }.
+
 For bilayer membrane simulation, including polarizability in the lipids is expected to provide more accurate descriptions, especially in situations pertaining to membrane binding processes, translocation of charged biomolecules across the membrane, and the behavior of molecules residing within membranes, such as membrane proteins~\cite{baker2015polarizable,Lemkul2016, li2017drude,chu2018anionicpolarizable,Lynch21,Chen2021,Melcr:2018a, melcr2019improved,nencini22}.
 %Indeed, even implicit inclusion of polarizability appears to provide a better description of cation binding to membranes~\cite{Melcr:2018a, melcr2019improved}. 
-The main lipid force fields with explicit electronic polarization are CHARMM-Drude~\cite{li2017drude, yu2023drude}, AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable}, and CHARMM-Fluctuating Charge (FQ)~\cite{lucas2012charge}.
+The available lipid force fields with explicit electronic polarization include CHARMM-Drude~\cite{li2017drude, yu2023drude}, AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable}, and CHARMM-Fluctuating Charge (FQ)~\cite{lucas2012charge}.
 The underlying treatment of polarizability differ among these models in the following way:  1) the classical Drude oscillator (CHARMM-Drude) models polarization by separating a core and a shell charge with a spring which stretches according to the envinronment giving the site a dipole moment~\cite{Lemkul2016},
@@
 Our previous efforts in benchmarking state-of-the-art non-polarizable force fields have demonstrated that the quality of predictions for important membrane properties greatly vary between different force fields, particularly for lipid headgroup conformational ensembles and ion binding affinities~\cite{Botan2015,Catte2016,Antila2019,bacle21,Antila2021,Antila2022,Databank}.
-This emphasizes the critical importance of rigorous comparative quality evaluations of simulations performed using different force fields, which helps to estimate the reliability of simulation results and select the best models for specific applications.
-This is particularly important when employing polarizable force fields, as they offer the potential to capture a broader range of physical phenomena pertinent to enhancing our understanding of in particular polar membrane regions, at an increased computational cost.
-Unfortunately, systematic benchmark studies assessing the quality of polarizable force fields are,
+Such benchmark studies are urgent also for polarizable lipid force fields,
+%are,  
 %is in great need. The few published studies reviewing the state of polarizable force fields for biomolecule simulations did not go beyond citing results from the existing studies
-to our knowledge, entirely missing from the field~\cite{inakollu2020polarisable,jing2019polarizable,baker2015polarizable}.
-Previous simulations using polarizable lipid force fields center around protein systems, with experimental validation (if performed at all) fovcusing on the properties of the proteins~\cite{jing2019polarizable, Lynch21}.
-Assessing the realism of the lipid models against experimental evidence from biophysical measurements is crucial since these membrane properties are foundational to more complicated applications, such as studies of transmembrane proteins.
-For example, it appears to be important that a wide variety of conformations is available for lipid headgroups in the model for realistic protein-lipid binding~\cite{bacle21}.
+%to our knowledge, entirely missing from the field
+%~\cite{inakollu2020polarisable,jing2019polarizable,baker2015polarizable}  
+%This emphasizes the critical importance of rigorous comparative quality evaluations of simulations performed using different force fields, which helps to estimate the reliability of simulation results and select the best models for specific applications.
+especially because their pledge is
+%important when employing polarizable force fields, as they offer the potential 
+to capture a broader range of physical phenomena pertinent to enhancing our understanding of in particular polar membrane regions.
+%, at an increased computational cost. 
+%Previous simulations using polarizable lipid force fields center around protein systems, with experimental validation (if performed at all) fovcusing on the properties of the proteins~\cite{jing2019polarizable, Lynch21}.
+Assessing quality of lipid models against experiments is crucial because membrane properties are fundamental in many applications where simulations are used to shed light on behaviour of complex biomolecular systems, such as studies of transmembrane proteins~\cite{jing2019polarizable, Lynch21}.
+For example, ion binding affinity regulates membrane surface charge and wide variety of conformations available for lipid headgroups appears to be important for modeling realistic protein-lipid binding~\cite{bacle21}.
 
-Here we set out to assess the quality of available polarizable force fields, on par with previous evaluations of non-polarizable force fields, using the resources given by the NMRLipids open collaboration framework (https://databank.nmrlipids.fi/).
-Specifically, we evaluated the quality of POPC, DOPC and POPE lipid bilayer simulations with CHARMM-Drude~\cite{li2017drude,yu2023drude} and AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable} parameters against experimental NMR and small angle X-ray scattering (SAXS) data.
-These two force fields are increasingly utilized in biomolecular simulations and have lipid parameters available for the selected systems with corresponding experimental data available from the NMRlipids databank~\cite{Databank}.
-Following the quality metrics defined in the NMRlipids project, structural properties and cation binding were evaluated against NMR order parameters and SAXS form factors~\cite{Botan2015,ollila16,Catte2016,Databank}, while lipid conformational dynamics was evaluated in comparison to data from NMR spin relaxation rate experiments~\cite{ferreira15,Antila2021}.
+Here we set out to assess the quality of actively developed polarizable lipid force fields, CHARMM-Drude~\cite{li2017drude,yu2023drude} and  AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable}, on par with previous evaluations of non-polarizable force fields, using the resources given by the NMRLipids open collaboration framework (\url{http://nmrlipids.blogspot.fi/}).
+These two force fields were selected for comparison because they are increasingly utilized in biomolecular simulations and have parameters available for the selected lipids with corresponding experimental data available from the NMRlipids databank~\cite{Databank}.
+We evaluated the structural quality of POPC, DOPC and POPE lipid bilayer simulations against experimental NMR and small angle X-ray scattering (SAXS) data using the quality metrics defined in the NMRlipids databank~\cite{Databank}.
+Cation binding to membranes was evaluated against NMR order parameters~\cite{Catte2016} and lipid conformational dynamics was evaluated in comparison to data from NMR spin relaxation rate experiments~\cite{ferreira15,Antila2021}.
 %Small angle scattering (SAXS) data is used to estimate the description of electron density profile of the membrane. 
@@
 These systems were selected based on simultaneous availability of force field parameters, and experimental data~\cite{Databank}.
-After the data production, simulation trajectories were added into the NMRlipids Databank (https://github.com/NMRLipids/Databank/) and the NMRlipids quality metrics was used to evaluate each simulation against experiments~\cite{Databank}.
-This metric measures the quality of conformational ensemble of individual lipids in two aspects: first, it evaluates agreement with C-H bond order parameters (from NMR), and second, it examines consistency of membrane dimensions compared to small angle X-ray scattering (SAXS) data~\cite{Databank,ollila16}.
+After the data production, simulation trajectories were added into the NMRlipids Databank and its quality metrics was used to evaluate each simulation against experiments~\cite{Databank}.
+This metric measures the quality of lipid bilayer simulations in two aspects: first, it evaluates the quality of conformational ensemble of individual lipids against C-H bond order parameters from NMR, and second, it examines consistency of membrane dimensions compared to small angle X-ray scattering (SAXS) data~\cite{Databank,ollila16}.
 %(measuring the electron density variation across the membrane). 
@@
     \end{tabular}
-    \caption{NMRlipids databank quality metrics~\cite{Databank} and area per lipids (APL) compared between the best simulations found from the NMRlipids databank (OPLS3e for POPC and GROMOS-CKP for POPE), simulations with CHARMM36 force field, and simulations with polarizable force fields.
+    \caption{NMRlipids databank quality metrics~\cite{Databank} and area per lipids (APL) compared between the best simulations found from the NMRlipids databank (OPLS3e~\cite{roos19} for POPC and GROMOS-CKP~\cite{Chandrasekhar03,kukol09,piggot12} for POPE), simulations with CHARMM36 force field, and simulations with polarizable force fields.
 Order  parameter quality measures, $P^{\mathrm{headgroup}}$, $P^{sn-1}$, and $P^{sn-2}$, present the average probability of order parameters within a given segment to agree with experiments (larger probabilities mean higher quality).
@@
 Direct comparisons between simulated and experimental data are visualized in Fig~\ref{fig:order_parameters} and the resulting quality metrics are shown in Table~\ref{tab:my_label}.
-Similarly to the non-polarizable CHARMM36 simulations with Gromacs, CHARMM-Drude simulation predicts slightly too packed membrane with excessively negative order parameters with respect to experiments and simulations with the highest quality in the NMRlipids databank (OPLS3e)~\cite{Databank}.
+Similarly to the non-polarizable CHARMM36 simulations, CHARMM-Drude simulation predicts slightly too packed membrane with excessively negative order parameters with respect to experiments and simulations with the highest quality in the NMRlipids databank (OPLS3e)~\cite{Databank}.
 However,
@@
 C-H bond order parameters are sensitive to the conformational ensemble but the correspondence is not unique.
-Instead, they essentially describe the average of the distributions and do not contain information of the dynamics of the confomational sampling.
+Instead, they essentially describe the average of the distributions and do not contain information of the dynamics of the conformational sampling.
 Therefore, a simulation that reproduces the order parameters has an ensemble that is only potentially correct, and even the correct ensemble may not have the experimentally observed dynamics.
@@
 Given the abundance of cations in biological systems, accurately capturing their interactions with membranes has the uttermost importance in MD simulations.
-A wealth of  experimental evidence shows that monovalent ions, such as sodium (except for lithium), exhibit very weak binding affinity of PC lipid bilayers, while multivalent ions like calcium bind more strongly~\cite{Catte2016}.
+A wealth of  experimental evidence shows that monovalent ions (except for lithium) exhibit very weak binding affinity of PC lipid bilayers, while multivalent ions like calcium bind more strongly~\cite{Catte2016}.
 However, simulations with non-polarizable force fields without any additional corrections systematically overestimate cation binding to lipid bilayers~\cite{Catte2016}.
@@
 %Finally, we address the premise that adding polarizability to force fields improves the depiction of ion binding to biomembranes in simulations. 
-To test this notion, we evaluated sodium and calcium cation binding to membranes against the experimental NMR data using the "lipid electrometer," which involves quantifing the amount of ion binding to the membrane by monitoring the change in lipid headgroup order parameters in response to increasing salt concentration~\cite{Catte2016}.
+To test this notion, we evaluated ion binding to membranes against the experimental NMR data using the "lipid electrometer," which involves quantifing the amount of ion binding to the membrane by monitoring the change in lipid headgroup order parameters in response to increasing salt concentration~\cite{Catte2016}.
 Changes in these order parameters induced by increasing NaCl and CaCl$_2$ are shown in Fig.~\ref{fig:popc_order_parameter_change} for simulations and experiments, whereas Fig.~\ref{fig:ion_density_profiles} shows the density profiles of ions with respect to the bilayer normal.
 Results from AMBER-based ECC-lipid model are shown for the reference that gives a good agreement with experiments for cation binding.
+We also show data from the non-polarizable CHARMM36 where NBFIX correction for the ion models was specifically developed to address overbinding~\cite{Venable2013,Han2018}.
 
@@
 	\includegraphics{Figures/order_parameter_change.png}
-	\caption{The change in the lipid head group order parameters $\beta$ (top row) $\alpha$ (bottom row) upon increasing ion concentration with respect to the simulations without salt. }
+	\caption{The change in the lipid head group order parameters $\beta$ (top row) $\alpha$ (bottom row) upon increasing ion concentration with respect to the simulations without salt.
+CHARMM36 and ECC data is reproduced using the Zenodo repositories Ref.~\cite{ollila_2015_32496, ollila_2015_32497, ollila_2015_32498, nencini_ricky_2019_3434396} and Ref.~\cite{melcr_josef_2017_3335503}, respectively.}
 	\label{fig:popc_order_parameter_change}
@@
 Please note that only 350~mM and 450~mM CaCl$_{2}$ concentrations are shown for the Drude and AMOEBA force fields, respectively, while 1000~mM NaCl concentration is shown for all force fields.
-CHARMM36 NaCl simulations are at 950~mM.}
+CHARMM36 NaCl simulations are at 950~mM.
+CHARMM36 data is reproduced using the Zenodo repositories Ref.~\cite{ollila_2015_32496, ollila_2015_32497, ollila_2015_32498, nencini_ricky_2019_3434396}}
     \label{fig:ion_density_profiles}
-            \todo{SAMULI: We should also show the ion density profiles along membrane normal and compare the ones with 1000 mM NaCl and 350 mM CaCl$_2$
-          (concentration in bulk water) to the best profiles in the literature (Fig. 5 in http://dx.doi.org/10.1021/acs.jpcb.7b12510).
-          This will show the actual difference in ion binding to the best performing models available.}
-          \todo{Batuhan: Data generated from http://dx.doi.org/10.1021/acs.jpcb.7b12510}
 \end{figure}
@@
 In strong contrast, experimentally one sees no sodium binding and some binding of calcium ions.
-Excessive sodium binding in the CHARMM-Drude model has been noticed before for systems containing peptides or amino acids~\cite{Ngo2019, Kav2022} and deep-eutectic solvents~\cite{shayestehpour2022ion}.
+Excessive sodium binding in the CHARMM-Drude model has been observed before for systems containing peptides or amino acids~\cite{Ngo2019, Kav2022} and deep-eutectic solvents~\cite{shayestehpour2022ion}.
 Excessive calcium interactions, in line with what we observe here for the lipids, have also been reported~\cite{Tan2022}.
@@
 Particularly, increasing $\alpha$ carbon order parameters and forking upon ion binding are not observed in experiments.
-This is in contrast with the results from non-polarizable simulations, where experimentally observed decrease of order parameters to more negative values upon ion binding were observed in all simulations, even though the binding affinity was often inaccurately predicted~\cite{Catte2016}.
+This is in contrast with the results from previous bechmarking of non-polarizable simulations~\cite{Catte2016}, where experimentally observed decrease of order parameters to more negative values upon ion binding were observed in all simulations, even though the binding affinity was often inaccurately predicted.
 This qualitative discrepancy in CHARMM-Drude2017 simulations most likely results from incorrect lipid headgroup conformational ensemble (see previous sections), which leads to inaccuracies in the structural response of the ensemble to ion binding.
@@
 %also appears to be better represented in the updated model. The 
-ions penetrate slightly
+ions penetrate 
 %reveal that binding of calcium has moved 
 deeper into the bilayer, and there is a stronger formation of ionic double layer by subsequent attraction of chloride into the bilayer.  
+In the non-polarizable counterpart, CHARMM36 model with the NBFIX correction, sodium biding is similar to the polarizable models outside small accumulation peak in the phosphate region, and stronger subsequent accumulation of anions outside the phosphate region.
+The structural response well in line with the experiments.
+The divalent cation biding for CHARMM36 is weakest of all models with binding depth, and following anion binding, quite similar the ECC.
+However, the structural response of the $\alpha$ carbons is weaker than in ECC and experiments.
+Contrasting the CHARMM36 with the CHARMM-Drude models and the ECC model results suggests that so far correcting the ion biding by scaling the Lennard-Jones parameters (NBFIX) of the non-polarizable models has resulted in slightly more realistic ion binding than introducing explicit polarizability to the model.
 %As opposed to the CHARMM36-Drude model, the order parameters of the AMOEBA model are almost completely unaffected by the ion addition, both for NaCl and CaCl$_{2}$.
@@
 In AMOEBA simulations, sodium binds weakly and does not affect order parameters, consistent with experiments and ECC-lipid simulations.
-However, order parameters do not change upon calcium binding in accordance with experiments and other simulations.
-This can be explained by the binding position of calcium, which is outside phosphate density peak in AMOEBA simulations, whereas calcium penetrates to phosphate region or deeper in other simulations and thereby reorients the headgroup dipole, giving rise to changes in order parameters~\cite{Catte2016}.
+However, order parameters also do not change upon calcium binding contrasting the experiments.
+This can be explained by the binding position of calcium, which is outside phosphate density peak in AMOEBA simulations.
+In simulations where calcium penetrates to phosphate region or deeper in similar amounts, it reorients the headgroup dipole, giving rise to changes in order parameters~\cite{Catte2016}.
 
@@
 This is of note, as polarizable force fields are often assumed superior and concequently a popupar choice in modelling, eg, ion channels~\cite{sun2017 jing2019polarizable, klesse2020induced}
-\todo{I think there is a mistake in the C36 data, firsly, the concentration should probably be 950mM instead of 1000mM and secondly the monovalent anion and cation curves look identical} ~\todo{batuhan: I fxed the anion and cation curves.
--> @batuhan will plot the c36 data in a different color for the 950mm case.s}.
-
 %\begin{figure}[!hbt]
@@
 
-In conclusion, incorporating explicit polarizability as done in CHARMM-Drude or AMOEBA force fields do not necessarily lead to improved description of cation binding to phospolipid membranes.
+In conclusion, incorporating explicit polarizability as done in CHARMM-Drude or AMOEBA force fields does not necessarily lead to improved description of cation binding to phospolipid membranes.
 However, it is difficult to isolate the explicit influence of polarizability per se on ion binding based on the results presented here.
@@
 We have rigorously compared simulation output of model lipid bilayers using polarizable lipid force fields with high-resolution experimental data as well as results from best-performing non-polarizable force fields.
-From this exercise, it is evident that current polarizable force fields for lipids fall short in important aspects.
-In most cases, the best non-polarizable models exhibits superior performance and accuracy across all the categories investigated in this study.
+From this, it is evident that current polarizable force fields for lipids fall short in important aspects.
+In most cases, the best non-polarizable models exhibits superior performance and accuracy in the categories investigated in this study.
 
 Evaluation order parameters and scattering form factors reveals that the AMOEBA force field has difficulties in producing membranes with reasonable physical properties.
-In constrast, the CHARMM-Drude2023 model has improved performance in this regard, notably surpassing its predecessor, CHARMM-Drude2017.
+In contrast, the CHARMM-Drude2023 model has improved performance in this regard, notably surpassing its predecessor, CHARMM-Drude2017.
 Nevertheless, top-performing non-polarizable models generally reproduce experimental conformations (order parameters) better, although not perfectly so.
@@
 In regard to conformational dynamics, the AMOEBA force field produces results, specifically in the headgroup region of PC lipids, that are either on par with or better than the best non-polarizable force fields.
-However, the CHARMM-Drude models exhibit an extremely slow dynamics, which can lead to deficient sampling and a substantial increase in simulation simulation time when using this model.
+However, the CHARMM-Drude models exhibit an extremely slow dynamics, which can lead to deficient sampling and a substantial increase in required simulation time when using this model.
 
-In general, the ECC model exhibits thet most accurate representation of ion binding, even though it lacks an explicit description of polarization.
-The non-polarizable CHARMM36, when used with the NBFix ion model designed to provide more accurate binding, does XXX.
-The polarizable AMOEBA and Drude-2023 models show similar ion binding levels to the ECC model, but they exhibit variation in binding depth and the consequent structural response of the lipids do not align with experiments.
+In general, the ECC model gives the most accurate representation of ion binding, even though it lacks an explicit description of polarization.
+The non-polarizable CHARMM36, when used with the NBFix ion model designed to provide more accurate binding, also rivals its polarizable counterparts.
+The polarizable AMOEBA and Drude-2023 models have similar ion binding levels to the ECC model, but they exhibit variation in binding depth, and the consequent structural response of the lipids do not align with experiments.
 The incorrect response to ion binding likely connects to the the other inaccuracies described above, including the misrepresentation of overall membrane structure and diversity in the conformational ensemble, as well as slow dynamics that lead to poor convergence of simulations.
@@
 Table~\ref{table:sim_details} for the full list of simulations conducted for this study and links to the openly available trajectories.
-Data not mentioned here was obtained by analysis pre-existing trajectories in the NMRlipids databank.
+Data not mentioned here was obtained by analysing pre-existing trajectories from the NMRlipids databank.
 
@@
 Qualities of SAXS form factors, $FF_{q}$, are measured from the distance between first form factor minimum in simulated vs.
-experimetal curve to avoid the effects from the simulation size dependency of the SAXS data.
+experimental curve to avoid the effects from the simulation size dependency of the SAXS data.
 %~\cite{Databank} one should focus on the extrema $q_{z}$[\AA$^{-1}$] when comparing the simulated and experimental data.
@@
 Simulated time is listed in column t$_{sim}$ and time used for analysis in t$_{analysis}$.
-Column "files" gives the reference to open access simulation data.}
+Column "files" gives the reference to open access simulation data.
+The salt concentrations calculated as [salt]=N$_c$×[water] / N$_w$, where [water] = 55.5 M}
 \label{table:sim_details}

### 2023-10-14 cover.tex

\documentclass[11pt]{letter}
\usepackage{graphicx,baskervald,microtype}
\usepackage{hyperref,amsmath,xcolor}
\hypersetup{
  colorlinks=true,
  linkcolor=black,
  urlcolor=black
}
\usepackage{pgfplots,marginnote}
\usepackage[top=0.5in,bottom=0.5in,left=1in,right=1in]{geometry}
%\newgeometry{margin=2.5cm}
\newcommand{\omamargin}[1]{\marginnote{\textbf{#1}}[7pt]}

\begin{document}
\reversemarginpar
\pagestyle{empty}
\noindent Dr. XXX\\
\noindent Department of XXX \\
\noindent University of XXX, XXX \\
\noindent email: XXX\\
\noindent +XXX\,XXX\,XXX\,XXX \\

Dear Editor,

Please find enclosed our manuscript titled
{XXX}, which we are submitting for your consideration to be published in \textit{XXX}.

In the manuscript, we answer the question "are polarizable molecular dynamics force fields better for modelling biomembranes (lipid bilayers) than their non-polarizable counterparts and in which ways". This has long thought to be the case as polarizable models have the potential for more accurate description of ionic interactions and the varying dielectric environment across a lipid bilayer. Systematic benchmarking of the models against both high-resolution experimental data and non-polarizable model has, however, been lacking.

We seek to fill this gap by using the NMRlipids open collaboration framework and the related NMRlipids databank. We test the models againts experimental NMR order parameters (describing the conformational space of lipid molecules), small-angle X-ray scattering curves (depicting the fourier tranformation of the electron density across the membrane), NMR relaxation and correlation rates (describing the conformational dynamics), and the change of NMR order parameters when salt is added to the system, which indicate the strength of ion binding. We find that across all these test, the best non-polarizable models beat the polarizable lipid models tested: AMOEBA, CHARMM-Drude2017/CHARMM-Drude2023. Although the polarizable models give good description in some cases, such as the ability of AMOEBA to capture monovalent ion binding and lipid headgroup dynamics, they also have some serious flaws, like the faulty description of the lipid tails in AMOEBA and vastly too slow conformational dynamics of CHARMM-Drude models. It is particularly important to bring attention the last result, since it means one has to run the simulation longer to obtain good quality data.

We believe the test performed in this paper will provide useful reference for anyone looking to use polarizable force field for membrane simulations. We hope it will also encourage the developers of these models to use the experimental data in the parametrization and/or carefully benchmach their models against it. The paper exemplifies the benefits of doing so by pointing out the improvement from the CHARMM-Drude2017 model to the 2023 version.

 In summary, we trust our manuscript fulfills the high standards for publication in %the \textit{Resource} section of
\textit{XXX} and leave it to your expert consideration.\\

%We now leave it for your own consideration.

Sincerely yours,

on behalf of the authors,\\

Dr. XXX

\end{document}

### 2023-10-27 manuscriptPOL.tex
Update manuscript

@@
 In the case of biomembranes in particular, they are expected to provide better description, eg., by capturing the varying dielectric environment across the lipid bilayer.
-Here, we use the framework established within the NMRlipids open science project and the related NMRlipids databank to test the commonly used polarizable lipid models, CHARMM-Drude and Amoeba, against high fidelity experimental data and best non-polarizable models.
-We find that, for each property tested, the best non-polarizable models surpass the polarizable ones.
+Here, we use the framework established within the NMRlipids open science project and the related NMRlipids databank to test the commonly used polarizable lipid models, CHARMM-Drude and AMOEBA, against high fidelity experimental data and best non-polarizable models.
+%We find that, for each property tested, the best non-polarizable models surpass the polarizable ones.
 The flaws detected range from inaccuracies in describing the conformational space of lipid tails (AMOEBA) to the drastically too slow conformational dynamics of CHARMM-Drude models.
@@
 
-The dielectric environment in and around a lipid bilayer varies dramatically as you move from the water phase through the interfacial headgroup region to the membrane interior.
-The water phase highly polar, while the interfacial region contains dipolar headgroups, and the membrane interior is predominantly hydrophobic and nonpolar due to the aliphatic lipid tails.
+%Molecular dynamics (MD) simulations, due to their ability to resolve Angstrom and femtosecond resolution dynamics and more accurate force fields, are an invaluable tool to understand many biological systems including ribosomes, ion-channels, actin filaments, amyloid-$\beta$, and G-coupled receptors.
+On the other hand, MD simulations' predictive power depends strongly on how MD force fields model the interactions between the atoms in the systems.
+Many MD force fields treat electronic polarizability, i.e.
+the electronic distribution's ability to respond to an electric field, rather implicitly.
+But, molecules are polarizable and they respond to the change of electric field in their environments.
+For instance, the electronic distribution of a peptide varies according to its conformation, the electronic distribution of a ligand changes upon binding to a protein, and the dielectric environment in and around a lipid bilayer varies dramatically as one moves from the water phase through the interfacial headgroup region to the membrane interior.
+A fundamental question in molecular dynamics is, hence, how to accurately include electronic polarizability into its force fields.
+A more practical question is, when electronic polarizability is in some way included into the force fields, how accurate do the force fields become describing the biomolecular systems.
+Here, we aim to tackle the latter and evaluate the accuracy of common state-of-the-art polarizable force fields in describing lipid bilayers.
+Are current explicitly polarizable force fields more accurate than implicitly or non-polarizable force fields?
 
-In conventional molecular dynamics (MD) simulations, electrostatic interactions are described by assigning static point charges to the atoms and molecules at the beginning of the simulation based on the force field definitions.
-Dynamic effects arising from molecular polarizability are traditionally not explicitly included and are only considered in an averaged fashion in the force field parametrization process when fitting to macroscopic observables.
+In conventional molecular dynamics (MD) simulations, electrostatic interactions are described by assigning static point charges to the atoms and molecules at the beginning of the simulation as defined in the inputted model---the force field.
+Dynamic effects arising from molecular polarizability are traditionally not explicitly included and only considered in an averaged fashion via force field parameterization process where parameters are obtained by fitting to macroscopic observables or to ab initio calculations.
 Significant efforts have been dedicated to introduce explicit polarizability into MD simulations, in the hopes of reaching a more accurate representation of the system~\cite{Thole1981,ando2001stable,Grossfield2003,
 lamoreux2003,Antila2013 ,Lemkul2016, baker2015polarizable,jing2019polarizable }.
+This effort has been motivated by the plethora of interactions where polarization is perceived to be a key contribution to correctly describing the system.
+These include description of water, ion hydration and ion binding to molecules, cation--$\pi$ and $\pi$-$\pi$ interactions~\cite{inakollu2020polarisable}.
+These lower-level interactions also translate to the behaviour of larger-scale biological systems, such as ion channels, where ion-selectivity and ion currents may be affected by polarization~\cite{roux2011ion, klesse2020induced, prajapati2020computational,Yue2022}, and telomeric DNA~\cite{Salsbury22} where the conformations adopted are mediated by ionic interactions.
 
-For bilayer membrane simulation, including polarizability in the lipids is expected to provide more accurate descriptions, especially in situations pertaining to membrane binding processes, translocation of charged biomolecules across the membrane, and the behavior of molecules residing within membranes, such as membrane proteins~\cite{baker2015polarizable,Lemkul2016, li2017drude,chu2018anionicpolarizable,Lynch21,Chen2021,Melcr:2018a, melcr2019improved,nencini22}.
-%Indeed, even implicit inclusion of polarizability appears to provide a better description of cation binding to membranes~\cite{Melcr:2018a, melcr2019improved}. 
+For bilayer membrane simulation, specifically,  the molecular (dielectric) environment varies dramatically while crossing the membrane from the water phase across the dipolar/charged lipid headgroup interface to the hydrophobic tail region.
+Therefore, including polarizability in the lipids is expected to provide more accurate description, especially in situations pertaining to membrane binding processes, membrane potential and translocation of charged biomolecules across the membrane, and the behavior of molecules residing within membranes, such as membrane proteins~\cite{baker2015polarizable,Lemkul2016, li2017drude,chu2018anionicpolarizable,Lynch21,Chen2021,Melcr:2018a, melcr2019improved,nencini22}.
+
 The available lipid force fields with explicit electronic polarization include CHARMM-Drude~\cite{li2017drude, yu2023drude}, AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable}, and CHARMM-Fluctuating Charge (FQ)~\cite{lucas2012charge}.
-The underlying treatment of polarizability differ among these models in the following way:  1) the classical Drude oscillator (CHARMM-Drude) models polarization by separating a core and a shell charge with a spring which stretches according to the envinronment giving the site a dipole moment~\cite{Lemkul2016},
+The underlying strategy for including explicit polarizability differ among these models:  1) the classical Drude oscillator (CHARMM-Drude) models polarization by separating a core and a shell charge with a spring which stretches according to the environment giving the site a fluctuating dipole moment~\cite{Lemkul2016},
 2) the induced point dipole/multipole (AMOEBA) uses polarizable point dipoles placed on chosen sites of the molecule~\cite{ponder2010current},
-and 3) the electronegativity equilization (FQ) employs atomic charges that are no longer constant but can change according to the electronegativities of the molecule atoms and the electric fields from the molecular enviroment~\cite{patel2004charmm}.
+and 3) the electronegativity equilization (FQ) employs atomic charges that are no longer constant but can change (redistribute within the molecule) according to the electronegativities of the molecule atoms and the electric fields from the molecular environment~\cite{patel2004charmm}.
 All of these approaches result in an increasing computational cost, e.g, by introducing new types of interactions, more interactions sites, or by requiring a shorter time step in the simulations.
-As an alternative approach, electronic continuum correction (ECC) has been proposed as a computational efficient method to implicitly include polarizability by scaling the atom partial charges~\cite{Melcr:2018a,dijon20,Antila2021}.
+As an alternative approach, electronic continuum correction (ECC) has been proposed as a computational efficient method to implicitly include polarizability by scaling the atom partial charges~\cite{Melcr:2018a,dijon20}.
 
@@
 
-Our previous efforts in benchmarking state-of-the-art non-polarizable force fields have demonstrated that the quality of predictions for important membrane properties greatly vary between different force fields, particularly for lipid headgroup conformational ensembles and ion binding affinities~\cite{Botan2015,Catte2016,Antila2019,bacle21,Antila2021,Antila2022,Databank}.
-Such benchmark studies are urgent also for polarizable lipid force fields,
-%are,  
-%is in great need. The few published studies reviewing the state of polarizable force fields for biomolecule simulations did not go beyond citing results from the existing studies
-%to our knowledge, entirely missing from the field
-%~\cite{inakollu2020polarisable,jing2019polarizable,baker2015polarizable}  
-%This emphasizes the critical importance of rigorous comparative quality evaluations of simulations performed using different force fields, which helps to estimate the reliability of simulation results and select the best models for specific applications.
-especially because their pledge is
+Our previous efforts in benchmarking state-of-the-art non-polarizable lipid force fields have demonstrated that the quality of predictions for important membrane properties greatly vary between different force fields, particularly for lipid headgroup conformational ensembles and ion binding affinities~\cite{Botan2015,Catte2016,Antila2019,bacle21,Antila2021,Antila2022,Databank}.
+The ability to capture these membrane properties correctly is important on its own but also creates the basis for the description of more complex systems: for example, ion binding affinity regulates membrane surface charge and wide variety of conformations available for lipid headgroups appears to be important for modeling realistic protein-lipid binding~\cite{bacle21}.
+Therefore, such benchmark studies are urgently needed also for polarizable lipid force fields, considering  the increased computational cost they come with, and most importantly, their pledge
 %important when employing polarizable force fields, as they offer the potential 
 to capture a broader range of physical phenomena pertinent to enhancing our understanding of in particular polar membrane regions.
-%, at an increased computational cost. 
-%Previous simulations using polarizable lipid force fields center around protein systems, with experimental validation (if performed at all) fovcusing on the properties of the proteins~\cite{jing2019polarizable, Lynch21}.
-Assessing quality of lipid models against experiments is crucial because membrane properties are fundamental in many applications where simulations are used to shed light on behaviour of complex biomolecular systems, such as studies of transmembrane proteins~\cite{jing2019polarizable, Lynch21}.
-For example, ion binding affinity regulates membrane surface charge and wide variety of conformations available for lipid headgroups appears to be important for modeling realistic protein-lipid binding~\cite{bacle21}.
 
-Here we set out to assess the quality of actively developed polarizable lipid force fields, CHARMM-Drude~\cite{li2017drude,yu2023drude} and  AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable}, on par with previous evaluations of non-polarizable force fields, using the resources given by the NMRLipids open collaboration framework (\url{http://nmrlipids.blogspot.fi/}).
-These two force fields were selected for comparison because they are increasingly utilized in biomolecular simulations and have parameters available for the selected lipids with corresponding experimental data available from the NMRlipids databank~\cite{Databank}.
-We evaluated the structural quality of POPC, DOPC and POPE lipid bilayer simulations against experimental NMR and small angle X-ray scattering (SAXS) data using the quality metrics defined in the NMRlipids databank~\cite{Databank}.
-Cation binding to membranes was evaluated against NMR order parameters~\cite{Catte2016} and lipid conformational dynamics was evaluated in comparison to data from NMR spin relaxation rate experiments~\cite{ferreira15,Antila2021}.
-%Small angle scattering (SAXS) data is used to estimate the description of electron density profile of the membrane. 
-Furthermore, we compare the results to those obtained from the best performing non-polarizable simulations from the NMRlipids Databank\cite{Databank}.
-
- %n lipid modelling, the first two have surpassed the third in popularity.
+Here we assess the quality of actively developed polarizable lipid force fields, CHARMM-Drude~\cite{li2017drude,yu2023drude} and  AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable}, using the resources and the framework from the NMRLipids open collaboration (\url{http://nmrlipids.blogspot.fi/}).
+These two force fields were selected for comparison because they are increasingly utilized in biomolecular simulations and have parameters available for several lipids for which corresponding experimental data is available in NMRlipids databank~\cite{Databank}.
+We assess the structural quality of POPC, DOPC and POPE lipid bilayer simulations against experimental NMR and small angle X-ray scattering (SAXS) data using the quality metrics defined in the NMRlipids databank~\cite{Databank}.
+Cation binding to membranes was evaluated against NMR order parameters~\cite{Catte2016} and lipid conformational dynamics is benchmarked to data from NMR spin relaxation rate experiments~\cite{ferreira15,Antila2021}.
+Furthermore, for each experimental benchmark we compare the results from polarizable models to those from the best performing non-polarizable simulations obtained from the NMRlipids Databank\cite{Databank}.
 
@@
         POPC & CHARMM36     & 0.70 & 0.54 & 0.69 & 1.16 & 65.0\\
-        POPC & CHARMM-Drude & 0.52 & 0.29 & 0.53 & 1.06 & 62.5\\
+        POPC & CHARMM-Drude2017 & 0.52 & 0.29 & 0.53 & 1.06 & 62.5\\
         POPC & CHARMM-Drude2023 & 0.63 & 0.60 & 0.57 & 0.96 & 64.5\\
@@
         POPE & CHARMM36   & 0.54 & 0.52 & 0.27 & 1.30 & 57.2 \\
-        POPE & CHARMM-Drude & 0.06 & 0.53 & 0.27 & 0.80 & 56.6  \\
+        POPE & CHARMM-Drude2017 & 0.06 & 0.53 & 0.27 & 0.80 & 56.6  \\
         POPE & CHARMM-Drude2023 & 0.28 & 0.59 & 0.54 & 0.00 & 61.4  \\
@@
     %\includegraphics{Figures/order_parameters.png}
-    \includegraphics[width=0.8\textwidth]{Figures/quality.pdf}
+    \includegraphics[width=0.9\textwidth]{Figures/quality.pdf}
     \caption{X-ray scattering form factors, and C-H bond order parameters for headgroup and glycerol backbone, and acyl chains (from top to bottom) compared between simulations and experiments using the NMRlipids databank.
@@
 Direct comparisons between simulated and experimental data are visualized in Fig~\ref{fig:order_parameters} and the resulting quality metrics are shown in Table~\ref{tab:my_label}.
-Similarly to the non-polarizable CHARMM36 simulations, CHARMM-Drude simulation predicts slightly too packed membrane with excessively negative order parameters with respect to experiments and simulations with the highest quality in the NMRlipids databank (OPLS3e)~\cite{Databank}.
+Similarly to the non-polarizable CHARMM36 simulations, CHARMM-Drude2017 simulation predicts slightly too packed membrane with excessively negative order parameters with respect to experiments and simulations with the highest quality in the NMRlipids databank (OPLS3e)~\cite{Databank}.
 However,
 %consistently outperform the polarizable force fields with respect to the NMR measurables for all parts of the molecule (higher $P$-values). For older CHARMM-Drude force field, 
-the quality of headgroup conformations in the CHARMM-Drude simulations is worse than in its non-polarizable counterpart.
+the quality of headgroup conformations in the CHARMM-Drude2017 simulations is worse than in its non-polarizable counterpart.
 This is likely because headgroup and glycerol backbone parameters were optimized to reproduce absolute average values of the experimentally determined order parameters, instead of taking into account the order parameter forking (meaning measurably different order parameters for different C--H bonds at a single carbon atom) and order parameter sign~\cite{Antila2022}.
@@
 %In the tail regions, the CHARMM-Drude2023 appears to be the best out of the polarizable models. 
-
+%
+%
 The unsatisfactory description of the lipid tail region and the area per lipid is further reflected in the inability of AMOEBA to capture the SAXS curves.
@@
 membrane properties. 
-The higher $FFq$ values (Table~\ref{tab:my_label}) of the polarizable force fields in general indicate that they are not able confer consistency with the SAXS data as well as the best non-polarizable force fields (OPLS3e).
+%The higher $FFq$ values (Table~\ref{tab:my_label}) of the polarizable force fields in general indicate that they are not able confer consistency with the SAXS data as well as the best non-polarizable force fields (OPLS3e).
 That said, CHARMM-Drude2023 is on par with or better than its non-polarizable counterpart CHARMM36 and the best out of all the models here in capturing the POPE SAXS data, benefiting from the improvements made in the description of acyl chains.
@@
 However, simulations with non-polarizable force fields without any additional corrections systematically overestimate cation binding to lipid bilayers~\cite{Catte2016}.
-Implicit inclusion of electronic continuum correction (ECC) to both ion and lipid parameters can substantially improve the situation~\cite{Melcr:2018a,melcr2019improved,bacle21}, suggesting that electronic polarizability plays an important role in ion binding to membranes, which might further lead one to expect that that simulations with polarizable force fields will perform better in this aspect and more accurately describe ion binding to membranes.
+Implicit inclusion of electronic continuum correction (ECC) to both ion and lipid parameters can substantially improve the situation~\cite{Melcr:2018a,melcr2019improved,bacle21}, suggesting that electronic polarizability plays an important role in ion binding to membranes, which might further lead one to expect that simulations with polarizable force fields will perform better in this aspect and more accurately describe ion binding to membranes.
 %Finally, we address the premise that adding polarizability to force fields improves the depiction of ion binding to biomembranes in simulations. 
 To test this notion, we evaluated ion binding to membranes against the experimental NMR data using the "lipid electrometer," which involves quantifing the amount of ion binding to the membrane by monitoring the change in lipid headgroup order parameters in response to increasing salt concentration~\cite{Catte2016}.
-Changes in these order parameters induced by increasing NaCl and CaCl$_2$ are shown in Fig.~\ref{fig:popc_order_parameter_change} for simulations and experiments, whereas Fig.~\ref{fig:ion_density_profiles} shows the density profiles of ions with respect to the bilayer normal.
-Results from AMBER-based ECC-lipid model are shown for the reference that gives a good agreement with experiments for cation binding.
+Changes in these order parameters induced by increasing NaCl or CaCl$_2$ concentration are shown in Fig.~\ref{fig:popc_order_parameter_change} for simulations and experiments, whereas Fig.~\ref{fig:ion_density_profiles} shows the density profiles of ions with respect to the bilayer normal.
+Results from AMBER-based ECC-lipid model are shown for the reference that gives a good agreement with experiments for cation binding~\cite{Melcr:2018a}.
 We also show data from the non-polarizable CHARMM36 where NBFIX correction for the ion models was specifically developed to address overbinding~\cite{Venable2013,Han2018}.
@@
 In strong contrast, experimentally one sees no sodium binding and some binding of calcium ions.
-Excessive sodium binding in the CHARMM-Drude model has been observed before for systems containing peptides or amino acids~\cite{Ngo2019, Kav2022} and deep-eutectic solvents~\cite{shayestehpour2022ion}.
-Excessive calcium interactions, in line with what we observe here for the lipids, have also been reported~\cite{Tan2022}.
-Furthermore, the response of headgroup order parameters to the bound ions is not in qualitative agreement with experiments for CHARMM-Drude2017 simulations according to Fig.~\ref{fig:popc_order_parameter_change}.
+Furthermore, the response of headgroup order parameters to the bound ions is not in qualitative agreement with experiments for CHARMM-Drude2017 simulations in Fig.~\ref{fig:popc_order_parameter_change}.
 Particularly, increasing $\alpha$ carbon order parameters and forking upon ion binding are not observed in experiments.
@@
 This qualitative discrepancy in CHARMM-Drude2017 simulations most likely results from incorrect lipid headgroup conformational ensemble (see previous sections), which leads to inaccuracies in the structural response of the ensemble to ion binding.
+Excessive sodium binding in the CHARMM-Drude model has been observed before for systems containing peptides or amino acids~\cite{Ngo2019, Kav2022} and deep-eutectic solvents~\cite{shayestehpour2022ion}.
+%Excessive calcium interactions, in line with what we observe here for the lipids, have also been reported~\cite{Tan2022}.
 
-In simulations with CHARMM-Drude2023 parameters, sodium ion binding is in line with ECC-lipid simulations when comparing density profiles,
+In simulations with CHARMM-Drude2023 parameters, sodium and calcium ion binding are in line with ECC-lipid simulations when comparing density profiles, although calcium binding affinity is slightly larger, ions penetrate deeper into the bilayer, and there is a stronger formation of ionic double layer by subsequent attraction of chloride into the bilayer in CHARMM-Drude2023 when compared with ECC-lipids.
 %force field: the change of the order parameters is diminished to be more in line with experiments and only the some accumulation of sodium ions near the phosphate oxygen is seen in the ion distributions.
 %However, this the ion layering is most prominently present in the original CHARMM36 force field, for for sodium and calcium ions. 
-but addition of sodium or calcium ions seems to induce forking of order parameters in contrast to experiments. 
+However, sodium or calcium ion binding seems to induce forking of lipid headgroup order parameters in CHARMM-Drude2023 simulations in contrast to experiments. 
 %at the $\beta$ carbon in particular. 
@@
 %, which would also not be fully reflected in the calculation of error estimates for the order parameters. 
-Based on density profiles, calcium binding in CHARMM-Drude2023 simulations is in line with the ECC-lipid model, although binding affinity is slightly larger, 
-%also appears to be better represented in the updated model. The 
-ions penetrate 
-%reveal that binding of calcium has moved 
-deeper into the bilayer, and there is a stronger formation of ionic double layer by subsequent attraction of chloride into the bilayer.  
-In the non-polarizable counterpart, CHARMM36 model with the NBFIX correction, sodium biding is similar to the polarizable models outside small accumulation peak in the phosphate region, and stronger subsequent accumulation of anions outside the phosphate region.
+%
+%Based on density profiles, calcium binding in CHARMM-Drude2023 simulations is in line with the ECC-lipid model,
+
+%In the non-polarizable counterpart, CHARMM36 model with the NBFIX correction, sodium biding is similar to the polarizable models outside small accumulation peak in the phosphate region, and stronger subsequent accumulation of anions outside the phosphate region.
 The structural response well in line with the experiments.
@@
 Contrasting the CHARMM36 with the CHARMM-Drude models and the ECC model results suggests that so far correcting the ion biding by scaling the Lennard-Jones parameters (NBFIX) of the non-polarizable models has resulted in slightly more realistic ion binding than introducing explicit polarizability to the model.
+%
 %As opposed to the CHARMM36-Drude model, the order parameters of the AMOEBA model are almost completely unaffected by the ion addition, both for NaCl and CaCl$_{2}$.
@@
 While this is in line with experiments regarding sodium ions, there should be some binding of calcium ions to the membrane.
+
 In AMOEBA simulations, sodium binds weakly and does not affect order parameters, consistent with experiments and ECC-lipid simulations.
-However, order parameters also do not change upon calcium binding contrasting the experiments.
+Calcium binding affinity is similar to ECC-lipids, but order parameters also do not change upon  binding contrasting the experiments.
 This can be explained by the binding position of calcium, which is outside phosphate density peak in AMOEBA simulations.
-In simulations where calcium penetrates to phosphate region or deeper in similar amounts, it reorients the headgroup dipole, giving rise to changes in order parameters~\cite{Catte2016}.
+In other simulations calcium penetrates to phosphate region or deeper and reorients the headgroup dipole, giving rise to changes in order parameters in line with the experimental data~\cite{Catte2016}.
 
@@
 
-In conclusion, incorporating explicit polarizability as done in CHARMM-Drude or AMOEBA force fields does not necessarily lead to improved description of cation binding to phospolipid membranes.
-However, it is difficult to isolate the explicit influence of polarizability per se on ion binding based on the results presented here.
-What is clear, however, is that the headgroup conformational ensemble, which presumably plays a role in ion binding, is poorly described by the CHARMM-Drude parameters.
+In conclusion, incorporating explicit polarizability as implemented in CHARMM-Drude or AMOEBA parameters used here do not necessarily lead to improved description of cation binding to phospolipid membranes.
+However, these parameters do not correctly capture the response of lipid headgroup to cation binding to membrane, which most likely arises from inaccuracies in lipid parameters.
+Because such inaccuracies can also affect ion binding, it is difficult to isolate the explicit influence of polarizability per se on ion binding based on the results presented here.
+%What is clear, however, is that the headgroup conformational ensemble, which presumably plays a role in ion binding, is poorly described by the CHARMM-Drude parameters.
 In addition, the AMOEBA parameters struggle to capture the basic membrane properties that are reproduced by all current state-of-the-art MD simulation models.
+Nevertheless, CHARMM-Drude2023 predicts higher affinity for calcium than for sodium, which is in line with the wide range of experimental data~\cite{Catte2016}.
+This is a major improvement to non-polarizable CHARMM36 and CHARMM-Drude2017 which predict approximately similar binding to these ions in contrast to experiments.
 
-\subsection{Discussion}
+%\subsection{Discussion}
 
-\textcolor{red}{For the ion binding discussion: drude calcium overbinds to proteins. this is known here https://pubs.acs.org/doi/full/10.1021/acsphyschemau.1c00039
-drude sodium overbinds to proteins, as shown here https://pubs.rsc.org/en/content/articlepdf/2022/ra/d2ra01478e and https://onlinelibrary.wiley.com/doi/abs/10.1002/adts.201800106.
+%\textcolor{red}{For the ion binding discussion: drude calcium overbinds to proteins. this is known here https://pubs.acs.org/doi/full/10.1021/acsphyschemau.1c00039
+%drude sodium overbinds to proteins, as shown here https://pubs.rsc.org/en/content/articlepdf/2022/ra/d2ra01478e and https://onlinelibrary.wiley.com/doi/abs/10.1002/adts.201800106.
 Venable also cautioned us against using any ion parameters with the drude force fields.
@@
 \subsection{Conclusions}
-We have rigorously compared simulation output of model lipid bilayers using polarizable lipid force fields with high-resolution experimental data as well as results from best-performing non-polarizable force fields.
-From this, it is evident that current polarizable force fields for lipids fall short in important aspects.
+Including electronic polarizability to MD simulation models of membranes is expected to improve the description of bilayer polar regions.
+However, performance of membrane simulations have not been evaluated with equal footing with non-polarizable membranes.
+Here, we used the quality evaluation metrics defined in the NMRlipids databank~\cite{Databank} together with additional analyzes on dynamics and ion binding to evaluate quality of two available polarizable lipid parameters, CHARMM-Drude~\cite{li2017drude, yu2023drude} and AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable}, against
+%We have rigorously compared simulation output of model lipid bilayers using polarizable lipid force fields with high-resolution 
+experimental data and results from the best-performing non-polarizable force fields. 
+Considering the complexity and cost of simulations with polarizable models, it is crucial to understand their accuracy with respect to experiments and best non-polarizable models when planning simulations or evaluating the reliability of simulation results.
+
+%From this, it is evident that current polarizable force fields for lipids fall short in important aspects.
 In most cases, the best non-polarizable models exhibits superior performance and accuracy in the categories investigated in this study.
+Our comparisons of lipid conformations and dynamics revealed room for improvement in current polarizable parameters to reach the level of the best currently available non-polarizable parameters.
+CHARMM36-Drude models predict slightly too ordered membrane and struggle to correctly capture headgroup conformations and dynamics.
+The former issue is present also in non-polarizable CHARMM36, while slow headgroup dynamics and incorrect response to ion binding is present only in polarizable CHARMM36-Drude simulations.
+%In contrast, the CHARMM-Drude2023 model has improved performance in this regard, notably surpassing its predecessor, CHARMM-Drude2017.
+Nevertheless, top-performing non-polarizable models generally reproduce experimental conformations (order parameters) better, although not perfectly so.
+%However, the CHARMM-Drude models exhibit an extremely slow dynamics, which can lead to deficient sampling and a substantial increase in required simulation time when using this model.
+%
+%Evaluation order parameters and scattering form factors reveals that 
+The tested AMOEBA parameters have difficulties to capture membrane ordering in acyl chain region,  %membranes with reasonable physical properties. 
+%In regard to conformational dynamics, the AMOEBA force field produces results, specifically in 
+yet the PC and PE headgroup properties have similar or almost simular quality 
+%are either on par with or better 
+than the best non-polarizable force fields.
 
-Evaluation order parameters and scattering form factors reveals that the AMOEBA force field has difficulties in producing membranes with reasonable physical properties.
-In contrast, the CHARMM-Drude2023 model has improved performance in this regard, notably surpassing its predecessor, CHARMM-Drude2017.
-Nevertheless, top-performing non-polarizable models generally reproduce experimental conformations (order parameters) better, although not perfectly so.
+Sodium and calcium binding to membranes were evaluated against headgroup order parameter changes upon addition of NaCl or CaCl$_2$ in experiments and also compared with the ECClipids model with implicitly included electronic polarizability that currently gives the most accurate representation of ion binding against experiments.
+%The non-polarizable CHARMM36, when used with the NBFix ion model designed to provide more accurate binding, also rivals its polarizable counterparts. 
+%The polarizable AMOEBA and 
+CHARMM-drude2023 model exhibits a major improvement to non-polarizable CHARMM and CHARMM-Drude2017 parameters as it correctly reproduces the stronger calcium binding with respect to sodium that was not present in older parameters.
+This difference between sodium and calcium binding is also present in simulations with AMOEBA parameters.
+However, calcium
+%have similar ion binding levels to the ECC model, but they exhibit variation in 4
+binding depth, affinity, and the consequent structural response of the lipids do not exactly align with experiments or ECClipids results neither in CHARMM-Drude2023 nor AMOEBA simulations.
+The incorrect response to ion binding likely connects to the other discussed inaccuracies related to lipid conformational ensembles, dynamics and membrane order.
+%structure and diversity in the conformational ensemble, as well as slow dynamics that lead to poor convergence of simulations.
 
-In regard to conformational dynamics, the AMOEBA force field produces results, specifically in the headgroup region of PC lipids, that are either on par with or better than the best non-polarizable force fields.
-However, the CHARMM-Drude models exhibit an extremely slow dynamics, which can lead to deficient sampling and a substantial increase in required simulation time when using this model.
+ In summary, the full potential and promise of explicitly polarizable lipid force fields to improve the description of polar membrane regions have not yet been fully realized.
+However, it seems likely that this is not due to implementation of polarizability, but rather arises from compatibility of other parameters, suchs as dihedral and Lennart-Jones, with polarizable electrons.
+Due to long history in developing these parameters for non-polarizable force fields, it is not surprising that they
+ %benefit from a longer history of intense scrutiny and
+ out-perform their polarizable counterparts in many respects, as we here show for the case of certain phospholipid bilayers. On the other hand, improvements  
+from CHARMM-Drude2017 to CHARMM-Drude2023 demonstrate the potential for tuning these parameters to improve polarizable force fields.
+Such endeavors are expected to substantially ease with emerging automized methods for parameter development~\cite{Antila2022,yu2023drude}.
 
-In general, the ECC model gives the most accurate representation of ion binding, even though it lacks an explicit description of polarization.
-The non-polarizable CHARMM36, when used with the NBFix ion model designed to provide more accurate binding, also rivals its polarizable counterparts.
-The polarizable AMOEBA and Drude-2023 models have similar ion binding levels to the ECC model, but they exhibit variation in binding depth, and the consequent structural response of the lipids do not align with experiments.
-The incorrect response to ion binding likely connects to the the other inaccuracies described above, including the misrepresentation of overall membrane structure and diversity in the conformational ensemble, as well as slow dynamics that lead to poor convergence of simulations.
-
- In summary, the full potential and promise of explicitly polarizable force fields have still not been fully realized.
-Non-polarizable force fields benefit from a longer history of intense scrutiny and out-perform their polarizable counterparts in many respects, as we here show for the case of certain phospholipid bilayers.
-Considering the complexity and cost of using these models, it is crucial to parameterize polarizability with more thorough consideration to the experimental benchmarks.
-Doing so holds great potential for improvement, as demonstrated by the refinement of the CHARMM-Drude2017 model to the most recent CHARMM-Drude2023 version.
+ %Doing so holds great potential for improvement, as demonstrated by the refinement of the CHARMM-Drude2017 model to the most recent CHARMM-Drude2023 version.
 
@@
 \begin{comment}
-
+    4
 \subsubsection{Building a polarizable force field}
@@
 
-Considering the increased interest in polarizable force fields in membrane simulations, we mention here some hurdles in their practical usage.
-Firstly, the increased level of detail obtained by explicit treatment of electronic polarizability always comes with increasing computational cost.
+While non-polarizable MD simulations of membranes can be nowadays routinely performed with several simulation packages and force fields, polarizable simulations bear many practical complications.
+%Secondly, there are compability issues with the most used MD engines and the polarizable models and that support might be version dependent. 
+Regarding currently popular MD simulation packages for membrane simulations, OpenMM supports both AMOEBA and Drude force fields, NAMD can only run the Drude force field whereas, GROMACS
+%---the most popular MD engine within the bilayer simulation field---
+only has limited support for the Drude polarizable force field via an unofficial Git-branch~\cite{drudegithub}. %and no support for AMOEBA. 
+TINKER is widely used with AMOEBA, but it does not support semi-isotropic pressure coupling, which %decouples the pressure tensor in $xy$ and $z$-direction and is 
+is required in membrane simulations~\cite{xxx}.
+%crucial for correctly modelling membrane systems~\cite{xxx}.
+%
+Therefore, we have selected to use OpenMM in all simulations in this work.
+%In general, easy access to MD engines, as well as adaptability of common analysis tools and libraries for polarizable force fields, will be crucial in facilitating more wide-spread utilization of these models.
+CHARMM-GUI can generate the topology and input parameters for the Drude force field, which greatly simplifies the employment of this model in particular~\cite{kognole2022charmm}.
+Standard protocols for AMOEBA lipid parameters are not available, but some parameters are available online~\cite{amoebagithub,klesse2020induced}.
+
+%Considering the increased interest in polarizable force fields in membrane simulations, we mention here some hurdles in their practical usage. Firstly, 
+Another practical issue to be considered in polarizable membrane simulations is the increased computational cost arising from increased level of details obtained by explicit treatment of electronic polarizability.
 For the Drude-based models, this slowdown occurs both because the addition of the Drude particles increases the number of interaction pairs and the employed extended dual-Langevin thermostat requires a shorter 1~fs integration time step, compared to the 2~fs typically used for non-polarizable membrane simulations.
@@
 
-Secondly, there are compability issues with the most used MD engines and the polarizable models and that support might be version dependent.
-Currently, only OpenMM supports both AMOEBA and Drude force fields.
-NAMD can only run the Drude force field whereas TINKER is only compatible with AMOEBA and does not support semi-isotropic pressure coupling, which decouples the pressure tensor in $xy$ and $z$-direction and is crucial for correctly modelling membrane systems~\cite{xxx}.
-GROMACS---the most popular MD engine within the bilayer simulation field---only has limited support for the Drude polarizable force field via an unofficial Git-branch~\cite{drudegithub} and no support for AMOEBA.
-
-In general, easy access to MD engines, as well as adaptability of common analysis tools and libraries for polarizable force fields, will be crucial in facilitating more wide-spread utilization of these models.
-Fortunately, CHARMM-GUI now can generate the topology and input parameters for the Drude force field, which greatly simplifies the employment of this model in particular~\cite{kognole2022charmm}.
+All simulations performed in this work are listed in Table~\ref{table:sim_details} with links to the openly available trajectory data.
+Data not mentioned here was obtained by analysing pre-existing trajectories from the NMRlipids databank.
 
 \subsection{Simulations with CHARMM-Drude parameters}
-
-Table~\ref{table:sim_details} for the full list of simulations conducted for this study and links to the openly available trajectories.
-Data not mentioned here was obtained by analysing pre-existing trajectories from the NMRlipids databank.
 
@@
 
-CHARMM-Drude 2023 force field parameters are currently not integrated into CHARMM-GUI.
-Therefore, the simulation setups with NaCl and CaCl$_{2}$ have been generated following the instructions in the original CHARMM-Drude 2023 paper~\cite{yu2023drude} using CHARMM~\cite{brooks2009charmm} and the last frame of 200~ns long CHARMM36 simulations.
-CHARMM-Drude 2023 simulations without any ions have been directly obtained from Zenodo~\cite{richard_m_venable_2023_7872447, richard_m_venable_2023_7871949}.
+CHARMM-Drude2023 force field parameters are currently not integrated into CHARMM-GUI.
+Therefore, the simulation setups with NaCl and CaCl$_{2}$ have been generated following the instructions in the original CHARMM-Drude2023 paper~\cite{yu2023drude} using CHARMM program~\cite{brooks2009charmm} and the last frame of 200~ns long CHARMM36 simulations (same as used for CHARMM-Drude2017).
+CHARMM-Drude2023 simulations without any ions have been directly obtained from Zenodo~\cite{richard_m_venable_2023_7872447, richard_m_venable_2023_7871949}.
 
@@
 Bonds containing the hydrogens were constrained.
-For CHARMM-Drude 2017, particle mesh Ewald was used to compute the Coulomb interactions.
-van der Waals interactions have been cut to 0 between 1.0~nm and 1.2~nm using a switching function.
-For the CHARMM-Drude 2023, Lennard-Jones particle-mesh Ewald (LJ-PME) method was used to compute the long-range dispersions~\cite{wennberg2013lennard}.
+For CHARMM-Drude2017, particle mesh Ewald was used to compute the Coulomb interactions.
+Van der Waals interactions have been cut to 0 between 1.0~nm and 1.2~nm using a switching function.
+For the CHARMM-Drude2023, Lennard-Jones particle-mesh Ewald (LJ-PME) method was used to compute the long-range dispersions~\cite{wennberg2013lennard}.
 Simulation frames have been saved in every 10~ps.

### 2023-12-01 SI.tex

--- 2023-09-11 manuscriptPOL.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass[journal=jacsat,manuscript=article,layout=singlecolumn]{achemso}
+\documentclass[journal=jctcce,manuscript=article,layout=singlecolumn]{achemso}
 
@@
 
-\author{Hanne Antila}
+\author{Hanne S. Antila}
+\affiliation{Department of Theory and Bio-Systems, Max Planck Institute of Colloids and Interfaces, 14476 Potsdam, Germany}
+\alsoaffiliation{Department of Biomedicine, University of Bergen, 5020 Bergen, Norway}
+\email{hanne.antila@uib.no}
+\author{Sneha Dixit}
+\affiliation{Department of Theory and Bio-Systems, Max Planck Institute of Colloids and Interfaces, 14476 Potsdam, Germany}
 \author{Batuhan Kav}
@@
 \author{Jesper J. Madsen}
+\affiliation{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida 33612, United States of America}
+\alsoaffiliation{Center for Global Health and Infectious Diseases Research, Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida 33612, United States of America}
 \author{Markus S. Miettinen}
 \author{O. H. Samuli Ollila}
+\affiliation{VTT Technical Research Centre of Finland, Espoo, Finland}
+\alsoaffiliation{Institute of Biotechnology, University of Helsinki, Helsinki, Finland}
 
@@
 %\title{NMRLipids VI: Polarizable Force Fields}
-\title{Do explicitly polarizable force fields give better description of a biomembrane?}
+%\title{Do explicitly polarizable force fields give better description of biomembranes?}
 %\title{Comparison of polarizable lipid models to high-resolution experimental data reveals room for improvement}
 %\title{Benchmarking polarizable lipid force field}
+\title{Supplementary Information: Evaluation of polarizable biomembrane simulations against experiments}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The "tocentry" environment can be used to create an entry for the
-%% graphical table of contents. It is given here as some journals
-%% require that it is printed as part of the abstract page. It will
-%% be automatically moved as appropriate.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\begin{tocentry}
-
-%\end{tocentry}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The abstract environment will automatically gobble the contents
-%% if an abstract is not used by the target journal.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{abstract}
-
-	For the initial project description, please refer to \\ https://github.com/batukav/NMRlipidsVIpolarizableFFs/blob/master/Manuscript/manuscript.pdf
-
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Start the main part of the manuscript here.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-
-The dielectric environment in and around a lipid bilayer varies dramatically as you move from the water phase through the interfacial headgroup region to the membrane interior.
-The water phase highly polar, while the interfacial region contains dipolar headgroups, and the membrane interior is predominantly hydrophobic and nonpolar due to the aliphatic lipid tails.
-
-In conventional molecular dynamics (MD) simulations, electrostatic interactions are described by assigning static point charges to the atoms and molecules at the beginning of the simulation based on the force field definitions.
-Dynamic effects arising from molecular polarizability are traditionally not explicitly included and are only considered in an averaged fashion in the force field parametrization process when fitting to macroscopic observables.
-Significant efforts have been dedicated to introduce explicit polarizability into MD simulations, in the hopes of reaching a more accurate representation of the system~\cite{Thole1981,ando2001stable,Grossfield2003,
-lamoreux2003,Antila2013 ,Lemkul2016, baker2015polarizable,jing2019polarizable }.
-For bilayer membrane simulation, including polarizability in the lipids is expected to provide more accurate descriptions, especially in situations pertaining to membrane binding processes, translocation of charged biomolecules across the membrane, and the behavior of molecules residing within membranes, such as membrane proteins~\cite{baker2015polarizable,Lemkul2016, li2017drude,chu2018anionicpolarizable,Lynch21,Chen2021,Melcr:2018a, melcr2019improved,nencini22}.
-%Indeed, even implicit inclusion of polarizability appears to provide a better description of cation binding to membranes~\cite{Melcr:2018a, melcr2019improved}. 
-The main lipid force fields with explicit electronic polarization are CHARMM-Drude~\cite{li2017drude, yu2023drude}, AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable}, and CHARMM-Fluctuating Charge (FQ)~\cite{lucas2012charge}.
-The underlying treatment of polarizability differ among these models in the following way:  1) the classical Drude oscillator (CHARMM-Drude) models polarization by separating a core and a shell charge with a spring which stretches according to the envinronment giving the site a dipole moment~\cite{Lemkul2016},
-2) the induced point dipole/multipole (AMOEBA) uses polarizable point dipoles placed on chosen sites of the molecule~\cite{ponder2010current},
-and 3) the electronegativity equilization (FQ) employs atomic charges that are no longer constant but can change according to the electronegativities of the molecule atoms and the electric fields from the molecular enviroment~\cite{patel2004charmm}.
-All of these approaches result in an increasing computational cost, e.g, by introducing new types of interactions, more interactions sites, or by requiring a shorter time step in the simulations.
-As an alternative approach, electronic continuum correction (ECC) has been proposed as a computational efficient method to implicitly include polarizability by scaling the atom partial charges~\cite{Melcr:2018a,dijon20,Antila2021}.
-
-%Hence, there certainly is room for improvement: efforts from the NMRlipids open science community have shown that the current non-polarizable lipid models cannot correctly reproduce the conformational ensemble~\cite{Botan2015,Antila2019}, conformational dynamics~\cite{Antila2021}, or ion biding~\cite{Catte2016} within MD simulations.
-
-Our previous efforts in benchmarking state-of-the-art non-polarizable force fields have demonstrated that the quality of predictions for important membrane properties greatly vary between different force fields, particularly for lipid headgroup conformational ensembles and ion binding affinities~\cite{Botan2015,Catte2016,Antila2019,bacle21,Antila2021,Antila2022,Databank}.
-This emphasizes the critical importance of rigorous comparative quality evaluations of simulations performed using different force fields, which helps to estimate the reliability of simulation results and select the best models for specific applications.
-This is particularly important when employing polarizable force fields, as they offer the potential to capture a broader range of physical phenomena pertinent to enhancing our understanding of in particular polar membrane regions, at an increased computational cost.
-Unfortunately, systematic benchmark studies assessing the quality of polarizable force fields are,
-%is in great need. The few published studies reviewing the state of polarizable force fields for biomolecule simulations did not go beyond citing results from the existing studies
-to our knowledge, entirely missing from the field~\cite{inakollu2020polarisable,jing2019polarizable,baker2015polarizable}.
-Previous simulations using polarizable lipid force fields center around protein systems, with experimental validation (if performed at all) fovcusing on the properties of the proteins~\cite{jing2019polarizable, Lynch21}.
-Assessing the realism of the lipid models against experimental evidence from biophysical measurements is crucial since these membrane properties are foundational to more complicated applications, such as studies of transmembrane proteins.
-For example, it appears to be important that a wide variety of conformations is available for lipid headgroups in the model for realistic protein-lipid binding~\cite{bacle21}.
-
-Here we set out to assess the quality of available polarizable force fields, on par with previous evaluations of non-polarizable force fields, using the resources given by the NMRLipids open collaboration framework (https://databank.nmrlipids.fi/).
-Specifically, we evaluated the quality of POPC, DOPC and POPE lipid bilayer simulations with CHARMM-Drude~\cite{li2017drude,yu2023drude} and AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable} parameters against experimental NMR and small angle X-ray scattering (SAXS) data.
-These two force fields are increasingly utilized in biomolecular simulations and have lipid parameters available for the selected systems with corresponding experimental data available from the NMRlipids databank~\cite{Databank}.
-Following the quality metrics defined in the NMRlipids project, structural properties and cation binding were evaluated against NMR order parameters and SAXS form factors~\cite{Botan2015,ollila16,Catte2016,Databank}, while lipid conformational dynamics was evaluated in comparison to data from NMR spin relaxation rate experiments~\cite{ferreira15,Antila2021}.
-%Small angle scattering (SAXS) data is used to estimate the description of electron density profile of the membrane. 
-Furthermore, we compare the results to those obtained from the best performing non-polarizable simulations from the NMRlipids Databank\cite{Databank}.
-
- %n lipid modelling, the first two have surpassed the third in popularity.
-
-%All simulation data, analysis scripts, and results are publicly available in their corresponding links. No plausible nor implausible request is required to download, use, and disseminate the data.
-
-\section{Results and Discussion}
-
-\subsection{Evaluation of lipid bilayer structural properties}
-
-To evaluate the structural properties of lipid bilayers in simulations with polarizable force fields, we simulated POPC and POPE lipid bilayers with CHARMM-Drude2017~\cite{li2017drude} and CHARMM-Drude2023~\cite{yu2023drude} parameters, and POPE and DOPC bilayers with AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable} parameters.
-These systems were selected based on simultaneous availability of force field parameters, and experimental data~\cite{Databank}.
-After the data production, simulation trajectories were added into the NMRlipids Databank (https://github.com/NMRLipids/Databank/) and the NMRlipids quality metrics was used to evaluate each simulation against experiments~\cite{Databank}.
-This metric measures the quality of conformational ensemble of individual lipids in two aspects: first, it evaluates agreement with C-H bond order parameters (from NMR), and second, it examines consistency of membrane dimensions compared to small angle X-ray scattering (SAXS) data~\cite{Databank,ollila16}.
-%(measuring the electron density variation across the membrane). 
-The former is divided into three parts that separately describe the average quality of headgroup, glycerol backbone and acyl chain qualities as described in the NMRlipids databank~\cite{Databank}.
-While order parameters primarily reflect lipid conformations, acyl chain order parameters are also a good proxy for membrane packing since a smaller area per lipid tends to increase the magnitude of the order parameters~\cite{Databank}.
-%one for each of the carbonyl tails, and the quality of each is evaluated separately.
-
-\begin{figure}[t]
-    \centering
-    %\includegraphics{Figures/order_parameters.png}
-    \includegraphics[width=0.4\textwidth]{Figures/molecules.pdf}
-    \caption{Molecules and atom labelling used.}
-    \label{fig:molecules}
-
-\end{figure}
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{c c c c c c c}
-        Lipid & Force field & $P^{\mathrm{headgroup}}$ & $P^{sn-1}$ & $P^{sn-2}$ & $FF_{q}$ & APL \\
-        \hline
-        POPC & OPLS3e       & 0.76 & 0.87 & 0.85 & 0.15 & 66.5\\
-        POPC & CHARMM36     & 0.70 & 0.54 & 0.69 & 1.16 & 65.0\\
-        POPC & CHARMM-Drude & 0.52 & 0.29 & 0.53 & 1.06 & 62.5\\
-        POPC & CHARMM-Drude2023 & 0.63 & 0.60 & 0.57 & 0.96 & 64.5\\
-        \hline
-        POPE & GROMOS-CKP & 0.29 & 0.83 & 0.48 & 0.40 & 59.6\\
-        POPE & CHARMM36   & 0.54 & 0.52 & 0.27 & 1.30 & 57.2 \\
-        POPE & CHARMM-Drude & 0.06 & 0.53 & 0.27 & 0.80 & 56.6  \\
-        POPE & CHARMM-Drude2023 & 0.28 & 0.59 & 0.54 & 0.00 & 61.4  \\
-        POPE & AMOEBA & 0.21 & 0.10 & 0.23 & 3.80 & 66.9\\
-        \hline
-        DOPC & AMOEBA & 0.60 & 0.60 & 0.54 & - & 70.2\\
-    \end{tabular}
-    \caption{NMRlipids databank quality metrics~\cite{Databank} and area per lipids (APL) compared between the best simulations found from the NMRlipids databank (OPLS3e for POPC and GROMOS-CKP for POPE), simulations with CHARMM36 force field, and simulations with polarizable force fields.
-Order  parameter quality measures, $P^{\mathrm{headgroup}}$, $P^{sn-1}$, and $P^{sn-2}$, present the average probability of order parameters within a given segment to agree with experiments (larger probabilities mean higher quality).
-Form factor quality measure, $FF_{q}$, presents the difference in essential features between simulated and experimental form factors (smaller values indicate higher quality).
-Experimental estimations for areas per lipid are POPC: 64.3$\pm1$\AA$^{2}$\cite{kucerka2011}, DOPC: 67.5$\pm1$\AA$^{2}$\cite{kucerka2008}, and POPE: 56.7$\pm$3\AA$^{2}$\cite{Rickeard2020}.}
-    \label{tab:my_label}
-\end{table}
-
-\begin{figure}[!p]
-    \centering
-    %\includegraphics{Figures/order_parameters.png}
-    \includegraphics[width=0.8\textwidth]{Figures/quality.pdf}
-    \caption{X-ray scattering form factors, and C-H bond order parameters for headgroup and glycerol backbone, and acyl chains (from top to bottom) compared between simulations and experiments using the NMRlipids databank.
-The atom labelling is as depicted in Fig.~\ref{fig:molecules}.
-Experimental data in this figure are originally reported in Refs.~\citenum{ferreira13,bacle21,Databank,NMRlipidsIII,kucerka2015}.
-For CHARMM-Drude2023 simulations we selected representative replicas among three available simulations, for example for differences between replicas for POPC see Fig.~\ref{fig:order_parameters_replicas}.}
-    \label{fig:order_parameters}
-
-\end{figure}
-
-Direct comparisons between simulated and experimental data are visualized in Fig~\ref{fig:order_parameters} and the resulting quality metrics are shown in Table~\ref{tab:my_label}.
-Similarly to the non-polarizable CHARMM36 simulations with Gromacs, CHARMM-Drude simulation predicts slightly too packed membrane with excessively negative order parameters with respect to experiments and simulations with the highest quality in the NMRlipids databank (OPLS3e)~\cite{Databank}.
-However,
-%consistently outperform the polarizable force fields with respect to the NMR measurables for all parts of the molecule (higher $P$-values). For older CHARMM-Drude force field, 
-the quality of headgroup conformations in the CHARMM-Drude simulations is worse than in its non-polarizable counterpart.
-This is likely because headgroup and glycerol backbone parameters were optimized to reproduce absolute average values of the experimentally determined order parameters, instead of taking into account the order parameter forking (meaning measurably different order parameters for different C--H bonds at a single carbon atom) and order parameter sign~\cite{Antila2022}.
-A better description of the headgroup and glycerol backbone is provided by CHARMM36-Drude2023, yet its quality is still below that offered by the non-polarizable CHARMM36 parameters (see also discussion about conformational dynamics below).
-Also the quality of membrane packing and acyl chain order are improved in CHARMM36-Drude2023 compared to the earlier version but again, it did not perform at a quality comparable to that of the best available simulations.
-%also provides better description of the carbon tails. 
-%In fact, for POPE sn-2 tail the CHARMM-Drude2023 gives the best description of the models investigated here.
-
- AMOEBA simulations capture the headgroup and glycerol backbone order parameters reasonably well with the exception of g$_1$,
-%ll for both DOPC and POPE but vastly overestimates one of the g$_3$ values, resulting in 
-where forking of order parameters is unacceptably large.
-However, experimentally observed low order parameters at double bond region in both DOPC acyl chains and {\it sn}-2 chain of POPE are not captured.
-These low order parameters signal an important mechanism through which acyl chain double bonds affect membrane properties~\cite{ollila07} and are well reproduced in all state-of-the art MD simulation parameters~\cite{ollila16}.
-Furthermore, AMOEBA parameters substantially overestimate the area per lipid in POPE simulations, leading to too disordered acyl chains.
-Similar issues are evident also in membrane data presented in the recent publication for the AMOEBA cholesterol model~\cite{Li23chol}.
-%also clearly struggles to depict the tail conformations, in particular in the double bond region of DOPC (Fig.~\ref{fig:order_parameters}). 
-%parameters do not capture the  
-%In the tail regions, the CHARMM-Drude2023 appears to be the best out of the polarizable models.
-
-The unsatisfactory description of the lipid tail region and the area per lipid is further reflected in the inability of AMOEBA to capture the SAXS curves.
-Thus, we can conclude that our simulations with the AMOEBA parameters did not reproduce essential
-membrane properties.
-The higher $FFq$ values (Table~\ref{tab:my_label}) of the polarizable force fields in general indicate that they are not able confer consistency with the SAXS data as well as the best non-polarizable force fields (OPLS3e).
-That said, CHARMM-Drude2023 is on par with or better than its non-polarizable counterpart CHARMM36 and the best out of all the models here in capturing the POPE SAXS data, benefiting from the improvements made in the description of acyl chains.
-
-\subsection{Evaluation of conformational dynamics of lipids}
-C-H bond order parameters are sensitive to the conformational ensemble but the correspondence is not unique.
-Instead, they essentially describe the average of the distributions and do not contain information of the dynamics of the confomational sampling.
-Therefore, a simulation that reproduces the order parameters has an ensemble that is only potentially correct, and even the correct ensemble may not have the experimentally observed dynamics.
-To further elucidate this aspect in the case of polarizable force fields, we compared NMR $^{13}$C spin relaxation $R_{1}$ rates and C-H bond effective correlation times $\tau_{\mathrm{eff}}$ with experiments~\cite{ferreira15} and best non-polarizable simulations from our previous study~\cite{Antila2021} for simulations with PC headgroups and the glycerol backbone in Fig.~\ref{fig:correlation_times}.
-This focus was chosen based on the availability of both experimental data and polarizable simulations.
-$R_{1}$ rates measured with typical magnetic field strengths are sensitive to rotational dynamics of C-H bonds with the timescales around $\sim$0.1-1~ns, while effective correlation times respond to a larger range of dynamical processes up to $\sim$1000~ns~\cite{ferreira15}.
-The effective correlation time can be interpreted as an average measure for how fast molecular conformations sample the phase space that leads to the average order parameters.
-
-The effective correlation times in CHARMM-Drude2017 and CHARMM-Drude2023 simulations are approximately two  and one orders of magnitude slower, respectively, than the values suggested by experiments and best available simulations (Fig.~\ref{fig:correlation_times}).
-This indicates that not only are the simulations with this polarizable force field computationally costlier for equivalent lengths of trajectory, but one would also have to simulate longer to obtain converged results.
-The non-polarizable counterpart of the Drude models, CHARMM36, exhibits much more realistic, faster effective correlation times.
-Also the dynamics in the fast range ($R_{1}$ rates), are on average slightly more realistic in the non-polarizable CHARMM36 compared to both of the polarizable counterparts.
-Interestingly, CHARMM-Drude models have been reported to have slower water hydrogen bonding dynamics around amino acids compared to their non-polarizable counterpart,~\cite{Ngo2019} which might align with an overall slower dynamics of the model in addition to enhanced water binding.
-
-In contrast to the Drude-based models investigated here, the $R_1$ spin relaxation times and effective correlation times from DOPC simulations with AMOEBA force field are close to experimental data from POPC and on par with the best non-polarizable models (Slipids and CHARMM36).
-The small difference in acyl chain composition (DOPC vs.
-POPC) is not expected to strongly affect headgroup dynamics due to effective decoupling between hydrophilic and hydrophobic membrane regions~\cite{Antila2022rot, Klauda08}.
-%appears to be faulty in the fast range ($R_{1}$ rates) and, furthermore, $\tau_{\mathrm{eff}}$ values reveal vastly too slow overall dynamics.
-The updated CHARMM-Drude2023 does not significantly improve the dynamics at the fast range investigated here.
-The effective correlation times of the newer Drude model are improved but still an order of magnitude too slow.
-This indicates that not only are the simulations with this polarizable force field computationally costly but one would have to simulate longer to obtain converged results.
-The non-polarizable counterpart of the Drude model, CHARMM36, exhibits much more realistic, faster effective correlation times.
-%Interestingly, it has also been observed that CHARMM-Drude models have slower dynamic of water hydrogen bonding around amino acids compared to their non-polarizable counterpart~\cite{Ngo2019} which might align with an overall slower dynamics of the model in addition to enhanced water binding.
-%Note that the experimental data here is from POPC.
-However, chemically identical the headgroup region the dynamics is assumed to be practically identical in DOPC, due to the decoupling of the dynamics between the headgroup and the glycerol regions~\cite{Antila2022rot, Klauda08}.
-
-%We include the same measurables calculated from non-polarizable CHARMM36 and SLipids force fields---the best MD force fields regarding the dynamics~\cite{Antila2021}---as a point of comparison.
-
-\begin{figure}[!hbt]
-	\centering
-	\includegraphics{Figures/both_inkscp.pdf}
-	\caption{Effective correlation times (top) and R$_1$ rates (bottom) for the polarizable models and best performing non-polarizable force fields.
-Note, top panel axis is logarithmic to visualize the slow Drude dynamics.
-Experimental values are obtained from Ref.~\citenum{Antila2022rot}.
-All simulations used here were salt free.}
-	\label{fig:correlation_times}
-\end{figure}
-
-\subsection{Cation binding to membranes in polarizable simulations}
-
-Given the abundance of cations in biological systems, accurately capturing their interactions with membranes has the uttermost importance in MD simulations.
-A wealth of  experimental evidence shows that monovalent ions, such as sodium (except for lithium), exhibit very weak binding affinity of PC lipid bilayers, while multivalent ions like calcium bind more strongly~\cite{Catte2016}.
-However, simulations with non-polarizable force fields without any additional corrections systematically overestimate cation binding to lipid bilayers~\cite{Catte2016}.
-Implicit inclusion of electronic continuum correction (ECC) to both ion and lipid parameters can substantially improve the situation~\cite{Melcr:2018a,melcr2019improved,bacle21}, suggesting that electronic polarizability plays an important role in ion binding to membranes, which might further lead one to expect that that simulations with polarizable force fields will perform better in this aspect and more accurately describe ion binding to membranes.
-%Finally, we address the premise that adding polarizability to force fields improves the depiction of ion binding to biomembranes in simulations. 
-To test this notion, we evaluated sodium and calcium cation binding to membranes against the experimental NMR data using the "lipid electrometer," which involves quantifing the amount of ion binding to the membrane by monitoring the change in lipid headgroup order parameters in response to increasing salt concentration~\cite{Catte2016}.
-Changes in these order parameters induced by increasing NaCl and CaCl$_2$ are shown in Fig.~\ref{fig:popc_order_parameter_change} for simulations and experiments, whereas Fig.~\ref{fig:ion_density_profiles} shows the density profiles of ions with respect to the bilayer normal.
-Results from AMBER-based ECC-lipid model are shown for the reference that gives a good agreement with experiments for cation binding.
-
-\begin{figure}[!hbt]
-	\centering
-	\includegraphics{Figures/order_parameter_change.png}
-	\caption{The change in the lipid head group order parameters $\beta$ (top row) $\alpha$ (bottom row) upon increasing ion concentration with respect to the simulations without salt. }
-	\label{fig:popc_order_parameter_change}
-\end{figure}
-
-\begin{figure}[!hbt]
-    \centering
-    \includegraphics{Figures/ion_density_profiles_with_chloride.eps}
-    \caption{Density profiles along the membrane normal for the CHARMM36 (first row)~\cite{Catte2016}, Drude 2017 and 2023 POPC (second and third rows), AMOEBA DOPC (fourth row), and electronic continuum correction (ECC, last row)~\cite{Melcr:2018a}.
-Solid and dashed lines are for the systems with the NaCl and CaCl$_{2}$, respectively.
-Please note that only 350~mM and 450~mM CaCl$_{2}$ concentrations are shown for the Drude and AMOEBA force fields, respectively, while 1000~mM NaCl concentration is shown for all force fields.
-CHARMM36 NaCl simulations are at 950~mM.}
-    \label{fig:ion_density_profiles}
-            \todo{SAMULI: We should also show the ion density profiles along membrane normal and compare the ones with 1000 mM NaCl and 350 mM CaCl$_2$
-          (concentration in bulk water) to the best profiles in the literature (Fig. 5 in http://dx.doi.org/10.1021/acs.jpcb.7b12510).
-          This will show the actual difference in ion binding to the best performing models available.}
-          \todo{Batuhan: Data generated from http://dx.doi.org/10.1021/acs.jpcb.7b12510}
-\end{figure}
-
-The CHARMM-Drude2017 predicts similar calcium ion density profile to ECC-lipids, in good agreement with experiments.
-However, the sodium binding is equally strong in these simulations in contrast with the experimental evidence~\cite{Catte2016}.
-%, evidenced both by large change in the order parameters and the ion density profiles.
-Based on the change in the order parameters, the sodium binding is even stronger than the divalent calcium binding.
-In strong contrast, experimentally one sees no sodium binding and some binding of calcium ions.
-Excessive sodium binding in the CHARMM-Drude model has been noticed before for systems containing peptides or amino acids~\cite{Ngo2019, Kav2022} and deep-eutectic solvents~\cite{shayestehpour2022ion}.
-Excessive calcium interactions, in line with what we observe here for the lipids, have also been reported~\cite{Tan2022}.
-Furthermore, the response of headgroup order parameters to the bound ions is not in qualitative agreement with experiments for CHARMM-Drude2017 simulations according to Fig.~\ref{fig:popc_order_parameter_change}.
-Particularly, increasing $\alpha$ carbon order parameters and forking upon ion binding are not observed in experiments.
-This is in contrast with the results from non-polarizable simulations, where experimentally observed decrease of order parameters to more negative values upon ion binding were observed in all simulations, even though the binding affinity was often inaccurately predicted~\cite{Catte2016}.
-This qualitative discrepancy in CHARMM-Drude2017 simulations most likely results from incorrect lipid headgroup conformational ensemble (see previous sections), which leads to inaccuracies in the structural response of the ensemble to ion binding.
-
-In simulations with CHARMM-Drude2023 parameters, sodium ion binding is in line with ECC-lipid simulations when comparing density profiles,
-%force field: the change of the order parameters is diminished to be more in line with experiments and only the some accumulation of sodium ions near the phosphate oxygen is seen in the ion distributions.
-%However, this the ion layering is most prominently present in the original CHARMM36 force field, for for sodium and calcium ions. 
-but addition of sodium or calcium ions seems to induce forking of order parameters in contrast to experiments. 
-%at the $\beta$ carbon in particular. 
-This might indicate inaccurate structural response to ion binding, but poor convergence of the simulations owing to the slow conformational dynamics detected above cannot be ruled out in this case or in the case of the older 2017 version.
-%, which would also not be fully reflected in the calculation of error estimates for the order parameters. 
-Based on density profiles, calcium binding in CHARMM-Drude2023 simulations is in line with the ECC-lipid model, although binding affinity is slightly larger, 
-%also appears to be better represented in the updated model. The 
-ions penetrate slightly
-%reveal that binding of calcium has moved 
-deeper into the bilayer, and there is a stronger formation of ionic double layer by subsequent attraction of chloride into the bilayer. 
-%As opposed to the CHARMM36-Drude model, the order parameters of the AMOEBA model are almost completely unaffected by the ion addition, both for NaCl and CaCl$_{2}$.
-The ion density profiles confirm that very little ions penetrate the membrane region where the lipid phosphates reside.
-While this is in line with experiments regarding sodium ions, there should be some binding of calcium ions to the membrane.
-In AMOEBA simulations, sodium binds weakly and does not affect order parameters, consistent with experiments and ECC-lipid simulations.
-However, order parameters do not change upon calcium binding in accordance with experiments and other simulations.
-This can be explained by the binding position of calcium, which is outside phosphate density peak in AMOEBA simulations, whereas calcium penetrates to phosphate region or deeper in other simulations and thereby reorients the headgroup dipole, giving rise to changes in order parameters~\cite{Catte2016}.
-
-%Again, the change in the order parameters indicate the non-polarizable CHARMM36 depicts sodium binding the best out of the models compared here, being slightly more realistic than its most recent polarizable counterpart, CHARMM-Drude2023.
-Regarding calcium binding, CHARMM36 captures the $\Delta^{\alpha}_{CH}$ about as well as the CHARMM-Drude2023, but the change in the $\beta$ carbon is better represented in the CHARMM-Drude2023 and AMOEBA models.
-This suggest that some accuracy might have been gained in modelling the divalent ion binding with polarizable models but the advantage is not clear.
-No force field, polarizable or non-polarizable, is able to model both monovalent and divalent ion binding flawlessly.
-This is of note, as polarizable force fields are often assumed superior and concequently a popupar choice in modelling, eg, ion channels~\cite{sun2017 jing2019polarizable, klesse2020induced}
-\todo{I think there is a mistake in the C36 data, firsly, the concentration should probably be 950mM instead of 1000mM and secondly the monovalent anion and cation curves look identical} ~\todo{batuhan: I fxed the anion and cation curves.
--> @batuhan will plot the c36 data in a different color for the 950mm case.s}.
-
-%\begin{figure}[!hbt]
-%	\centering
-%	\includegraphics{dihedral_distributions_for_all.png}
-%	\caption{The distributions of the torsion angles for the head group atoms}
-%	\label{fig:dihedral}
-%        \todo{SAMULI: I am not sure if we need this figure at all.
-%          Or were you planning to discuss why order parameter response to ions is qualitatively incorrect in Drude PC?
-%          If yes, then maybe remove POPE from this figure.
-%        }
-%\end{figure}
-
-%\clearpage
-
-%\clearpage
-
-In conclusion, incorporating explicit polarizability as done in CHARMM-Drude or AMOEBA force fields do not necessarily lead to improved description of cation binding to phospolipid membranes.
-However, it is difficult to isolate the explicit influence of polarizability per se on ion binding based on the results presented here.
-What is clear, however, is that the headgroup conformational ensemble, which presumably plays a role in ion binding, is poorly described by the CHARMM-Drude parameters.
-In addition, the AMOEBA parameters struggle to capture the basic membrane properties that are reproduced by all current state-of-the-art MD simulation models.
-
-\subsection{Discussion}
-
-\textcolor{red}{For the ion binding discussion: drude calcium overbinds to proteins. this is known here https://pubs.acs.org/doi/full/10.1021/acsphyschemau.1c00039
-drude sodium overbinds to proteins, as shown here https://pubs.rsc.org/en/content/articlepdf/2022/ra/d2ra01478e and https://onlinelibrary.wiley.com/doi/abs/10.1002/adts.201800106.
-Venable also cautioned us against using any ion parameters with the drude force fields.
-This is important as 1- polarizable force fields are supposed to be better for ion parameters and 2- they are generally used to study the ion channels.}
-
-\subsection{Conclusions}
-We have rigorously compared simulation output of model lipid bilayers using polarizable lipid force fields with high-resolution experimental data as well as results from best-performing non-polarizable force fields.
-From this exercise, it is evident that current polarizable force fields for lipids fall short in important aspects.
-In most cases, the best non-polarizable models exhibits superior performance and accuracy across all the categories investigated in this study.
-
-Evaluation order parameters and scattering form factors reveals that the AMOEBA force field has difficulties in producing membranes with reasonable physical properties.
-In constrast, the CHARMM-Drude2023 model has improved performance in this regard, notably surpassing its predecessor, CHARMM-Drude2017.
-Nevertheless, top-performing non-polarizable models generally reproduce experimental conformations (order parameters) better, although not perfectly so.
-
-In regard to conformational dynamics, the AMOEBA force field produces results, specifically in the headgroup region of PC lipids, that are either on par with or better than the best non-polarizable force fields.
-However, the CHARMM-Drude models exhibit an extremely slow dynamics, which can lead to deficient sampling and a substantial increase in simulation simulation time when using this model.
-
-In general, the ECC model exhibits thet most accurate representation of ion binding, even though it lacks an explicit description of polarization.
-The non-polarizable CHARMM36, when used with the NBFix ion model designed to provide more accurate binding, does XXX.
-The polarizable AMOEBA and Drude-2023 models show similar ion binding levels to the ECC model, but they exhibit variation in binding depth and the consequent structural response of the lipids do not align with experiments.
-The incorrect response to ion binding likely connects to the the other inaccuracies described above, including the misrepresentation of overall membrane structure and diversity in the conformational ensemble, as well as slow dynamics that lead to poor convergence of simulations.
-
- In summary, the full potential and promise of explicitly polarizable force fields have still not been fully realized.
-Non-polarizable force fields benefit from a longer history of intense scrutiny and out-perform their polarizable counterparts in many respects, as we here show for the case of certain phospholipid bilayers.
-Considering the complexity and cost of using these models, it is crucial to parameterize polarizability with more thorough consideration to the experimental benchmarks.
-Doing so holds great potential for improvement, as demonstrated by the refinement of the CHARMM-Drude2017 model to the most recent CHARMM-Drude2023 version.
-
-\section{Methods}
-%\subsection{Practical pitfalls of polarizable biomembrane simulations}
-
-\begin{comment}
-
-\subsubsection{Building a polarizable force field}
-
-The atomic point charges describing the electrostatics in non-polarizable force fields are typically parameterized by fitting to the electrostatic potential energy surface around one or few conformers obtained from ab initio calculations.
-One then proceeds to build on top of that the parameters for Lennard-Jones interactions and the dihedrals using a mixture of experimental (such as heats of vaporizations) and quantum mechanical information.
-For a polarizable force fields, this process has to be redone as in the first step the polarizable contribution has to be separated from the "static" one.
-Furthermore, for the polarization models considered here, an additional cutoff/damping scheme needs to be introduced to prevent polarization catastrophe~\cite{Thole1981}---a phenomenon where close-by dipoles feed of each others and their dipole moments (and energy) grows into infinity.
-
-Both polarizable lipid force fields considered here use a bottom up approach where parameters are first obtained for smaller molecules or molecular fragments from which the lipid models are then built.
-In the CHARMM-Drude model, the polarizabilities, the Thole damping factors used to prevent polarization catastrophe, and the partial charges where determined by fitting to a discretized QM potential energy surface around the molecule.
-More specifically, the perturbations of the potential energy surface in the presence of a small charge where observed to separate the dynamic polarizability contribution.
-They then proceeded to use enthalpies of vaporation along with QM calculations of molecular interaction energies to obtain the Lennard-Jones parameters, and on top of that dihedral potentials were introduced to reproduce the QM potential energy profiles along a dihedral rotation.
-Finally, the dihedrals of the lipid head groups where refined targetting the NMR order parameters.
-
-In the AMOEBA force field, a set of atomic multipoles were first obtained based on the QM electron densities using Stone's distributed multipole analysis~\cite{Stone1981}.
-The atoms were then assigned polarizabilities according to the AMOEBA atom types and induced dipole contribution was self consistently subtracted from the permanent dipoles~\cite{shi2013proteinamoeba}.
-The charges (monopoles) were then kept constant while the multipoles were refined against the electrostatic potential energy surface around few conformers of the molecular fragments.
-%Multipoles are allowed to induce only dipoles at polarizable sites outside their immediate environment (polarization group).
-After electrostatic parameters were set, the Lennard-Jones parameters were taken from previous AMOEBA parameterizations of smaller molecules containing same functional groups~\cite{Ren2011polorganic,shi2011hydration}, and then refined against molecule - water (gas phase) interaction energies.
-After settling all the nonbonded parameters, dihedral potentials were parameterised based on the QM conformational energy profiles.
-\end{comment}
-\begin{comment}
-- polarization contribution has to be separated from point charge electrostatics in the parametrization process.
-- for the models considered here (Drude, IDP) one has to prevent the polarization catastrophe somehow (Thole damping parameter)
-- Brief explanation how the two FFs studied here were built.
-
-Notes on CHARMM-Drude parametrization:
-- bottom up approach. Permantent charge, polarizability and screening factors from QM. Polarizability decoupled from permanent by fitting to changes in PES? 
-- Non-bonded electrostatics parameters calculated based on QM.
-- LJ from enthalpies of vaporization together .with QM data
-- Dihedrals from QM scans for model compounds, then refined to according to the order parameters of the headgroup using re-weighted MD and an genetic algorithm.
-
-Notes on AMOEBA parametrization (this is NOT well explained in the paper)
--bottom up approach
--Stone distributed multipole analysis to get the multipoles from the QM electron densities. Divided into permanent and induced.
--Then iterative fitting to QM-PES for electrostatics
--WdW from 1) optimization to hydration energy calculations and 2) QM fragment-water gas phase calculations.
--dihedrals optimized to QM potentials
--no mention of the screening for polarization? Do they use some AMOEBA value
-\end{comment}
-\subsection{Using a polarizable force field for membrane simulations}
-
-Considering the increased interest in polarizable force fields in membrane simulations, we mention here some hurdles in their practical usage.
-Firstly, the increased level of detail obtained by explicit treatment of electronic polarizability always comes with increasing computational cost.
-For the Drude-based models, this slowdown occurs both because the addition of the Drude particles increases the number of interaction pairs and the employed extended dual-Langevin thermostat requires a shorter 1~fs integration time step, compared to the 2~fs typically used for non-polarizable membrane simulations.
-The AMOEBA force field can use a multi time-step integration algorithm where the non-electronic interactions are iterated with a 2~fs time step whereas more computationally unstable polarization terms are iterated with a smaller time step.
-However, the multi-time step scheme  only partly mitigates the computational cost: one typically experiences almost a ten-fold decrease in the simulation speed while using the polarizable force fields.
-
-Secondly, there are compability issues with the most used MD engines and the polarizable models and that support might be version dependent.
-Currently, only OpenMM supports both AMOEBA and Drude force fields.
-NAMD can only run the Drude force field whereas TINKER is only compatible with AMOEBA and does not support semi-isotropic pressure coupling, which decouples the pressure tensor in $xy$ and $z$-direction and is crucial for correctly modelling membrane systems~\cite{xxx}.
-GROMACS---the most popular MD engine within the bilayer simulation field---only has limited support for the Drude polarizable force field via an unofficial Git-branch~\cite{drudegithub} and no support for AMOEBA.
-
-In general, easy access to MD engines, as well as adaptability of common analysis tools and libraries for polarizable force fields, will be crucial in facilitating more wide-spread utilization of these models.
-Fortunately, CHARMM-GUI now can generate the topology and input parameters for the Drude force field, which greatly simplifies the employment of this model in particular~\cite{kognole2022charmm}.
-
-\subsection{Simulations with CHARMM-Drude parameters}
-
-Table~\ref{table:sim_details} for the full list of simulations conducted for this study and links to the openly available trajectories.
-Data not mentioned here was obtained by analysis pre-existing trajectories in the NMRlipids databank.
-
-CHARMM-Drude2017 simulations were performed with OpenMM 7.5.0~\cite{eastman2017openmm} using parameters extracted with \textit{Membrane Builder}~\cite{wu2014charmm,jo2009charmm,jo2007automated,lee2018charmm} and \textit{Drude Prepper}~\cite{kognole2022charmm} from CHARMM-GUI~\cite{jo2008charmm,lee2016charmm}.
-Before starting the simulations, membrane structures were equilibrated for 200~ns using the non-polarizable CHARMM36 force field~\cite{klauda2010update}, and the last frames of these simulations were used to generate the starting structures for the polarizable force field simulations.
-
-CHARMM-Drude 2023 force field parameters are currently not integrated into CHARMM-GUI.
-Therefore, the simulation setups with NaCl and CaCl$_{2}$ have been generated following the instructions in the original CHARMM-Drude 2023 paper~\cite{yu2023drude} using CHARMM~\cite{brooks2009charmm} and the last frame of 200~ns long CHARMM36 simulations.
-CHARMM-Drude 2023 simulations without any ions have been directly obtained from Zenodo~\cite{richard_m_venable_2023_7872447, richard_m_venable_2023_7871949}.
-
-A dual Langevin thermostat was employed to keep the Drude particles and the rest of the system at 1.0~K and 303~K, respectively.
-A Drude-hardwall of 0.02~nm was utilized to keep the Drude particles close to their parent atoms.
-Semi-isotropic Monte Carlo barostat was used to couple pressure to 1~bar independently in membrane plane and normal directions.
-%while the pressure fluctuations along the $xy$-plane were coupled.
-A constant surface tension of 0.0~dyne/cm was applied throughout the simulations.
-Bonds containing the hydrogens were constrained.
-For CHARMM-Drude 2017, particle mesh Ewald was used to compute the Coulomb interactions.
-van der Waals interactions have been cut to 0 between 1.0~nm and 1.2~nm using a switching function.
-For the CHARMM-Drude 2023, Lennard-Jones particle-mesh Ewald (LJ-PME) method was used to compute the long-range dispersions~\cite{wennberg2013lennard}.
-Simulation frames have been saved in every 10~ps.
-
-\subsection{Simulations with AMOEBA parameters}
-
-AMOEBA force field parameters for OpenMM were obtained from GitHub~\cite{amoebagithub,klesse2020induced}.
-All AMOEBA simulations have been run on OpenMM 7.1.1~\cite{eastman2017openmm}.
-The same initial structures as the CHARMM-Drude simulations have been used to create the simulation setups.
-A multi-time step Langevin integrator was used to iterate the bonded and non-bonded interactions in every 0.5~fs and 2.0~fs, respectively.
-A non-bonded cutoff of 1.2~nm has been applied while semi-isotropic Monte Carlo barostat was used to couple pressure to 1~bar independently in membrane plane and normal directions.
-%keeping the pressure at 1~bar using a Monte Carlo barostat.
-Pressure fluctuations in the $xy$-plane has been coupled using a semi-isotropic pressure coupling and a constant surface tension at 0.0~dyn/cm has been applied throughout the simulations.
-Simulation frames have been saved in every 10~ps. Further simulations details can be found in the input files of the respective simulations (Table~\ref{table:sim_details})
-
-\subsection{Analysis of simulations}
- All simulations were first uploaded to the NMRlipids databank~\cite{Databank}. 
-Area per lipids, X-ray scattering form factors, and order parameters ($S_{CH}$) are automatically calculated by the NMRLipids Databank~\cite{Databank} and were extracted from therein.
-Quality evaluation metrics were quantified as detailed in the NMRlipids databank~\cite{Databank} with the exception that POPC simulations at 303\,K were coupled with the experimental data measured at 300\,K, while in the NMRlipids databank simulations are coupled to experiments with the maximum temperature difference of two degrees.
-Shortly, order parameter qualities, $P^{\mathrm{headgroup}}$, $P^{sn-1}$, and $P^{sn-2}$, presents an average probability for order parameters of a given segment to locate within experimental values, taking the error bars of both values into the account.
-Qualities of SAXS form factors, $FF_{q}$, are measured from the distance between first form factor minimum in simulated vs.
-experimetal curve to avoid the effects from the simulation size dependency of the SAXS data.
-%~\cite{Databank} one should focus on the extrema $q_{z}$[\AA$^{-1}$] when comparing the simulated and experimental data.
-In particular $FFq$ only measures the difference between the first minima of the simulated and experimental curves.
-Note that in quantifying the bilayer electron densities for calculation of SAXS curves  the NMRlipids databank analysis algorithm places electrons as point charges at the atom centers without considering redistribution of charge density due to the polarizability.
-Nevertheless, we expect that this approximation does not have significant effect on the resulting SAXS form factors.
-
-The $R_{1}$ rates and effective correlation times, along with the accompanying error estimates, are quantified from the trajectories as elaborated in Ref.~\citenum{Antila2021}.
-In-house python script available at XXX was used for this purpose.
-Source codes for all scripts and a detailed methodological explanation of the computed values are available in the referenced cited herein.
-
-\begin{comment}
-Order parameters are calculated using the script "OrderParameters.py" under\\ "https://github.com/batukav/NMRlipidsVIpolarizableFFs/tree/master/scripts". \\
-The order parameters are first calculated for each lipid as an average over time, and then averaged over the lipids.
-The error estimates are calculated as the standard error of the mean from the second average.
-Dihedral distributions are calculated using the script "calcDihedral.py" under\\ "https://github.com/batukav/NMRlipidsVIpolarizableFFs/tree/master/scripts" .\\
-The $R_{1}$ rates and effective correlation times, along with the accompanying error estimates, are quantified from the trajectories as elaborated in Ref.~\citenum{Antila2021}.
-In-house python script available at XXX was used for this purpose.
-\subsection{Simulation Details and data availability}
-This project is performed using the "NMRLipids Databank" format and all related files, except the trajectories, are available under\\ "https://github.com/batukav/NMRlipidsVIpolarizableFFs/"
-\end{comment}
-
-\newpage
-%\begin{adjustbox}{angle=90}
-\begin{table}[]
-\begin{small}
-\begin{tabular}{cccccccccc}
-	Lipid/ion                & force field  & Ion (M) & N$_{l}$ & N$_{w}$ & N$_{ion}$ & T (K) & t$_{sim}$ (ns) & t$_{analysis}$ (ns) & files \\ \hline
-\multirow{2}{*}{POPC}            & Drude2017 & 0      & 144       & 6400       & 0         & 303    & 500              & 400         &          \cite{kav_batuhan_2021_7607436}    \\ 
-                                 & Drude2023 & 0      & 72       & 2239       & 0         & 303      & 300              & 200         &          \cite{kav_batuhan_2023_7916287, richard_m_venable_2023_7871949}    \\ \hline
-\multirow{3}{*}{POPE}                              & Drude2017 & 0      & 144       & 6400       & 0         & 308.15    & 350              & 300         &          \cite{kav_batuhan_2021_7604627}    \\                                            & Drude2023 & 0       & 72        & 2304      &  0         & 303    & 300             & 200         &           \cite{richard_m_venable_2023_7872447,kav_batuhan_2023_7916494} \\
-                                                  & AMOEBA     & 0       & 72        & 2880      &  0        & 303     & 305.94          & 305.94      & \cite{kav_batuhan_2023_7622838} \\ \hline
-	\multirow{5}{*}{POPC:NaCl}        & Drude2017 & 0.350   & 144      & 6400     & 41         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_7586915}   \\
-				  & Drude2017 & 0.450   & 144      & 6400     & 51         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_7591753}   \\
-				  & Drude2017 & 0.650   & 144      & 6400     & 77         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_7596011}   \\
-				  & Drude2017 & 1.0     & 144      & 6400     & 115        & 303   & 500             & 400                  & \cite{kav_batuhan_2020_7600326}   \\ 
-                    & Drude2023 & 1.0      & 128       & 6400      & 41          & 303    & 223.97      &       220.14       & \cite{kav_batuhan_2023_8000095} \\     
-                    & Drude2023 & 1.0      & 128       & 6400      & 115          & 303    & 220.14      &       220.14       & \cite{kav_batuhan_2023_8000133} \\ \hline
-
-	\multirow{5}{*}{POPC:CaCl$_{2}$} & Drude2017 & 0.350   & 144      & 6400     & 41         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_7600827}   \\
-				  & Drude2017 & 0.450   & 144      & 6400     & 52         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_7605016}   \\
-				  & Drude2017 & 0.650   & 144      & 6400     & 76         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_7604040} \\
-				  & Drude2017 & 1.0   & 144      & 6400     & 114         & 303   & 500             & 400                  & \cite{kav_batuhan_2023_7658975}\\
-                    & Drude2023 & 0.350 & 128      & 6400     & 41          & 303   & 219.59          & 219.59               & \cite{kav_batuhan_2023_8000065}\\
-                    & Drude2023 & 0.790 & 128      & 6400     & 91          & 303   & 214.85          & 214.85               & \cite{kav_batuhan_2023_7992137} \\ \hline
-DOPC           & AMOEBA   & 0 & 72 & 2880 & 0 & 303 & 201.61 & 201.61 & \cite{kav_batuhan_2022_7604681} \\ \hline
-\multirow{2}{*}{DOPC:NaCl} & AMOEBA & 0.450 & 72 & 2880 & 17 & 303 & 218.41 & 218.41 & \cite{kav_batuhan_2022_7604711} \\
-                           & AMOEBA & 1.0 & 72 & 2880 &  35 & 303 & 201.61 & 201.61 & \cite{kav_batuhan_2023_7625844} \\ \hline
-\multirow{2}{*}{DOPC:CaCl$_{2}$} & AMOEBA  & 0.450 & 72 & 2880 & 16 & 303 & 218.41 & 218.41 & \cite{kav_batuhan_2022_7604842} \\
-                                 & AMOEBA & 1.0 & 72 & 2880  & 36 & 303 & 218.41 & 218.41 & \cite{kav_batuhan_2022_7604810}
-
-\end{tabular}
-\end{small}
-\caption{Systems simulated for this work.
-Column N$_l$ gives the number of lipids, N$_w$ gives the number of water molecules, T(K) denotes the temperature in kelvins.
-Simulated time is listed in column t$_{sim}$ and time used for analysis in t$_{analysis}$.
-Column "files" gives the reference to open access simulation data.}
-\label{table:sim_details}
-\end{table}
-\todo{check the simualation settings (especially the temperature}
-%\end{adjustbox}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The "Acknowledgement" section can be given in all manuscript
-%% classes.  This should be given within the "acknowledgement"
-%% environment, which will make the correct section or running title.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{acknowledgement}
-
-J.J.M. thanks Huiying Chu (from the Li lab) for technical discussions and Sameer Varma for useful suggestions.
-
-\end{acknowledgement}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The same is true for Supporting Information, which should use the
-%% suppinfo environment.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{suppinfo}
-\end{suppinfo}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The appropriate \bibliography command should be placed here.
-%% Notice that the class file automatically sets \bibliographystyle
-%% and also names the section correctly.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\bibliography{references.bib}
-
-\section{Supplementary figures}
-
+\renewcommand{\thefigure}{S\arabic{figure}}
+\setcounter{figure}{0} 
 \begin{figure}[!h]
@@
     \includegraphics[width=0.8\textwidth]{Figures/quality_Drude2023_replicas.pdf}
-    \caption{X-ray scattering form factors, and C-H bond order parameters for headgroup and glycerol backbone, and acyl chains (from top to bottom) compared between simulations and experiments using the NMRlipids databank.
-The atom labelling is as depicted in Fig.~\ref{fig:molecules}.}
+    \caption{Comparison of different replicas from CHARMM-Drude2023 simulations against experimental X-ray scattering form factors and C-H bond order parameters from the NMRlipids databank.}
     \label{fig:order_parameters_replicas}
@@
 \end{figure}
-
 \end{document}

### 2023-12-01 cover.tex

@@
 Please find enclosed our manuscript titled
-{XXX}, which we are submitting for your consideration to be published in \textit{XXX}.
+{XXX}, which we are submitting for your consideration to be published as an article in \textit{XXX}.
 
 In the manuscript, we answer the question "are polarizable molecular dynamics force fields better for modelling biomembranes (lipid bilayers) than their non-polarizable counterparts and in which ways".
-This has long thought to be the case as polarizable models have the potential for more accurate description of ionic interactions and the varying dielectric environment across a lipid bilayer.
+This has long thought to be the case as polarizable models have the potential for more accurate description of ionic interactions and the varying (dielectric) environment across a lipid bilayer.
 Systematic benchmarking of the models against both high-resolution experimental data and non-polarizable model has, however, been lacking.
 
-We seek to fill this gap by using the NMRlipids open collaboration framework and the related NMRlipids databank.
-We test the models againts experimental NMR order parameters (describing the conformational space of lipid molecules), small-angle X-ray scattering curves (depicting the fourier tranformation of the electron density across the membrane), NMR relaxation and correlation rates (describing the conformational dynamics), and the change of NMR order parameters when salt is added to the system, which indicate the strength of ion binding.
-We find that across all these test, the best non-polarizable models beat the polarizable lipid models tested: AMOEBA, CHARMM-Drude2017/CHARMM-Drude2023.
-Although the polarizable models give good description in some cases, such as the ability of AMOEBA to capture monovalent ion binding and lipid headgroup dynamics, they also have some serious flaws, like the faulty description of the lipid tails in AMOEBA and vastly too slow conformational dynamics of CHARMM-Drude models.
-It is particularly important to bring attention the last result, since it means one has to run the simulation longer to obtain good quality data.
+We seek to fill this gap by using the NMRlipids open collaboration framework and the related, novel NMRlipids databank.
+We test the models against experimental NMR (nuclear magnetic resonance) order parameters (describing the conformational space of lipid molecules), small-angle X-ray scattering curves (which measure the Fourier transformation of density of scatterers and thus capture more macroscopic membrane structure), NMR relaxation and correlation rates (describing the conformational dynamics), and the change of NMR order parameters when salt is added to the system, which indicate the strength of ion binding.
+We find that across all these test, the best non-polarizable models beat the polarizable lipid models tested: AMOEBA-based parametrization, CHARMM-Drude2017, and CHARMM-Drude2023.
+Although the polarizable models give good description in some cases, such as the ability of AMOEBA to capture monovalent ion binding and lipid headgroup dynamics, they also have some serious flaws, like the faulty description of the lipid tails in AMOEBA and the vastly too slow conformational dynamics of both CHARMM-Drude models.
+It is particularly important to bring attention the last result, since it means one has to run the simulation longer to obtain good quality data on top of the already-increased computational cost owing to the implicit polarizability.
 
-We believe the test performed in this paper will provide useful reference for anyone looking to use polarizable force field for membrane simulations.
-We hope it will also encourage the developers of these models to use the experimental data in the parametrization and/or carefully benchmach their models against it.
-The paper exemplifies the benefits of doing so by pointing out the improvement from the CHARMM-Drude2017 model to the 2023 version.
+We believe the test performed in this paper will provide an useful resource for anyone looking to use polarizable force field for membrane simulations.
+We hope it will also encourage the developers of these models to use the experimental data in the parametrization and carefully benchmark their models against it.
+The manuscript exemplifies the benefits of doing so by pointing out the improvement from the CHARMM-Drude2017 model to the 2023 version.
 

### 2023-12-01 manuscriptPOL.tex

@@
 %\title{NMRLipids VI: Polarizable Force Fields}
-\title{Do explicitly polarizable force fields give better description of biomembranes?}
+%\title{Do explicitly polarizable force fields give better description of biomembranes?}
 %\title{Comparison of polarizable lipid models to high-resolution experimental data reveals room for improvement}
 %\title{Benchmarking polarizable lipid force field}
+\title{Evaluating polarizable biomembrane simulations against experiments}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{abstract}
-With the increase of available computational capabilities, polarizable molecular dynamics force fields have gained popularity in modelling biomolecular systems.
-In the case of biomembranes in particular, they are expected to provide better description, eg., by capturing the varying dielectric environment across the lipid bilayer.
-Here, we use the framework established within the NMRlipids open science project and the related NMRlipids databank to test the commonly used polarizable lipid models, CHARMM-Drude and AMOEBA, against high fidelity experimental data and best non-polarizable models.
+%Classical molecular dynamics (MD) simulations are powerful tool to model membranes and other complex biomolecular systems with high spatiotemporal resolution.
+However, conventional MD simulations assume static point charges to the atoms, thereby neglecting the electronic polarizability effects.
+Because electronic polarization effects play key roles in many biomolecular systems,
+%. With the remarkable advances in available computational capabilities, 
+%polarizable molecular dynamics force fields are gaining popularity for more accurate modeling of biomolecular systems.
+For these advances, it is crucial to evaluate how accuracy of models against experiments evolves when computational cost and complexity of simulations are increasing.
+
+%Notably, in the context of biomembranes, these force fields  are expected to offer an improved description, eg., by more accurately capturing the dynamic dielectric environment across the lipid bilayer.
+Owing to the increase of available computational capabilities and the potential for providing more accurate description, polarizable molecular dynamics force fields are gaining popularity in modelling biomolecular systems.
+It is, however, crucial to evaluate how much precision is truly gained with the increased cost and complexity of the simulation.
+Here, we leverage the NMRlipids open collaboration and databank to assess the performance available polarizable lipid models, CHARMM-Drude and AMOEBA-based parameters, against high-fidelity experimental data and compare them to top-performing non-polarizable models.
+While some improvement in description of ion binding to membranes is observed in the most recent CHARMM-Drude parameters, and the conformational dynamics of AMOEBA-based parameters are excellent, best non-polarizable models tend to outperformed their polarizable counterparts for each property we explored.
+The identified shortcomings range from inaccuracies in describing the conformational space of lipids to excessively slow conformational dynamics.
+Our results provide valuable insights
+for further refinement of polarizable lipid force fields and for selecting the best simulation parameters for specific applications. 
 %We find that, for each property tested, the best non-polarizable models surpass the polarizable ones.
@@
 
-In conventional molecular dynamics (MD) simulations, electrostatic interactions are described by assigning static point charges to the atoms and molecules at the beginning of the simulation as defined in the inputted model---the force field.
-Dynamic effects arising from molecular polarizability are traditionally not explicitly included and only considered in an averaged fashion via force field parameterization process where parameters are obtained by fitting to macroscopic observables or to ab initio calculations.
-Significant efforts have been dedicated to introduce explicit polarizability into MD simulations, in the hopes of reaching a more accurate representation of the system~\cite{Thole1981,ando2001stable,Grossfield2003,
-lamoreux2003,Antila2013 ,Lemkul2016, baker2015polarizable,jing2019polarizable }.
-This effort has been motivated by the plethora of interactions where polarization is perceived to be a key contribution to correctly describing the system.
-These include description of water, ion hydration and ion binding to molecules, cation--$\pi$ and $\pi$-$\pi$ interactions~\cite{inakollu2020polarisable}.
-These lower-level interactions also translate to the behaviour of larger-scale biological systems, such as ion channels, where ion-selectivity and ion currents may be affected by polarization~\cite{roux2011ion, klesse2020induced, prajapati2020computational,Yue2022}, and telomeric DNA~\cite{Salsbury22} where the conformations adopted are mediated by ionic interactions.
+Classical molecular dynamics (MD) simulations are nowadays widely and almost routinely used to model wide range of biomolecular complexes~\cite{gupta22}.
+In conventional MD simulations, electrostatic interactions are described by assigning static point charges to the atoms and molecules at the beginning of the simulation as defined in the inputted model---the force field.
+Dynamic effects arising from electronic polarizability are traditionally not explicitly included and only considered in an averaged fashion within the force field parameterization process where parameters are obtained by fitting to macroscopic observables or to \textit{ab initio} calculations.
+However, electronic polarization is perceived to be a key contribution to correctly describe many biomolecular systems.
+These include description of water, ion hydration and ion binding to molecules, cation--$\pi$ and $\pi$-$\pi$ interactions~\cite{inakollu2020polarisable}, and general co-operativity in interactions~\cite{nochebuena23}.
+These lower-level interactions also translate to the behaviour of larger-scale biomolecular systems, such as ion channels, where ion-selectivity and ion currents may be affected by polarization~\cite{roux2011ion, klesse2020induced, prajapati2020computational,Yue2022}, and telomeric DNA~\cite{Salsbury22} where the conformations adopted are mediated by ionic interactions.
+Therefore, significant efforts have been dedicated to introduce explicit polarizability into MD simulations in the hopes of reaching a more accurate representation of the system~\cite{Thole1981,ando2001stable,Grossfield2003,lamoreux2003,Antila2013,Lemkul2016,baker2015polarizable,jing2019polarizable}.
 
-For bilayer membrane simulation, specifically,  the molecular (dielectric) environment varies dramatically while crossing the membrane from the water phase across the dipolar/charged lipid headgroup interface to the hydrophobic tail region.
+In a bilayer membrane, specifically, the molecular (dielectric) environment varies dramatically while crossing the membrane from the water phase across the dipolar/charged lipid headgroup interface to the hydrophobic tail region.
 Therefore, including polarizability in the lipids is expected to provide more accurate description, especially in situations pertaining to membrane binding processes, membrane potential and translocation of charged biomolecules across the membrane, and the behavior of molecules residing within membranes, such as membrane proteins~\cite{baker2015polarizable,Lemkul2016, li2017drude,chu2018anionicpolarizable,Lynch21,Chen2021,Melcr:2018a, melcr2019improved,nencini22}.
+However, the quality of polarizable lipid models have not been evaluated on an equal footing with non-polarizable models.
 
-The available lipid force fields with explicit electronic polarization include CHARMM-Drude~\cite{li2017drude, yu2023drude}, AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable}, and CHARMM-Fluctuating Charge (FQ)~\cite{lucas2012charge}.
+The available lipid force fields with explicit electronic polarization include CHARMM-Drude~\cite{li2017drude, yu2023drude}, AMOEBA-based~\cite{chu2018anionicpolarizable,chu2018polarizable}, and CHARMM-Fluctuating Charge (FQ)~\cite{lucas2012charge} parameters.
 The underlying strategy for including explicit polarizability differ among these models:  1) the classical Drude oscillator (CHARMM-Drude) models polarization by separating a core and a shell charge with a spring which stretches according to the environment giving the site a fluctuating dipole moment~\cite{Lemkul2016},
-2) the induced point dipole/multipole (AMOEBA) uses polarizable point dipoles placed on chosen sites of the molecule~\cite{ponder2010current},
+2) the induced point dipole/multipole, Atomic Multipole Optimized Energetics for BiomoleculAr (AMOEBA), uses polarizable point dipoles placed on chosen sites of the molecule~\cite{ponder2010current},
 and 3) the electronegativity equilization (FQ) employs atomic charges that are no longer constant but can change (redistribute within the molecule) according to the electronegativities of the molecule atoms and the electric fields from the molecular environment~\cite{patel2004charmm}.
@@
 
-%Hence, there certainly is room for improvement: efforts from the NMRlipids open science community have shown that the current non-polarizable lipid models cannot correctly reproduce the conformational ensemble~\cite{Botan2015,Antila2019}, conformational dynamics~\cite{Antila2021}, or ion biding~\cite{Catte2016} within MD simulations.
-
 Our previous efforts in benchmarking state-of-the-art non-polarizable lipid force fields have demonstrated that the quality of predictions for important membrane properties greatly vary between different force fields, particularly for lipid headgroup conformational ensembles and ion binding affinities~\cite{Botan2015,Catte2016,Antila2019,bacle21,Antila2021,Antila2022,Databank}.
-The ability to capture these membrane properties correctly is important on its own but also creates the basis for the description of more complex systems: for example, ion binding affinity regulates membrane surface charge and wide variety of conformations available for lipid headgroups appears to be important for modeling realistic protein-lipid binding~\cite{bacle21}.
+The ability to capture these membrane properties correctly is important on its own but also creates the basis for the description of more complex systems: for example, ion binding affinity regulates membrane surface charge and having wide variety of conformations available for lipid headgroups appears to be important for capturing realistic protein-lipid interactions~\cite{bacle21}.
 Therefore, such benchmark studies are urgently needed also for polarizable lipid force fields, considering  the increased computational cost they come with, and most importantly, their pledge
@@
 
-Here we assess the quality of actively developed polarizable lipid force fields, CHARMM-Drude~\cite{li2017drude,yu2023drude} and  AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable}, using the resources and the framework from the NMRLipids open collaboration (\url{http://nmrlipids.blogspot.fi/}).
+Here we assess the quality of the actively developed polarizable lipid force fields, CHARMM-Drude~\cite{li2017drude,yu2023drude} and  AMOEBA-based~\cite{chu2018anionicpolarizable,chu2018polarizable} parameters, using the resources and the framework from the NMRlipids open collaboration (\url{http://nmrlipids.blogspot.fi/}).
 These two force fields were selected for comparison because they are increasingly utilized in biomolecular simulations and have parameters available for several lipids for which corresponding experimental data is available in NMRlipids databank~\cite{Databank}.
 We assess the structural quality of POPC, DOPC and POPE lipid bilayer simulations against experimental NMR and small angle X-ray scattering (SAXS) data using the quality metrics defined in the NMRlipids databank~\cite{Databank}.
-Cation binding to membranes was evaluated against NMR order parameters~\cite{Catte2016} and lipid conformational dynamics is benchmarked to data from NMR spin relaxation rate experiments~\cite{ferreira15,Antila2021}.
+Cation binding to membranes is evaluated against salt-induced changes NMR order parameters~\cite{Catte2016} and lipid conformational dynamics is benchmarked to data from NMR spin relaxation rate experiments~\cite{ferreira15,Antila2021}.
 Furthermore, for each experimental benchmark we compare the results from polarizable models to those from the best performing non-polarizable simulations obtained from the NMRlipids Databank\cite{Databank}.
+Our results will act as a useful reference for selecting the best polarizable lipid models for wide range of applications and for guiding the development of polarizable force field parameters for lipids.
 
-%All simulation data, analysis scripts, and results are publicly available in their corresponding links. No plausible nor implausible request is required to download, use, and disseminate the data.
+\section{Methods}
+
+\subsection{Using a polarizable force field for membrane simulations}
+
+While non-polarizable MD simulations of membranes can be nowadays routinely performed with several simulation packages and force fields, polarizable simulations bear many practical complications.
+%Secondly, there are compability issues with the most used MD engines and the polarizable models and that support might be version dependent. 
+Out of the currently popular MD simulation packages for membrane simulations, OpenMM~\cite{eastman2017openmm} supports both AMOEBA and Drude force fields, NAMD~\cite{phillips2020scalable} can only run the Drude force field, whereas GROMACS~\cite{abraham2015gromacs}
+%---the most popular MD engine within the bilayer simulation field---
+has only limited support for the Drude polarizable force field via an unofficial Git-branch~\cite{drudegithub}.  
+TINKER is widely used with AMOEBA, but it does not support semi-isotropic pressure coupling, which %decouples the pressure tensor in $xy$ and $z$-direction and is 
+is required in membrane simulations.%~\cite{xxx}.
+%crucial for correctly modelling membrane systems~\cite{xxx}.
+%
+Consequently, we have selected OpenMM for simulations in this work.
+%In general, easy access to MD engines, as well as adaptability of common analysis tools and libraries for polarizable force fields, will be crucial in facilitating more wide-spread utilization of these models.
+
+Another practical issue is availability of parameters for the molecules of interest.
+CHARMM-GUI can generate the topology and input parameters for the Drude force field, which greatly simplifies the employment of this model in particular~\cite{kognole2022charmm}.
+Standard protocols for AMOEBA lipid parameters are not available, but some parameters can be found online~\cite{amoebagithub,klesse2020induced}.
+
+%Considering the increased interest in polarizable force fields in membrane simulations, we mention here some hurdles in their practical usage. Firstly, 
+Lastly, while conducting polarizable membrane simulations, one should consider the increased computational cost arising from the explicit treatment of electronic polarizability.
+For the Drude-based models, this slowdown occurs both because the addition of the Drude particles increases the number of interaction pairs and the employed extended dual-Langevin thermostat requires a shorter 1~fs integration time step, compared to the 2~fs typically used for non-polarizable membrane simulations.
+The AMOEBA force field can use a multi time-step integration algorithm where the non-electronic interactions are iterated with a 2~fs time step whereas more computationally unstable polarization terms are iterated with a smaller time step.
+However, the multi-time step scheme  only partly mitigates the computational cost: one typically experiences almost a ten-fold decrease in the simulation speed while using the polarizable force fields.
+
+All simulations performed in this work are listed in Table~\ref{table:sim_details} with links to the openly available trajectory data.
+Data not mentioned here was obtained by analysing pre-existing trajectories from the NMRlipids databank.
+
+\subsection{Simulations with CHARMM-Drude parameters}
+
+CHARMM-Drude2017 simulations were performed with OpenMM 7.5.0~\cite{eastman2017openmm} using parameters extracted with \textit{Membrane Builder}~\cite{wu2014charmm,jo2009charmm,jo2007automated,lee2018charmm} and \textit{Drude Prepper}~\cite{kognole2022charmm} from CHARMM-GUI~\cite{jo2008charmm,lee2016charmm}.
+Before starting the simulations, membrane structures were equilibrated for 200~ns using the non-polarizable CHARMM36 force field~\cite{klauda2010update}, and the last frames of these simulations were used to generate the starting structures for the polarizable force field simulations.
+Ion parameters are obtained from Ref.~\citenum{lin2018polarizable} and SWM4-NDP water model is employed in all Drude simulations~\cite{lamoureux2006polarizable}.
+
+CHARMM-Drude2023 force field parameters are currently not integrated into CHARMM-GUI.
+Therefore, the simulation setups with NaCl and CaCl$_{2}$~\cite{lin2018polarizable} using SWM4-NDP water model~\cite{lamoureux2006polarizable} have been generated following the instructions in the original CHARMM-Drude2023 paper~\cite{yu2023drude} using CHARMM program~\cite{brooks2009charmm} and the last frame of 200~ns long CHARMM36 simulations (same for CHARMM-Drude2017).
+The salt-free CHARMM-Drude2023 simulations were obtained from Zenodo~\cite{richard_m_venable_2023_7872447, richard_m_venable_2023_7871949}.
+
+A dual Langevin thermostat was employed to keep the Drude particles and the rest of the system at 1.0~K and 303~K, respectively.
+A Drude-hardwall of 0.02~nm was utilized to keep the Drude particles close to their parent atoms.
+Semi-isotropic Monte Carlo barostat was used to couple pressure to 1~bar independently in membrane plane and normal directions.
+%while the pressure fluctuations along the $xy$-plane were coupled.
+A constant surface tension of 0.0~dyne/cm was applied throughout the simulations.
+Bonds containing the hydrogens were constrained.
+For CHARMM-Drude2017, particle mesh Ewald was used to compute the Coulomb interactions and Van der Waals interactions were set to 0 between 1.0~nm and 1.2~nm using a switching function.
+For the CHARMM-Drude2023, Lennard-Jones particle-mesh Ewald (LJ-PME) method was used to compute the long-range dispersions~\cite{wennberg2013lennard}.
+Simulation frames were saved in every 10~ps.
+
+\subsection{Simulations with AMOEBA-based parameters}
+
+For simulations with the AMOEBA-based force field, we used the OpenMM implementation of parameters developed by Chu \textit{et al.}~\cite{chu2018polarizable} available at GitHub~\cite{amoebagithub,klesse2020induced}.
+All AMOEBA simulations were run using OpenMM 7.1.1~\cite{eastman2017openmm}.
+The same initial structures as in CHARMM-Drude simulations were used.
+A multi-time step Langevin integrator was used to iterate the bonded and non-bonded interactions in every 0.5~fs and 2.0~fs, respectively.
+A non-bonded cutoff of 1.2~nm was applied while semi-isotropic Monte Carlo barostat was used to couple pressure to 1~bar independently in membrane plane and normal directions.
+Ion and water parameters were obtained from Ref.~\citenum{shi2013proteinamoeba}.
+%keeping the pressure at 1~bar using a Monte Carlo barostat.
+Pressure fluctuations in the $xy$-plane has been coupled using a semi-isotropic pressure coupling and a constant surface tension at 0.0~dyn/cm has been applied throughout the simulations.
+Simulation frames were saved in every 10~ps. Further simulations details can be found in the input files of the respective simulations (Table~\ref{table:sim_details}).
+
+\subsection{Analysis of simulations}
+All simulations were first uploaded to the NMRlipids databank~\cite{Databank}.
+Area per lipids, small angle X-ray scattering form factors, and order parameters ($S_{CH}$) are automatically calculated by the NMRlipids Databank~\cite{Databank} and were extracted from therein.
+Quality evaluation metrics were quantified as detailed in the NMRlipids databank~\cite{Databank} with the exception that POPC simulations at 303\,K were paired with the experimental data measured at 300\,K, while in the NMRlipids databank simulations are paired with experiments with the maximum temperature difference of two degrees.
+The order parameter qualities, $P^{\mathrm{headgroup}}$, $P^{sn-1}$, and $P^{sn-2}$, presents an average probability for order parameters of a given segment to locate within experimental values, taking the error bars of both values into account.
+Qualities of SAXS form factors, $FF_{q}$, depict the distance of the first form factor minima between simulation and experiment.
+This avoids the effects from the simulation size dependency of the SAXS data~\cite{Databank}.
+%~\cite{Databank} one should focus on the extrema $q_{z}$[\AA$^{-1}$] when comparing the simulated and experimental data.
+In particular $FFq$ only measures the difference between the first minima of the simulated and experimental curves.
+Note that in quantifying the bilayer electron densities for calculation of SAXS curves, the NMRlipids databank analysis algorithm places electrons as point charges at the atom centers without considering redistribution of charge density due to the polarizability.
+Nevertheless, we expect this approximation not to have significant effect on the resulting SAXS form factors.
+
+The density profiles were calculated using  MDAnalysis~\cite{michaud2011mdanalysis,gowers2016mdanalysis}/NumPy~\cite{harris2020array} for CHARMM-Drude and AMOEBA simulations, and 
+\textit{gmx density} command from Gromacs package for the CHARMM36 and ECClipids simulations, for which the trajectories where extracted from the NMRlipids databank.
+In all calculations, the center of mass of the system is translated to origin coordinates.
+All data is normalized such that the area under the curve is equal to unity.
+The $R_{1}$ rates and effective correlation times, along with the accompanying error estimates, were quantified from the trajectories as elaborated in Ref.~\citenum{Antila2021}.
+In-house python script available at \url{https://github.com/NMRLipids/NMRlipidsVIpolarizableFFs/tree/master/scripts/correlation_times} was used for this purpose.
+
+%\newpage
+%\begin{adjustbox}{angle=90}
+\begin{table}[ht!]
+\begin{small}
+\begin{tabular}{cccccccccc}
+	Lipid/ion                & force field  & Ion (M) & N$_{l}$ & N$_{w}$ & N$_{ion}$ & T (K) & t$_{sim}$ (ns) & t$_{analysis}$ (ns) & files [Ref.]\\ \hline
+\multirow{2}{*}{POPC}            & Drude2017 & 0      & 144       & 6400       & 0         & 303    & 500              & 400         &          \citenum{kav_batuhan_2021_7607436}    \\ 
+                                 & Drude2023 & 0      & 72       & 2239       & 0         & 303      & 300              & 200         &          \citenum{kav_batuhan_2023_7916287}, \citenum{richard_m_venable_2023_7871949}    \\ \hline
+\multirow{3}{*}{POPE}                              & Drude2017 & 0      & 144       & 6400       & 0         & 308.15    & 350              & 300         &          \citenum{kav_batuhan_2021_7604627}    \\                                            & Drude2023 & 0       & 72        & 2304      &  0         & 303    & 300             & 200         &           \citenum{richard_m_venable_2023_7872447}, \citenum{kav_batuhan_2023_7916494} \\
+                                                  & AMOEBA     & 0       & 72        & 2880      &  0        & 303     & 305.94          & 305.94      & \citenum{kav_batuhan_2023_7622838} \\ \hline
+	\multirow{5}{*}{POPC:NaCl}        & Drude2017 & 0.350   & 144      & 6400     & 41         & 303   & 500             & 400                  & \citenum{kav_batuhan_2020_7586915}   \\
+				  & Drude2017 & 0.450   & 144      & 6400     & 51         & 303   & 500             & 400                  & \citenum{kav_batuhan_2020_7591753}   \\
+				  & Drude2017 & 0.650   & 144      & 6400     & 77         & 303   & 500             & 400                  & \citenum{kav_batuhan_2020_7596011}   \\
+				  & Drude2017 & 1.0     & 144      & 6400     & 115        & 303   & 500             & 400                  & \citenum{kav_batuhan_2020_7600326}   \\ 
+                    & Drude2023 & 0.350      & 128       & 6400      & 41          & 303    & 223.97      &       220.14       & \citenum{kav_batuhan_2023_8000095} \\     
+                    & Drude2023 & 1.0      & 128       & 6400      & 115          & 303    & 220.14      &       220.14       & \citenum{kav_batuhan_2023_8000133} \\ \hline
+
+	\multirow{5}{*}{POPC:CaCl$_{2}$} & Drude2017 & 0.350   & 144      & 6400     & 41         & 303   & 500             & 400                  & \citenum{kav_batuhan_2020_7600827}   \\
+				  & Drude2017 & 0.450   & 144      & 6400     & 52         & 303   & 500             & 400                  & \citenum{kav_batuhan_2020_7605016}   \\
+				  & Drude2017 & 0.650   & 144      & 6400     & 76         & 303   & 500             & 400                  & \citenum{kav_batuhan_2020_7604040} \\
+				  & Drude2017 & 1.0   & 144      & 6400     & 114         & 303   & 500             & 400                  & \citenum{kav_batuhan_2023_7658975}\\
+                    & Drude2023 & 0.350 & 128      & 6400     & 41          & 303   & 219.59          & 219.59               & \citenum{kav_batuhan_2023_8000065}\\
+                    & Drude2023 & 0.790 & 128      & 6400     & 91          & 303   & 214.85          & 214.85               & \citenum{kav_batuhan_2023_7992137} \\ \hline
+DOPC           & AMOEBA   & 0 & 72 & 2880 & 0 & 303 & 201.61 & 201.61 & \citenum{kav_batuhan_2022_7604681} \\ \hline
+\multirow{2}{*}{DOPC:NaCl} & AMOEBA & 0.450 & 72 & 2880 & 17 & 303 & 218.41 & 218.41 & \citenum{kav_batuhan_2022_7604711} \\
+                           & AMOEBA & 1.0 & 72 & 2880 &  35 & 303 & 201.61 & 201.61 & \citenum{kav_batuhan_2023_7625844} \\ \hline
+\multirow{2}{*}{DOPC:CaCl$_{2}$} & AMOEBA  & 0.450 & 72 & 2880 & 16 & 303 & 218.41 & 218.41 & \citenum{kav_batuhan_2022_7604842} \\
+                                 & AMOEBA & 1.0 & 72 & 2880  & 36 & 303 & 218.41 & 218.41 & \citenum{kav_batuhan_2022_7604810}
+
+\end{tabular}
+\end{small}
+\caption{Systems simulated for this work.
+Column N$_l$ gives the number of lipids, N$_w$ gives the number of water molecules, T(K) denotes the temperature in kelvins.
+Simulated time is listed in column t$_{sim}$ and time used for analysis in t$_{analysis}$.
+Column "files" gives the reference to open access simulation data.
+The salt concentrations calculated as [salt]=N$_c$×[water] / N$_w$, where [water] = 55.5 M}
+\label{table:sim_details}
+\end{table}
+\todo{check the simualation settings (especially the temperature}
 
@@
 
-To evaluate the structural properties of lipid bilayers in simulations with polarizable force fields, we simulated POPC and POPE lipid bilayers with CHARMM-Drude2017~\cite{li2017drude} and CHARMM-Drude2023~\cite{yu2023drude} parameters, and POPE and DOPC bilayers with AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable} parameters.
-These systems were selected based on simultaneous availability of force field parameters, and experimental data~\cite{Databank}.
-After the data production, simulation trajectories were added into the NMRlipids Databank and its quality metrics was used to evaluate each simulation against experiments~\cite{Databank}.
-This metric measures the quality of lipid bilayer simulations in two aspects: first, it evaluates the quality of conformational ensemble of individual lipids against C-H bond order parameters from NMR, and second, it examines consistency of membrane dimensions compared to small angle X-ray scattering (SAXS) data~\cite{Databank,ollila16}.
-%(measuring the electron density variation across the membrane). 
-The former is divided into three parts that separately describe the average quality of headgroup, glycerol backbone and acyl chain qualities as described in the NMRlipids databank~\cite{Databank}.
+To evaluate the structural properties of lipid bilayers in simulations with polarizable force fields, we simulated POPC and POPE lipid bilayers with CHARMM-Drude2017~\cite{li2017drude} and CHARMM-Drude2023~\cite{yu2023drude} parameters, and POPE and DOPC bilayers with AMOEBA-based~\cite{chu2018anionicpolarizable,chu2018polarizable} parameters.
+These systems were selected based on simultaneous availability of force field parameters and experimental data~\cite{Databank}.
+After the data production, simulation trajectories were added into the NMRlipids Databank and its quality metric were used to evaluate each simulation against experiments~\cite{Databank}.
+The metric measures the quality of lipid bilayer simulations in two aspects: first, it evaluates the quality of conformational ensemble of individual lipids against C-H bond order parameters from NMR, and second, it examines consistency of membrane dimensions compared to small angle X-ray scattering  data~\cite{Databank,ollila16}.
+The former is divided into three parts that separately describe the average quality of headgroup, glycerol backbone and acyl chain regions.
 While order parameters primarily reflect lipid conformations, acyl chain order parameters are also a good proxy for membrane packing since a smaller area per lipid tends to increase the magnitude of the order parameters~\cite{Databank}.
@@
 
-\begin{figure}[t]
-    \centering
+%\begin{figure}[t]
+%    \centering
     %\includegraphics{Figures/order_parameters.png}
-    \includegraphics[width=0.4\textwidth]{Figures/molecules.pdf}
-    \caption{Molecules and atom labelling used.}
-    \label{fig:molecules}
-
-\end{figure}
+    %\includegraphics[width=0.4\textwidth]{Figures/molecules.pdf}
+    %\caption{Molecules and atom labelling used.}
+%    \label{fig:molecules}
+%
+%\end{figure}
 
@@
     \caption{X-ray scattering form factors, and C-H bond order parameters for headgroup and glycerol backbone, and acyl chains (from top to bottom) compared between simulations and experiments using the NMRlipids databank.
-The atom labelling is as depicted in Fig.~\ref{fig:molecules}.
 Experimental data in this figure are originally reported in Refs.~\citenum{ferreira13,bacle21,Databank,NMRlipidsIII,kucerka2015}.
-For CHARMM-Drude2023 simulations we selected representative replicas among three available simulations, for example for differences between replicas for POPC see Fig.~\ref{fig:order_parameters_replicas}.}
+For CHARMM-Drude2023 simulations we selected representative replicas among three available one (for all POPC replicas, see SI Fig.
+S1).
+Right most panel shows the modelled lipids and the used carbon naming scheme.}
     \label{fig:order_parameters}
@@
 Direct comparisons between simulated and experimental data are visualized in Fig~\ref{fig:order_parameters} and the resulting quality metrics are shown in Table~\ref{tab:my_label}.
+Differences of headgroup conformations between force fields are shown in terms of dihedral distributions in Supplementary Information Fig.~S2.
 Similarly to the non-polarizable CHARMM36 simulations, CHARMM-Drude2017 simulation predicts slightly too packed membrane with excessively negative order parameters with respect to experiments and simulations with the highest quality in the NMRlipids databank (OPLS3e)~\cite{Databank}.
@@
 the quality of headgroup conformations in the CHARMM-Drude2017 simulations is worse than in its non-polarizable counterpart.
-This is likely because headgroup and glycerol backbone parameters were optimized to reproduce absolute average values of the experimentally determined order parameters, instead of taking into account the order parameter forking (meaning measurably different order parameters for different C--H bonds at a single carbon atom) and order parameter sign~\cite{Antila2022}.
+This is likely because headgroup and glycerol backbone parameters were optimized to reproduce absolute average values of the experimentally determined order parameters, instead of taking into account the order parameter forking (measurably different order parameters for different C--H bonds at a single carbon atom) and the order parameter sign~\cite{Antila2022}.
 A better description of the headgroup and glycerol backbone is provided by CHARMM36-Drude2023, yet its quality is still below that offered by the non-polarizable CHARMM36 parameters (see also discussion about conformational dynamics below).
-Also the quality of membrane packing and acyl chain order are improved in CHARMM36-Drude2023 compared to the earlier version but again, it did not perform at a quality comparable to that of the best available simulations.
+Also the quality of membrane packing and acyl chain order are improved in CHARMM36-Drude2023 compared to the earlier version but again, it is outperformed by the best available simulations (OPLS3e in Table~\ref{tab:my_label}).
 %also provides better description of the carbon tails. 
@@
 
- AMOEBA simulations capture the headgroup and glycerol backbone order parameters reasonably well with the exception of g$_1$,
+ AMOEBA-based simulations capture the headgroup and glycerol backbone order parameters reasonably well with the exception of g$_1$,
 %ll for both DOPC and POPE but vastly overestimates one of the g$_3$ values, resulting in 
@@
 These low order parameters signal an important mechanism through which acyl chain double bonds affect membrane properties~\cite{ollila07} and are well reproduced in all state-of-the art MD simulation parameters~\cite{ollila16}.
-Furthermore, AMOEBA parameters substantially overestimate the area per lipid in POPE simulations, leading to too disordered acyl chains.
-Similar issues are evident also in membrane data presented in the recent publication for the AMOEBA cholesterol model~\cite{Li23chol}.
+Furthermore, AMOEBA-based parameters substantially overestimate the area per lipid in POPE simulations, which is connected to too disordered acyl chains.
+Similar issues are evident also in the membrane data presented in the recent publication for the AMOEBA-based cholesterol model~\cite{Li23chol}.
 %also clearly struggles to depict the tail conformations, in particular in the double bond region of DOPC (Fig.~\ref{fig:order_parameters}). 
@@
 %
-The unsatisfactory description of the lipid tail region and the area per lipid is further reflected in the inability of AMOEBA to capture the SAXS curves.
-Thus, we can conclude that our simulations with the AMOEBA parameters did not reproduce essential
-membrane properties. 
+The unsatisfactory description of the lipid tail region and the area per lipid is further reflected in the inability of AMOEBA-based parameters to capture the SAXS curves.
+Thus, we can conclude that the AMOEBA-based parameters used in our simulations did not reproduce essential
+membrane properties at the level of the state-of-the-art lipid parameters. 
 %The higher $FFq$ values (Table~\ref{tab:my_label}) of the polarizable force fields in general indicate that they are not able confer consistency with the SAXS data as well as the best non-polarizable force fields (OPLS3e).
@@
 Instead, they essentially describe the average of the distributions and do not contain information of the dynamics of the conformational sampling.
-Therefore, a simulation that reproduces the order parameters has an ensemble that is only potentially correct, and even the correct ensemble may not have the experimentally observed dynamics.
-To further elucidate this aspect in the case of polarizable force fields, we compared NMR $^{13}$C spin relaxation $R_{1}$ rates and C-H bond effective correlation times $\tau_{\mathrm{eff}}$ with experiments~\cite{ferreira15} and best non-polarizable simulations from our previous study~\cite{Antila2021} for simulations with PC headgroups and the glycerol backbone in Fig.~\ref{fig:correlation_times}.
-This focus was chosen based on the availability of both experimental data and polarizable simulations.
-$R_{1}$ rates measured with typical magnetic field strengths are sensitive to rotational dynamics of C-H bonds with the timescales around $\sim$0.1-1~ns, while effective correlation times respond to a larger range of dynamical processes up to $\sim$1000~ns~\cite{ferreira15}.
+Therefore, a simulation that reproduces the order parameters has an ensemble that is only potentially correct, and even the correct ensemble may not be sampled at the experimentally observed dynamics.
+To further elucidate this aspect in the case of polarizable force fields, we compared their $^{13}$C NMR spin relaxation rates, $R_{1}$, and C--H bond effective correlation times, $\tau_{\mathrm{eff}}$, with experiments~\cite{ferreira15} and best non-polarizable simulations from our previous study~\cite{Antila2021} in Fig.~\ref{fig:correlation_times}.
+Here, we focus on PC headgroups and the glycerol backbone due to the availability of both experimental data and polarizable simulations.
+$R_{1}$ rates measured with typical magnetic field strengths are sensitive to rotational dynamics of C--H bonds with the timescales around $\sim$0.1-1~ns, while effective correlation times respond to a larger range of dynamical processes up to $\sim$1000~ns~\cite{ferreira15}.
 The effective correlation time can be interpreted as an average measure for how fast molecular conformations sample the phase space that leads to the average order parameters.
 
-The effective correlation times in CHARMM-Drude2017 and CHARMM-Drude2023 simulations are approximately two  and one orders of magnitude slower, respectively, than the values suggested by experiments and best available simulations (Fig.~\ref{fig:correlation_times}).
+The effective correlation times in CHARMM-Drude2017 and CHARMM-Drude2023 simulations are approximately two and one orders of magnitude slower, respectively, than the values extracted from experiments and best available simulations (Fig.~\ref{fig:correlation_times}).
 This indicates that not only are the simulations with this polarizable force field computationally costlier for equivalent lengths of trajectory, but one would also have to simulate longer to obtain converged results.
 The non-polarizable counterpart of the Drude models, CHARMM36, exhibits much more realistic, faster effective correlation times.
-Also the dynamics in the fast range ($R_{1}$ rates), are on average slightly more realistic in the non-polarizable CHARMM36 compared to both of the polarizable counterparts.
+Also the dynamics in the fast range ($R_{1}$ rates), are on average slightly more realistic in CHARMM36 simulations compared to both of its polarizable counterparts.
 Interestingly, CHARMM-Drude models have been reported to have slower water hydrogen bonding dynamics around amino acids compared to their non-polarizable counterpart,~\cite{Ngo2019} which might align with an overall slower dynamics of the model in addition to enhanced water binding.
 
-In contrast to the Drude-based models investigated here, the $R_1$ spin relaxation times and effective correlation times from DOPC simulations with AMOEBA force field are close to experimental data from POPC and on par with the best non-polarizable models (Slipids and CHARMM36).
+In contrast to the Drude-based models investigated here, the $R_1$ spin relaxation times and effective correlation times from DOPC simulations with AMOEBA-based force field are close to experimental data from POPC and on par with the best non-polarizable models (Slipids and CHARMM36).
 The small difference in acyl chain composition (DOPC vs.
-POPC) is not expected to strongly affect headgroup dynamics due to effective decoupling between hydrophilic and hydrophobic membrane regions~\cite{Antila2022rot, Klauda08}.
+POPC) is not expected to  affect headgroup dynamics due to effective decoupling between hydrophilic and hydrophobic membrane regions~\cite{Antila2022rot, Klauda08}.
 %appears to be faulty in the fast range ($R_{1}$ rates) and, furthermore, $\tau_{\mathrm{eff}}$ values reveal vastly too slow overall dynamics.
@@
 Experimental values are obtained from Ref.~\citenum{Antila2022rot}.
+For simulated $\tau_{\rm{eff}}$, the data point quantifies
+the average over C–H bonds.
+If $\tau_{\rm{eff}}$ could not be determined for all bonds due to slow convergence with respect to simulation time, only the range from the mean of the lower to the mean of the upper error estimates is shown.
 All simulations used here were salt free.}
@@
 
-Given the abundance of cations in biological systems, accurately capturing their interactions with membranes has the uttermost importance in MD simulations.
+Given the abundance of cations in biological systems, accurately capturing their interactions with membranes in simulations has the uttermost importance.
 A wealth of  experimental evidence shows that monovalent ions (except for lithium) exhibit very weak binding affinity of PC lipid bilayers, while multivalent ions like calcium bind more strongly~\cite{Catte2016}.
 However, simulations with non-polarizable force fields without any additional corrections systematically overestimate cation binding to lipid bilayers~\cite{Catte2016}.
-Implicit inclusion of electronic continuum correction (ECC) to both ion and lipid parameters can substantially improve the situation~\cite{Melcr:2018a,melcr2019improved,bacle21}, suggesting that electronic polarizability plays an important role in ion binding to membranes, which might further lead one to expect that simulations with polarizable force fields will perform better in this aspect and more accurately describe ion binding to membranes.
+Implicit inclusion of electronic continuum correction (ECC) to both ion and lipid parameters can substantially improve the situation~\cite{Melcr:2018a,melcr2019improved,bacle21}, suggesting that electronic polarizability plays an important role in ion binding to membranes.
+This might further lead one to expect that simulations with explicitly polarizable force fields will more accurately describe ion binding to membranes.
 %Finally, we address the premise that adding polarizability to force fields improves the depiction of ion binding to biomembranes in simulations. 
-To test this notion, we evaluated ion binding to membranes against the experimental NMR data using the "lipid electrometer," which involves quantifing the amount of ion binding to the membrane by monitoring the change in lipid headgroup order parameters in response to increasing salt concentration~\cite{Catte2016}.
-Changes in these order parameters induced by increasing NaCl or CaCl$_2$ concentration are shown in Fig.~\ref{fig:popc_order_parameter_change} for simulations and experiments, whereas Fig.~\ref{fig:ion_density_profiles} shows the density profiles of ions with respect to the bilayer normal.
-Results from AMBER-based ECC-lipid model are shown for the reference that gives a good agreement with experiments for cation binding~\cite{Melcr:2018a}.
+To test this notion, we evaluated ion binding to membranes against the experimental NMR data using the "lipid electrometer", in which the amount of ion binding to the membrane is quantified by monitoring the change in lipid headgroup order parameters in response to increasing salt concentration~\cite{seelig87,Catte2016}.
+Changes in these order parameters, as induced by increasing NaCl or CaCl$_2$ concentration, are shown in Fig.~\ref{fig:popc_order_parameter_change} for simulations and experiments, whereas Fig.~\ref{fig:ion_density_profiles} shows the density profiles of ions with respect to the bilayer normal.
+Results from AMBER-based ECClipids model are presented as a reference which gives a good agreement with experiments for cation binding~\cite{Melcr:2018a}.
 We also show data from the non-polarizable CHARMM36 where NBFIX correction for the ion models was specifically developed to address overbinding~\cite{Venable2013,Han2018}.
@@
 	\centering
-	\includegraphics{Figures/order_parameter_change.png}
+	\includegraphics[width=0.95\textwidth]{Figures/order_parameter_change.png}
 	\caption{The change in the lipid head group order parameters $\beta$ (top row) $\alpha$ (bottom row) upon increasing ion concentration with respect to the simulations without salt.
-CHARMM36 and ECC data is reproduced using the Zenodo repositories Ref.~\cite{ollila_2015_32496, ollila_2015_32497, ollila_2015_32498, nencini_ricky_2019_3434396} and Ref.~\cite{melcr_josef_2017_3335503}, respectively.}
+CHARMM36 and ECClipids data is reproduced using the Zenodo repositories at Refs.~\citenum{ollila_2015_32496, ollila_2015_32497, ollila_2015_32498, nencini_ricky_2019_3434396} and Ref.~\citenum{melcr_josef_2017_3335503}, respectively.}
 	\label{fig:popc_order_parameter_change}
@@
     \includegraphics{Figures/ion_density_profiles_with_chloride.eps}
-    \caption{Density profiles along the membrane normal for the CHARMM36 (first row)~\cite{Catte2016}, Drude 2017 and 2023 POPC (second and third rows), AMOEBA DOPC (fourth row), and electronic continuum correction (ECC, last row)~\cite{Melcr:2018a}.
-Solid and dashed lines are for the systems with the NaCl and CaCl$_{2}$, respectively.
-Please note that only 350~mM and 450~mM CaCl$_{2}$ concentrations are shown for the Drude and AMOEBA force fields, respectively, while 1000~mM NaCl concentration is shown for all force fields.
-CHARMM36 NaCl simulations are at 950~mM.
-CHARMM36 data is reproduced using the Zenodo repositories Ref.~\cite{ollila_2015_32496, ollila_2015_32497, ollila_2015_32498, nencini_ricky_2019_3434396}}
+    \caption{Density profiles along the membrane normal for the CHARMM36~\cite{Catte2016}, CHARMM-Drude2017, CHARMM-Drude2023, AMOEBA, and ECClipids~\cite{Melcr:2018a}.
+Data with circle markers and dashed lines in first and second column denote density profiles calculated from the water oxygens.
+Solid and dashed lines in the third column denote Cl$^{-1}$ density from NaCl and CaCl$_{2}$, respectively.
+Note that for CaCl$_{2}$ 350~mM (AMOEBA and CHARMM36) and 450~mM (Drude models and ECC) concentrations are shown, while 1000~mM NaCl concentration is shown for all force fields, except CHARMM36 simulations are at 950~mM NaCl.
+CHARMM36 data is reproduced using the Zenodo repositories Ref.~\citenum{ollila_2015_32496, ollila_2015_32497, ollila_2015_32498, nencini_ricky_2019_3434396}.
+Results are from POPC simulations for all force fields other than AMOEBA (DOPC).}
     \label{fig:ion_density_profiles}
@@
 
-The CHARMM-Drude2017 predicts similar calcium ion density profile to ECC-lipids, in good agreement with experiments.
-However, the sodium binding is equally strong in these simulations in contrast with the experimental evidence~\cite{Catte2016}.
+The CHARMM-Drude2017 predicts similar calcium ion density profile to ECClipids--a model which is in good agreement with experiments.
+However, the sodium binding is equally strong in CHARMM-Drude2017 simulations in contrast with the experimental evidence~\cite{Catte2016}.
 %, evidenced both by large change in the order parameters and the ion density profiles.
@@
 In strong contrast, experimentally one sees no sodium binding and some binding of calcium ions.
-Furthermore, the response of headgroup order parameters to the bound ions is not in qualitative agreement with experiments for CHARMM-Drude2017 simulations in Fig.~\ref{fig:popc_order_parameter_change}.
-Particularly, increasing $\alpha$ carbon order parameters and forking upon ion binding are not observed in experiments.
+and the response of headgroup order parameters to the bound ions is not in qualitative agreement with experiments ( Fig.~\ref{fig:popc_order_parameter_change}).
+In particular, increasing $\alpha$ carbon order parameters and forking upon ion binding are not observed in experiments.
 This is in contrast with the results from previous bechmarking of non-polarizable simulations~\cite{Catte2016}, where experimentally observed decrease of order parameters to more negative values upon ion binding were observed in all simulations, even though the binding affinity was often inaccurately predicted.
-This qualitative discrepancy in CHARMM-Drude2017 simulations most likely results from incorrect lipid headgroup conformational ensemble (see previous sections), which leads to inaccuracies in the structural response of the ensemble to ion binding.
+This qualitative discrepancy in CHARMM-Drude2017 simulations may result from incorrect lipid headgroup conformational ensemble (see previous sections), which leads to inaccuracies in the structural response of the ensemble to ion binding.
 Excessive sodium binding in the CHARMM-Drude model has been observed before for systems containing peptides or amino acids~\cite{Ngo2019, Kav2022} and deep-eutectic solvents~\cite{shayestehpour2022ion}.
@@
 
-In simulations with CHARMM-Drude2023 parameters, sodium and calcium ion binding are in line with ECC-lipid simulations when comparing density profiles, although calcium binding affinity is slightly larger, ions penetrate deeper into the bilayer, and there is a stronger formation of ionic double layer by subsequent attraction of chloride into the bilayer in CHARMM-Drude2023 when compared with ECC-lipids.
+In simulations with CHARMM-Drude2023 parameters, sodium and calcium ion binding are in line with ECClipids simulations when comparing density profiles, although calcium binding affinity is slightly larger, ions penetrate deeper into the bilayer, and there is a stronger formation of ionic double layer and deeper penetration of chloride into the bilayer in CHARMM-Drude2023 than in ECClipids.
+Interestingly, these features are observed in all simulations with calcium using polarizable force fields.
 %force field: the change of the order parameters is diminished to be more in line with experiments and only the some accumulation of sodium ions near the phosphate oxygen is seen in the ion distributions.
 %However, this the ion layering is most prominently present in the original CHARMM36 force field, for for sodium and calcium ions. 
-However, sodium or calcium ion binding seems to induce forking of lipid headgroup order parameters in CHARMM-Drude2023 simulations in contrast to experiments. 
+However, sodium or calcium ion binding seems to again induce forking of lipid headgroup order parameters in CHARMM-Drude2023 simulations in contrast to experiments. 
 %at the $\beta$ carbon in particular. 
@@
 
-%In the non-polarizable counterpart, CHARMM36 model with the NBFIX correction, sodium biding is similar to the polarizable models outside small accumulation peak in the phosphate region, and stronger subsequent accumulation of anions outside the phosphate region.
-The structural response well in line with the experiments.
-The divalent cation biding for CHARMM36 is weakest of all models with binding depth, and following anion binding, quite similar the ECC.
+In the non-polarizable counterpart, CHARMM36 model with the NBFIX correction, sodium biding is similar to the ECClipids and CHARMM-Drude2023 models, and subsequent accumulation of anions outside the phosphate region is stronger.
+The structural response is well in line with the experiments: it is similar to the ECClipids model but weaker than for CHARMM-Drude2023.
+The divalent cation binding for CHARMM36 is weaker than in ECClipids simulations, and order parameter response is smaller than in ECClipids and in experiments for the $\alpha$ carbon.
+Comparing calcium distributions from CHARMM36 simulations with CHARMM-Drude2023 model suggests that the polarizable model may better capture the difference in the relative amounts of sodium and calcium bound than correcting the ion binding by scaling the Lennard-Jones parameters (NBFIX).
+That said, the overall structural response to ion binding in CHARMM36 appears more realistic.
+%is closes to the ECC in magnitude and binding depth, but again CHARMM36 produces comparatively strong subsequent Cl accumulation.
 However, the structural response of the $\alpha$ carbons is weaker than in ECC and experiments.
-Contrasting the CHARMM36 with the CHARMM-Drude models and the ECC model results suggests that so far correcting the ion biding by scaling the Lennard-Jones parameters (NBFIX) of the non-polarizable models has resulted in slightly more realistic ion binding than introducing explicit polarizability to the model.
+Comparing the CHARMM36  structural response with the results from CHARMM-Drude models suggests that so far correcting the ion biding by scaling the Lennard-Jones parameters (NBFIX) of the non-polarizable models has resulted in slightly more realistic ion binding than introducing explicit polarizability to the model.
 %
@@
 
-In AMOEBA simulations, sodium binds weakly and does not affect order parameters, consistent with experiments and ECC-lipid simulations.
-Calcium binding affinity is similar to ECC-lipids, but order parameters also do not change upon  binding contrasting the experiments.
+In AMOEBA-based simulations, sodium binds weakly and does not affect order parameters, consistent with experiments and ECClipids simulations.
+Calcium binding affinity is similar to ECClipids, but order parameters do not change upon binding contrasting the experiments.
 This can be explained by the binding position of calcium, which is outside phosphate density peak in AMOEBA simulations.
@@
 
-In conclusion, incorporating explicit polarizability as implemented in CHARMM-Drude or AMOEBA parameters used here do not necessarily lead to improved description of cation binding to phospolipid membranes.
-However, these parameters do not correctly capture the response of lipid headgroup to cation binding to membrane, which most likely arises from inaccuracies in lipid parameters.
+In conclusion, incorporating explicit polarizability as implemented in CHARMM-Drude or AMOEBA-based parameters used here do not necessarily lead to improved description of cation binding to phospholipid membranes.
+These parameters do not correctly capture the response of lipid headgroup to cation binding to membrane, which most likely arises from inaccuracies in lipid parameters.
 Because such inaccuracies can also affect ion binding, it is difficult to isolate the explicit influence of polarizability per se on ion binding based on the results presented here.
@@
 In addition, the AMOEBA parameters struggle to capture the basic membrane properties that are reproduced by all current state-of-the-art MD simulation models.
-Nevertheless, CHARMM-Drude2023 predicts higher affinity for calcium than for sodium, which is in line with the wide range of experimental data~\cite{Catte2016}.
-This is a major improvement to non-polarizable CHARMM36 and CHARMM-Drude2017 which predict approximately similar binding to these ions in contrast to experiments.
+%Nevertheless, CHARMM-Drude2023 predicts higher affinity for calcium than for sodium, which is in line with the wide range of experimental data~\cite{Catte2016}.
+%This is a major improvement to non-polarizable CHARMM36 and CHARMM-Drude2017 which predict approximately similar binding to these ions in contrast to experiments.
 
@@
 
-%\textcolor{red}{For the ion binding discussion: drude calcium overbinds to proteins. this is known here https://pubs.acs.org/doi/full/10.1021/acsphyschemau.1c00039
-%drude sodium overbinds to proteins, as shown here https://pubs.rsc.org/en/content/articlepdf/2022/ra/d2ra01478e and https://onlinelibrary.wiley.com/doi/abs/10.1002/adts.201800106.
-Venable also cautioned us against using any ion parameters with the drude force fields.
-This is important as 1- polarizable force fields are supposed to be better for ion parameters and 2- they are generally used to study the ion channels.}
-
 \subsection{Conclusions}
-Including electronic polarizability to MD simulation models of membranes is expected to improve the description of bilayer polar regions.
-However, performance of membrane simulations have not been evaluated with equal footing with non-polarizable membranes.
-Here, we used the quality evaluation metrics defined in the NMRlipids databank~\cite{Databank} together with additional analyzes on dynamics and ion binding to evaluate quality of two available polarizable lipid parameters, CHARMM-Drude~\cite{li2017drude, yu2023drude} and AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable}, against
+Including electronic polarizability to MD simulation models of membranes is expected to improve the description of bilayer polar regions and their interactions with charged molecules, thereby making MD simulations of complex biomolecular systems more realistic.
+However, quality of polarizable membrane simulations have not been evaluated on equal footing with non-polarizable ones.
+Here, we used the quality evaluation metrics defined in the NMRlipids databank~\cite{Databank} together with additional analyzes on dynamics and ion binding to evaluate the performance of two available polarizable lipid models, CHARMM-Drude~\cite{li2017drude, yu2023drude} and AMOEBA-based~\cite{chu2018anionicpolarizable,chu2018polarizable} parameters, against
 %We have rigorously compared simulation output of model lipid bilayers using polarizable lipid force fields with high-resolution 
 experimental data and results from the best-performing non-polarizable force fields. 
-Considering the complexity and cost of simulations with polarizable models, it is crucial to understand their accuracy with respect to experiments and best non-polarizable models when planning simulations or evaluating the reliability of simulation results.
+Considering the complexity and additional computational cost of simulations with polarizable models, it is crucial to understand their accuracy with respect to experiments and choose the best models according to their respective strengths when planning simulations.
 
@@
 In most cases, the best non-polarizable models exhibits superior performance and accuracy in the categories investigated in this study.
-Our comparisons of lipid conformations and dynamics revealed room for improvement in current polarizable parameters to reach the level of the best currently available non-polarizable parameters.
-CHARMM36-Drude models predict slightly too ordered membrane and struggle to correctly capture headgroup conformations and dynamics.
-The former issue is present also in non-polarizable CHARMM36, while slow headgroup dynamics and incorrect response to ion binding is present only in polarizable CHARMM36-Drude simulations.
+Our comparisons of lipid conformations and dynamics ahow there is room for improvement in current polarizable parameterizations to reach the level of the best currently available non-polarizable parameters.
+Although the most recent CHARMM-Drude23 has improved the description of the molecular conformations and dynamics, both tested CHARMM-Drude models predict slightly too ordered membrane core and vastly too slow headgroup dynamics.
+%, struggle to correctly capture headgroup conformations and dynamics.
+The former issue is present also in non-polarizable CHARMM36, while slow headgroup dynamics and incorrect response to ion binding is present only in polarizable CHARMM-Drude simulations.
 %In contrast, the CHARMM-Drude2023 model has improved performance in this regard, notably surpassing its predecessor, CHARMM-Drude2017.
@@
 %Evaluation order parameters and scattering form factors reveals that 
-The tested AMOEBA parameters have difficulties to capture membrane ordering in acyl chain region,  %membranes with reasonable physical properties. 
+The tested AMOEBA-based parameters have difficulties capturing membrane ordering in acyl chain region and other more general membrane properties interconnected to chain ordering.
+%membranes with reasonable physical properties.
 %In regard to conformational dynamics, the AMOEBA force field produces results, specifically in 
-yet the PC and PE headgroup properties have similar or almost simular quality 
+Yet, the PC and PE headgroup conformations and dynamics have similar or almost similar quality 
 %are either on par with or better 
@@
 
-Sodium and calcium binding to membranes were evaluated against headgroup order parameter changes upon addition of NaCl or CaCl$_2$ in experiments and also compared with the ECClipids model with implicitly included electronic polarizability that currently gives the most accurate representation of ion binding against experiments.
-%The non-polarizable CHARMM36, when used with the NBFix ion model designed to provide more accurate binding, also rivals its polarizable counterparts. 
+Sodium and calcium binding to membranes in simulations were evaluated using experimentally observed headgroup order parameter changes upon addition of NaCl or CaCl$_2$.
+The binding in explicitly polarizable models was compared with the ECClipids model with implicitly included electronic polarizability that currently gives the most accurate response to ion binding against experiments, and to the non-polarizable CHARMM36, which when used with the NBFix ion model designed to provide more accurate binding also rivals its polarizable counterparts.
 %The polarizable AMOEBA and 
-CHARMM-drude2023 model exhibits a major improvement to non-polarizable CHARMM and CHARMM-Drude2017 parameters as it correctly reproduces the stronger calcium binding with respect to sodium that was not present in older parameters.
-This difference between sodium and calcium binding is also present in simulations with AMOEBA parameters.
+That said, CHARMM-drude2023 model provides an improved description of the stronger calcium binding compared to sodium.
+This difference between sodium and calcium binding is also present in simulations with AMOEBA-based parameters.
 However, calcium
-%have similar ion binding levels to the ECC model, but they exhibit variation in 4
+%have similar ion binding levels to the ECC model, but they exhibit variation in 
 binding depth, affinity, and the consequent structural response of the lipids do not exactly align with experiments or ECClipids results neither in CHARMM-Drude2023 nor AMOEBA simulations.
-The incorrect response to ion binding likely connects to the other discussed inaccuracies related to lipid conformational ensembles, dynamics and membrane order.
+The incorrect response to ion binding likely connects to the other discussed inaccuracies related to lipid conformational ensembles, dynamics, and membrane order.
 %structure and diversity in the conformational ensemble, as well as slow dynamics that lead to poor convergence of simulations.
 
- In summary, the full potential and promise of explicitly polarizable lipid force fields to improve the description of polar membrane regions have not yet been fully realized.
-However, it seems likely that this is not due to implementation of polarizability, but rather arises from compatibility of other parameters, suchs as dihedral and Lennart-Jones, with polarizable electrons.
-Due to long history in developing these parameters for non-polarizable force fields, it is not surprising that they
- %benefit from a longer history of intense scrutiny and
+ In summary, the full potential and promise of explicitly polarizable lipid force fields to improve the description of bilayer membranes have not yet been fully realized.
+However, it seems likely that this is not an inherent flaw in polarizability but rather in the current parameterization and their incombapility with parameters describing other interactions, such as those for Lennard-Jones interactions.
+Non-polarizable force fields benefit from a longer history of development and intense scrutiny, it is not surprising that they currently
  out-perform their polarizable counterparts in many respects, as we here show for the case of certain phospholipid bilayers. On the other hand, improvements  
 from CHARMM-Drude2017 to CHARMM-Drude2023 demonstrate the potential for tuning these parameters to improve polarizable force fields.
-Such endeavors are expected to substantially ease with emerging automized methods for parameter development~\cite{Antila2022,yu2023drude}.
+Such endeavors are expected to substantially ease with emerging automated methods for parameter development~\cite{Antila2022,yu2023drude}.
 
- %Doing so holds great potential for improvement, as demonstrated by the refinement of the CHARMM-Drude2017 model to the most recent CHARMM-Drude2023 version.
-
-\section{Methods}
-%\subsection{Practical pitfalls of polarizable biomembrane simulations}
-
+%\end{adjustbox}
 \begin{comment}
-    4
-\subsubsection{Building a polarizable force field}
-
-The atomic point charges describing the electrostatics in non-polarizable force fields are typically parameterized by fitting to the electrostatic potential energy surface around one or few conformers obtained from ab initio calculations.
-One then proceeds to build on top of that the parameters for Lennard-Jones interactions and the dihedrals using a mixture of experimental (such as heats of vaporizations) and quantum mechanical information.
-For a polarizable force fields, this process has to be redone as in the first step the polarizable contribution has to be separated from the "static" one.
-Furthermore, for the polarization models considered here, an additional cutoff/damping scheme needs to be introduced to prevent polarization catastrophe~\cite{Thole1981}---a phenomenon where close-by dipoles feed of each others and their dipole moments (and energy) grows into infinity.
-
-Both polarizable lipid force fields considered here use a bottom up approach where parameters are first obtained for smaller molecules or molecular fragments from which the lipid models are then built.
-In the CHARMM-Drude model, the polarizabilities, the Thole damping factors used to prevent polarization catastrophe, and the partial charges where determined by fitting to a discretized QM potential energy surface around the molecule.
-More specifically, the perturbations of the potential energy surface in the presence of a small charge where observed to separate the dynamic polarizability contribution.
-They then proceeded to use enthalpies of vaporation along with QM calculations of molecular interaction energies to obtain the Lennard-Jones parameters, and on top of that dihedral potentials were introduced to reproduce the QM potential energy profiles along a dihedral rotation.
-Finally, the dihedrals of the lipid head groups where refined targetting the NMR order parameters.
-
-In the AMOEBA force field, a set of atomic multipoles were first obtained based on the QM electron densities using Stone's distributed multipole analysis~\cite{Stone1981}.
-The atoms were then assigned polarizabilities according to the AMOEBA atom types and induced dipole contribution was self consistently subtracted from the permanent dipoles~\cite{shi2013proteinamoeba}.
-The charges (monopoles) were then kept constant while the multipoles were refined against the electrostatic potential energy surface around few conformers of the molecular fragments.
-%Multipoles are allowed to induce only dipoles at polarizable sites outside their immediate environment (polarization group).
-After electrostatic parameters were set, the Lennard-Jones parameters were taken from previous AMOEBA parameterizations of smaller molecules containing same functional groups~\cite{Ren2011polorganic,shi2011hydration}, and then refined against molecule - water (gas phase) interaction energies.
-After settling all the nonbonded parameters, dihedral potentials were parameterised based on the QM conformational energy profiles.
+%this is not needed for jctc
+\section{Author contributions}
+H.S.A. analyzed the correlation and relaxation times from the simulations and wrote the first draft of the manuscript.
+S.D.
+assisted in generating the initial systems setups for the simulations of POPC and POPE bilayers using AMOEBA based force field (including identifying errors in the force field files received), and wrote the initial scripts required to generate OpenMM and Tinker suitable file formats of the simulation systems.
+J.J.M.
+contributed to early testing (debugging, setup, and running) of simulations with the AMOEBA based force field using Tinker9 and Tinker-OpenMM as well as manuscript editing.
+B.K. conceptualized and initiated the project, did the literature research, created and ran the simulations, analyzed the results, and wrote the first draft of the manuscript.
+O.H.S.O made the quality evaluation using the NMRlipids databank, and assisted in conceptualizing the project and in writing the manuscript. 
 \end{comment}
-\begin{comment}
-- polarization contribution has to be separated from point charge electrostatics in the parametrization process.
-- for the models considered here (Drude, IDP) one has to prevent the polarization catastrophe somehow (Thole damping parameter)
-- Brief explanation how the two FFs studied here were built.
-
-Notes on CHARMM-Drude parametrization:
-- bottom up approach. Permantent charge, polarizability and screening factors from QM. Polarizability decoupled from permanent by fitting to changes in PES? 
-- Non-bonded electrostatics parameters calculated based on QM.
-- LJ from enthalpies of vaporization together .with QM data
-- Dihedrals from QM scans for model compounds, then refined to according to the order parameters of the headgroup using re-weighted MD and an genetic algorithm.
-
-Notes on AMOEBA parametrization (this is NOT well explained in the paper)
--bottom up approach
--Stone distributed multipole analysis to get the multipoles from the QM electron densities. Divided into permanent and induced.
--Then iterative fitting to QM-PES for electrostatics
--WdW from 1) optimization to hydration energy calculations and 2) QM fragment-water gas phase calculations.
--dihedrals optimized to QM potentials
--no mention of the screening for polarization? Do they use some AMOEBA value
-\end{comment}
-\subsection{Using a polarizable force field for membrane simulations}
-
-While non-polarizable MD simulations of membranes can be nowadays routinely performed with several simulation packages and force fields, polarizable simulations bear many practical complications.
-%Secondly, there are compability issues with the most used MD engines and the polarizable models and that support might be version dependent. 
-Regarding currently popular MD simulation packages for membrane simulations, OpenMM supports both AMOEBA and Drude force fields, NAMD can only run the Drude force field whereas, GROMACS
-%---the most popular MD engine within the bilayer simulation field---
-only has limited support for the Drude polarizable force field via an unofficial Git-branch~\cite{drudegithub}. %and no support for AMOEBA. 
-TINKER is widely used with AMOEBA, but it does not support semi-isotropic pressure coupling, which %decouples the pressure tensor in $xy$ and $z$-direction and is 
-is required in membrane simulations~\cite{xxx}.
-%crucial for correctly modelling membrane systems~\cite{xxx}.
-%
-Therefore, we have selected to use OpenMM in all simulations in this work.
-%In general, easy access to MD engines, as well as adaptability of common analysis tools and libraries for polarizable force fields, will be crucial in facilitating more wide-spread utilization of these models.
-CHARMM-GUI can generate the topology and input parameters for the Drude force field, which greatly simplifies the employment of this model in particular~\cite{kognole2022charmm}.
-Standard protocols for AMOEBA lipid parameters are not available, but some parameters are available online~\cite{amoebagithub,klesse2020induced}.
-
-%Considering the increased interest in polarizable force fields in membrane simulations, we mention here some hurdles in their practical usage. Firstly, 
-Another practical issue to be considered in polarizable membrane simulations is the increased computational cost arising from increased level of details obtained by explicit treatment of electronic polarizability.
-For the Drude-based models, this slowdown occurs both because the addition of the Drude particles increases the number of interaction pairs and the employed extended dual-Langevin thermostat requires a shorter 1~fs integration time step, compared to the 2~fs typically used for non-polarizable membrane simulations.
-The AMOEBA force field can use a multi time-step integration algorithm where the non-electronic interactions are iterated with a 2~fs time step whereas more computationally unstable polarization terms are iterated with a smaller time step.
-However, the multi-time step scheme  only partly mitigates the computational cost: one typically experiences almost a ten-fold decrease in the simulation speed while using the polarizable force fields.
-
-All simulations performed in this work are listed in Table~\ref{table:sim_details} with links to the openly available trajectory data.
-Data not mentioned here was obtained by analysing pre-existing trajectories from the NMRlipids databank.
-
-\subsection{Simulations with CHARMM-Drude parameters}
-
-CHARMM-Drude2017 simulations were performed with OpenMM 7.5.0~\cite{eastman2017openmm} using parameters extracted with \textit{Membrane Builder}~\cite{wu2014charmm,jo2009charmm,jo2007automated,lee2018charmm} and \textit{Drude Prepper}~\cite{kognole2022charmm} from CHARMM-GUI~\cite{jo2008charmm,lee2016charmm}.
-Before starting the simulations, membrane structures were equilibrated for 200~ns using the non-polarizable CHARMM36 force field~\cite{klauda2010update}, and the last frames of these simulations were used to generate the starting structures for the polarizable force field simulations.
-
-CHARMM-Drude2023 force field parameters are currently not integrated into CHARMM-GUI.
-Therefore, the simulation setups with NaCl and CaCl$_{2}$ have been generated following the instructions in the original CHARMM-Drude2023 paper~\cite{yu2023drude} using CHARMM program~\cite{brooks2009charmm} and the last frame of 200~ns long CHARMM36 simulations (same as used for CHARMM-Drude2017).
-CHARMM-Drude2023 simulations without any ions have been directly obtained from Zenodo~\cite{richard_m_venable_2023_7872447, richard_m_venable_2023_7871949}.
-
-A dual Langevin thermostat was employed to keep the Drude particles and the rest of the system at 1.0~K and 303~K, respectively.
-A Drude-hardwall of 0.02~nm was utilized to keep the Drude particles close to their parent atoms.
-Semi-isotropic Monte Carlo barostat was used to couple pressure to 1~bar independently in membrane plane and normal directions.
-%while the pressure fluctuations along the $xy$-plane were coupled.
-A constant surface tension of 0.0~dyne/cm was applied throughout the simulations.
-Bonds containing the hydrogens were constrained.
-For CHARMM-Drude2017, particle mesh Ewald was used to compute the Coulomb interactions.
-Van der Waals interactions have been cut to 0 between 1.0~nm and 1.2~nm using a switching function.
-For the CHARMM-Drude2023, Lennard-Jones particle-mesh Ewald (LJ-PME) method was used to compute the long-range dispersions~\cite{wennberg2013lennard}.
-Simulation frames have been saved in every 10~ps.
-
-\subsection{Simulations with AMOEBA parameters}
-
-AMOEBA force field parameters for OpenMM were obtained from GitHub~\cite{amoebagithub,klesse2020induced}.
-All AMOEBA simulations have been run on OpenMM 7.1.1~\cite{eastman2017openmm}.
-The same initial structures as the CHARMM-Drude simulations have been used to create the simulation setups.
-A multi-time step Langevin integrator was used to iterate the bonded and non-bonded interactions in every 0.5~fs and 2.0~fs, respectively.
-A non-bonded cutoff of 1.2~nm has been applied while semi-isotropic Monte Carlo barostat was used to couple pressure to 1~bar independently in membrane plane and normal directions.
-%keeping the pressure at 1~bar using a Monte Carlo barostat.
-Pressure fluctuations in the $xy$-plane has been coupled using a semi-isotropic pressure coupling and a constant surface tension at 0.0~dyn/cm has been applied throughout the simulations.
-Simulation frames have been saved in every 10~ps. Further simulations details can be found in the input files of the respective simulations (Table~\ref{table:sim_details})
-
-\subsection{Analysis of simulations}
- All simulations were first uploaded to the NMRlipids databank~\cite{Databank}. 
-Area per lipids, X-ray scattering form factors, and order parameters ($S_{CH}$) are automatically calculated by the NMRLipids Databank~\cite{Databank} and were extracted from therein.
-Quality evaluation metrics were quantified as detailed in the NMRlipids databank~\cite{Databank} with the exception that POPC simulations at 303\,K were coupled with the experimental data measured at 300\,K, while in the NMRlipids databank simulations are coupled to experiments with the maximum temperature difference of two degrees.
-Shortly, order parameter qualities, $P^{\mathrm{headgroup}}$, $P^{sn-1}$, and $P^{sn-2}$, presents an average probability for order parameters of a given segment to locate within experimental values, taking the error bars of both values into the account.
-Qualities of SAXS form factors, $FF_{q}$, are measured from the distance between first form factor minimum in simulated vs.
-experimental curve to avoid the effects from the simulation size dependency of the SAXS data.
-%~\cite{Databank} one should focus on the extrema $q_{z}$[\AA$^{-1}$] when comparing the simulated and experimental data.
-In particular $FFq$ only measures the difference between the first minima of the simulated and experimental curves.
-Note that in quantifying the bilayer electron densities for calculation of SAXS curves  the NMRlipids databank analysis algorithm places electrons as point charges at the atom centers without considering redistribution of charge density due to the polarizability.
-Nevertheless, we expect that this approximation does not have significant effect on the resulting SAXS form factors.
-
-The $R_{1}$ rates and effective correlation times, along with the accompanying error estimates, are quantified from the trajectories as elaborated in Ref.~\citenum{Antila2021}.
-In-house python script available at XXX was used for this purpose.
-Source codes for all scripts and a detailed methodological explanation of the computed values are available in the referenced cited herein.
-
-\begin{comment}
-Order parameters are calculated using the script "OrderParameters.py" under\\ "https://github.com/batukav/NMRlipidsVIpolarizableFFs/tree/master/scripts". \\
-The order parameters are first calculated for each lipid as an average over time, and then averaged over the lipids.
-The error estimates are calculated as the standard error of the mean from the second average.
-Dihedral distributions are calculated using the script "calcDihedral.py" under\\ "https://github.com/batukav/NMRlipidsVIpolarizableFFs/tree/master/scripts" .\\
-The $R_{1}$ rates and effective correlation times, along with the accompanying error estimates, are quantified from the trajectories as elaborated in Ref.~\citenum{Antila2021}.
-In-house python script available at XXX was used for this purpose.
-\subsection{Simulation Details and data availability}
-This project is performed using the "NMRLipids Databank" format and all related files, except the trajectories, are available under\\ "https://github.com/batukav/NMRlipidsVIpolarizableFFs/"
-\end{comment}
-
-\newpage
-%\begin{adjustbox}{angle=90}
-\begin{table}[]
-\begin{small}
-\begin{tabular}{cccccccccc}
-	Lipid/ion                & force field  & Ion (M) & N$_{l}$ & N$_{w}$ & N$_{ion}$ & T (K) & t$_{sim}$ (ns) & t$_{analysis}$ (ns) & files \\ \hline
-\multirow{2}{*}{POPC}            & Drude2017 & 0      & 144       & 6400       & 0         & 303    & 500              & 400         &          \cite{kav_batuhan_2021_7607436}    \\ 
-                                 & Drude2023 & 0      & 72       & 2239       & 0         & 303      & 300              & 200         &          \cite{kav_batuhan_2023_7916287, richard_m_venable_2023_7871949}    \\ \hline
-\multirow{3}{*}{POPE}                              & Drude2017 & 0      & 144       & 6400       & 0         & 308.15    & 350              & 300         &          \cite{kav_batuhan_2021_7604627}    \\                                            & Drude2023 & 0       & 72        & 2304      &  0         & 303    & 300             & 200         &           \cite{richard_m_venable_2023_7872447,kav_batuhan_2023_7916494} \\
-                                                  & AMOEBA     & 0       & 72        & 2880      &  0        & 303     & 305.94          & 305.94      & \cite{kav_batuhan_2023_7622838} \\ \hline
-	\multirow{5}{*}{POPC:NaCl}        & Drude2017 & 0.350   & 144      & 6400     & 41         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_7586915}   \\
-				  & Drude2017 & 0.450   & 144      & 6400     & 51         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_7591753}   \\
-				  & Drude2017 & 0.650   & 144      & 6400     & 77         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_7596011}   \\
-				  & Drude2017 & 1.0     & 144      & 6400     & 115        & 303   & 500             & 400                  & \cite{kav_batuhan_2020_7600326}   \\ 
-                    & Drude2023 & 1.0      & 128       & 6400      & 41          & 303    & 223.97      &       220.14       & \cite{kav_batuhan_2023_8000095} \\     
-                    & Drude2023 & 1.0      & 128       & 6400      & 115          & 303    & 220.14      &       220.14       & \cite{kav_batuhan_2023_8000133} \\ \hline
-
-	\multirow{5}{*}{POPC:CaCl$_{2}$} & Drude2017 & 0.350   & 144      & 6400     & 41         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_7600827}   \\
-				  & Drude2017 & 0.450   & 144      & 6400     & 52         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_7605016}   \\
-				  & Drude2017 & 0.650   & 144      & 6400     & 76         & 303   & 500             & 400                  & \cite{kav_batuhan_2020_7604040} \\
-				  & Drude2017 & 1.0   & 144      & 6400     & 114         & 303   & 500             & 400                  & \cite{kav_batuhan_2023_7658975}\\
-                    & Drude2023 & 0.350 & 128      & 6400     & 41          & 303   & 219.59          & 219.59               & \cite{kav_batuhan_2023_8000065}\\
-                    & Drude2023 & 0.790 & 128      & 6400     & 91          & 303   & 214.85          & 214.85               & \cite{kav_batuhan_2023_7992137} \\ \hline
-DOPC           & AMOEBA   & 0 & 72 & 2880 & 0 & 303 & 201.61 & 201.61 & \cite{kav_batuhan_2022_7604681} \\ \hline
-\multirow{2}{*}{DOPC:NaCl} & AMOEBA & 0.450 & 72 & 2880 & 17 & 303 & 218.41 & 218.41 & \cite{kav_batuhan_2022_7604711} \\
-                           & AMOEBA & 1.0 & 72 & 2880 &  35 & 303 & 201.61 & 201.61 & \cite{kav_batuhan_2023_7625844} \\ \hline
-\multirow{2}{*}{DOPC:CaCl$_{2}$} & AMOEBA  & 0.450 & 72 & 2880 & 16 & 303 & 218.41 & 218.41 & \cite{kav_batuhan_2022_7604842} \\
-                                 & AMOEBA & 1.0 & 72 & 2880  & 36 & 303 & 218.41 & 218.41 & \cite{kav_batuhan_2022_7604810}
-
-\end{tabular}
-\end{small}
-\caption{Systems simulated for this work.
-Column N$_l$ gives the number of lipids, N$_w$ gives the number of water molecules, T(K) denotes the temperature in kelvins.
-Simulated time is listed in column t$_{sim}$ and time used for analysis in t$_{analysis}$.
-Column "files" gives the reference to open access simulation data.
-The salt concentrations calculated as [salt]=N$_c$×[water] / N$_w$, where [water] = 55.5 M}
-\label{table:sim_details}
-\end{table}
-\todo{check the simualation settings (especially the temperature}
-%\end{adjustbox}
 
@@
 \begin{acknowledgement}
-
 J.J.M. thanks Huiying Chu (from the Li lab) for technical discussions and Sameer Varma for useful suggestions.
+B.K.
+thanks Huiying Chu and Guohou Li for providing AMOEBA based force field parameters and technical discussions.
+B.K also thanks Gianni Klesse for providing the AMOEBA based force field parameters in OpenMM format and technical help with running the simulations.
+O.H.S.O. acknowledges CSC -- IT Center for Science for computational resources and the Research Council of Finland for funding (grant nos. 315596, 319902, 345631 \& 356568).
 
@@
 \begin{suppinfo}
+Supporting Information: NMR order parameters and SAXS form factors for additional CHARMM-Drude2023 replicas and dihedral angle distributions for the headgroup from the studied polarizable models and CHARMM36.
 \end{suppinfo}
@@
 
-\section{Supplementary figures}
-
-\begin{figure}[!h]
-    \centering
-    %\includegraphics{Figures/order_parameters.png}
-    \includegraphics[width=0.8\textwidth]{Figures/quality_Drude2023_replicas.pdf}
-    \caption{X-ray scattering form factors, and C-H bond order parameters for headgroup and glycerol backbone, and acyl chains (from top to bottom) compared between simulations and experiments using the NMRlipids databank.
-The atom labelling is as depicted in Fig.~\ref{fig:molecules}.}
-    \label{fig:order_parameters_replicas}
-
-\end{figure}
-
-\begin{figure}[!hbt]
-	\centering
-	\includegraphics{Figures/dihedral_distributions_without_ions.eps}
-	\caption{The distributions of the torsion angles for the head group atoms.
-For each torsion angle, the upper and lower rows contain the DOPC (AMOEBA) - POPC (Drude/Drude2023) and POPE lipids, respectively.}
-	\label{fig:dihedral_no_salt}
-\end{figure}
-
 \end{document}
